### 2017-07-12 seminar_cor/seminar_cor_haegele.tex
renamed to seminar_cor_heagele.tex, continued with section 2

%% bare_jrnl_compsoc.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% Computer Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[10pt,journal,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\usepackage{graphicx}
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

% my own stuff
\newcommand{\xray}{X-ray~}
\newcommand{\xrays}{X-rays~}
\usepackage{cleveref}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Determination of Center of Rotation \\ in Computed Tomography}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{David H{\"a}gele}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Seminar on Image Processing - Algorithms and Architectures - Univerity of Stuttgart - IPVS/PAS - SS17}%
{H{\"a}gele: Determination of Center of Rotation in Computed Tomography}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Computer Science, Image Processing, Computed Tomography, Tomographic Reconstruction
\end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{I}{n} computed tomography (CT), a set of set of projections is acquired by meassuring the attenuation of \xrays through the scanned object at different rotation angles.
The \xrays are emmited by a source on one side of the scanned object, intersect with the object, and hit a detector on the other side.
The resulting meassurement of the detector is a projection of the object's internals.
Rotating the object and repeating the process results in a set of projections, which can be used to reconstruct an image of the internal structure of the scanned object. 
For this reconstruction it is mandatory to know the center of rotation (COR), or rather the location of its projection, as slight deviations from the actual COR produce severe artifacts in the reconstructed image.
In general, the COR is assumed to be projected onto the center of the detector, but this assumption can be violated on badly calibrated CT systems. 
A correct calibration can also become almost impossible when the \xray source is very close to the scanned object, as is the case for industrial scanners with resolutions on the micrometer scale \cite{azevedo90}.
In this paper, different techniques and algorithms for the determination of the center of rotation are explained, which can be found in the literature.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.
% However, the Computer Society has been known to put floats at the bottom.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Prerequisites}
In order to explain the problem of inaccurate COR assumption in tomographic reconstruction, some prerquisite knowledge about CT will be briefly explained in this section.

\subsection{Image Acquisition in CT}
This subsection will briefly explain how images are typically aquired in CT.

In \xray imaging, an \xray source (e.g. \xray tube) is used to generate photons which then run through an object.
The capability of the photons to penetrate matter, depends on the material, which means that the decrease in number of photons along the incident beam is different for different materials.
The decrease in the number of photons, or attenuation, results from the photons being absorbed or scattered as they travel through the matters~\cite{Buzug2008_chap1}.
Using an \xray detector, the photon count after matter penetration can be measured and thus the attenuation along the beam from the source to the detector.
The measured quantity can be interpreted as the projection of all attenuations along the ray.

However, a single projection is not sufficient to deduce the attenuation strengths within the object. 
To locate regions of high and low attenuation, multiple projections from different sides of the object are required.
Therefore the object is placed on a rotary table or the imaging aparatus (\xray source and detector) is rotated around the object.

\subsection{CT Geometries}
This subsection will briefly explain the most common geometries in CT, that is the setup of \xray source and detector and the ray projections which result from it.

\subsubsection{Parallel Beam}
The simplest geometry in CT (from a mathematical point of view) is the parallel beam geometry.
Here all rays are parallel that are emitted from the same side of the object.
\Cref{fig:parallelbeam} shows a schematic of this geometry with the resulting signal.
%
\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{img/parallel_beam_geometry}
\caption{Parallel beam geometry with the source moving perpendicular to the ray direction for sampling a 1-dimensional image (graph). \cite{Buzug2008_chap4}}
\label{fig:parallelbeam}
\end{figure}
%
This projection geometry can be mathematically expressed with the Radon transform.
When we think of the object as a function of attenuation coefficients $f(x,y)$ and the measured signal as a function of projected coefficients $R(r,\alpha)$ where $r$ is the position on the detector and $\alpha$ is the rotation angle at which the signal is obtained, then
\begin{equation}
R(r,\alpha) = \int_{-\infty}^{\infty} f(r\;cos\;\alpha + t\;sin\;\alpha, r\;sin\;\alpha-t\;cos\;\alpha)\;dt.
\end{equation} 
The parameter $t$ is the position on the ray that hits the detector at $r$ and angle $\alpha$, which makes it clear that the projection $R(r,\alpha)$ is the integral along a ray through the object $f(x,y)$.

\subsubsection{Fan Beam}

\subsubsection{Cone Beam}

\subsection{Reconstruction}

\subsubsection{Filtered Back Projection}

\section{COR Determination}

\subsection{Sinogram Based Methods}

\subsubsection{High Density Feature}

\subsubsection{Center of Mass}

\subsubsection{Cross-Correlation}

\subsection{Reconstruction Based Methods}

\subsubsection{Integral of Negativity}

\subsubsection{Histogram Entropy}

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{}

% use section* for acknowledgment
%\ifCLASSOPTIONcompsoc
%  % The Computer Society usually uses the plural form
%  \section*{Acknowledgments}
%\else
%  % regular IEEE prefers the singular form
%  \section*{Acknowledgment}
%\fi

%The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{bibliography}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}

### 2017-07-13 seminar_cor/seminar_cor_haegele.tex
added ligature package

@@
 \usepackage{cleveref}
+\usepackage{cmap}
+\usepackage{csquotes}
 
@@
 
+Unfortuantely, parallel beam geometry is impractical as a pencil beam \xray source (source emmiting only one ray in a specific direction) would have to be moved parallel to the detector to collect all samples for a single rotation angle, which is very time consuming.
+However, parallel beam geometry occurs at synchrotron facilities \cite{something}.
+
 \subsubsection{Fan Beam}

### 2017-07-14 seminar_cor/seminar_cor_haegele.tex
new images, more content (started fan beam)

@@
 Here all rays are parallel that are emitted from the same side of the object.
-\Cref{fig:parallelbeam} shows a schematic of this geometry with the resulting signal.
+\Cref{subfig:parallelbeam} shows a schematic of this geometry with the resulting signal.
 %
@@
 \centering
-\includegraphics[width=\linewidth]{img/parallel_beam_geometry}
-\caption{Parallel beam geometry with the source moving perpendicular to the ray direction for sampling a 1-dimensional image (graph). \cite{Buzug2008_chap4}}
-\label{fig:parallelbeam}
+\subfloat[Parallel Beam]{\includegraphics[width=0.58\linewidth]{img/parallel_beam_geometry}%
+\label{subfig:parallelbeam}}
+\hfil
+\subfloat[Fan Beam]{\includegraphics[width=0.4\linewidth, trim=0 -6cm 0 0]{img/fanbeamgeometry}%
+\label{fig_second_case}}
+\caption{Parallel beam and fan beam geometry.
+In parallel beam the source moving perpendicular to the ray direction for sampling a 1-dimensional image (graph).
+In fan beam, rays are not perpendicular but instead emmited in fan blade form.
+$\gamma$ is the rotation angle at which the signal is obtained.
+\cite{Buzug2008_chap3,Buzug2008_chap4}}
+\label{fig:parallalandfanbeam}
+\end{figure}
+
+This projection geometry can be mathematically expressed with the Radon transform.
+When we think of the object as a function of attenuation coefficients $f(x,y)$ and the measured signal as a function of projected coefficients $R(r,\gamma)$ where $r$ is the position on the detector and $\gamma$ is the rotation angle at which the signal is obtained, then
+\begin{equation}
+R(r,\gamma) = \int_{-\infty}^{\infty} f(r\;cos(\gamma) + t\;sin(\gamma), r\;sin(\gamma)-t\;cos(\gamma))\;dt.
+\end{equation} 
+The parameter $t$ is the position on the ray that hits the detector at $r$ and angle $\gamma$, which makes it clear that the projection $R(r,\gamma)$ is the integral along a ray through the object $f(x,y)$.
+The sampling of an object in this fashion, results in a so called sinogram.
+The contribution of a specific area of the object to the individual projections, follows a sinusoidal path along the $\gamma$-axis as can be seen in \cref{fig:sinogram_parallel}
+%
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{img/phantomandradon.png}
+\caption{
+A \enquote{smilie} phantom and its corresponding sinogram with $\gamma~\in~[0^\circ,360^\circ]$.
+An imaginary CT would have started with source on top and detector on bottom and then rotated clockwise around the phantom.
+}
+\label{fig:sinogram_parallel}
 \end{figure}
 %
-This projection geometry can be mathematically expressed with the Radon transform.
-When we think of the object as a function of attenuation coefficients $f(x,y)$ and the measured signal as a function of projected coefficients $R(r,\alpha)$ where $r$ is the position on the detector and $\alpha$ is the rotation angle at which the signal is obtained, then
-\begin{equation}
-R(r,\alpha) = \int_{-\infty}^{\infty} f(r\;cos\;\alpha + t\;sin\;\alpha, r\;sin\;\alpha-t\;cos\;\alpha)\;dt.
-\end{equation} 
-The parameter $t$ is the position on the ray that hits the detector at $r$ and angle $\alpha$, which makes it clear that the projection $R(r,\alpha)$ is the integral along a ray through the object $f(x,y)$.
 
@@
 \subsubsection{Fan Beam}
+A more practical geometry is the fan beam, in which a source emits a flat beam that extends to the sides (like a fan blade).
 

### 2017-07-14 seminar_cor/seminar_cor_haegele.tex
new phantom, also fanbeam sinogram

@@
 \subfloat[Fan Beam]{\includegraphics[width=0.4\linewidth, trim=0 -6cm 0 0]{img/fanbeamgeometry}%
-\label{fig_second_case}}
+\label{subfig:fanbeam}}
 \caption{Parallel beam and fan beam geometry. 
@@
 $\gamma$ is the rotation angle at which the signal is obtained. 
-\cite{Buzug2008_chap3,Buzug2008_chap4}}
+Figures from~\cite{Buzug2008_chap3,Buzug2008_chap4}}
 \label{fig:parallalandfanbeam}
@@
 A \enquote{smilie} phantom and its corresponding sinogram with $\gamma~\in~[0^\circ,360^\circ]$.
-An imaginary CT would have started with source on top and detector on bottom and then rotated clockwise around the phantom.
+An imaginary CT would have started with source on top and detector on bottom and then rotated clockwise around the phantom to obtain the shown sinogram.
 }
@@
 A more practical geometry is the fan beam, in which a source emits a flat beam that extends to the sides (like a fan blade). 
+\Cref{subfig:fanbeam} shows a schematic of this geometry.
+Here the projection for a single rotation angle can be obtained at once.
 

### 2017-07-15 seminar_cor/seminar_cor_haegele.tex
fan beam done, cone beam almnost

@@
 
-Unfortuantely, parallel beam geometry is impractical as a pencil beam \xray source (source emmiting only one ray in a specific direction) would have to be moved parallel to the detector to collect all samples for a single rotation angle, which is very time consuming.
+Unfortuantely, parallel beam geometry is impractical as a pencil beam \xray source (source emmiting only one ray in a specific direction) would have to be moved parallel to the detector to collect all samples for a single rotation angle, which is very time consuming \cite{Buzug2008_chap7}.
 However, parallel beam geometry occurs at synchrotron facilities \cite{something}.
@@
 Here the projection for a single rotation angle can be obtained at once. 
+The geometry can be defined by the COR it rotates around, the offset of the source to the COR and its aperture angle.
+In clinical practise curved detector rays are used, as shown in \cref{subfig:fanbeam}, to get constant aperture angle steps for each detector cell.
+%
+\begin{figure}[!h]
+\centering
+\subfloat[]{\includegraphics[width=0.43\linewidth]{eps/rebinning.eps}%
+\label{subfig:rebinning1}}
+\hfil
+\subfloat[]{\includegraphics[width=0.51\linewidth]{eps/rebinning_geo.eps}%
+\label{subfig:rebinning2}}
+\caption{
+On the left a schematic is shown with two fan beams with aperture angle $\varphi$ at different rotation angles.
+It can be sen that the outer rays of the two are parallel.
+The maximum distance of two parallel beams is limited by $\varphi$, and the maximum rotation angle $\theta$ at which two
+parallel rays can be obtained is given by $\theta = \psi_{min}+\psi_{max}$ which are the angles from central to outer rays.
+On the right a schematic is shown that points out the correspondence between fan and parallel beam geometry.
+It can be seen that for the red ray which is the ray at rotation angle $\theta$ with deviation angle $\psi$ from the central ray, there exists an angle $\gamma$ in parallel beam geometry so that it is perpendicular to the $r$-axis (i.e.
+parallel beam detector).
+Figures from~\cite{Buzug2008_chap7}
+}
+\label{fig:rebinning}
+\end{figure}
+%
+
+As each ray projection for a single rotation angle in this geometry corresponds to a different ray directions (i.e.
+rays are not parallel here), the resulting sinogram differs from the sinogram of parallel beam geometry.
+When analyzing this geometry further, it can be seen that the ray projections of a parallel beam geometry are inherent and can be found at different rotation angles.
+Through a process called rebinning, the sinogram of the fan beam can be transformed to parallel beam.
+\Cref{fig:rebinning} illustrates this concept.
 
 \subsubsection{Cone Beam}
+The previously introduced geometries are both 2-dimensional (2D) as they described projections within an x-y-plane onto a one dimensional detector.
+To obtain a 3-dimensional (3D) image the scanned object would have to be moved perpendicular to the plane to obtain multiple slice images.
+With the 3D cone beam geometry, this is not necessary as it uses a cone shaped beam that projects onto a 2D flat panel detector.
+%
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{img/conebeam_centralslice.png}
+\caption{
+A cone beam setup with an object on a rotary table.
+It can be seen that only the projection of the central slice is equivalent to a projection in fan beam geometry.
+Other projections that are parallel to the s-axis of the detector use samples at multiple heights of the object.
+Also the projection of the rotation axis onto the detector is shown which in this case is offset from detectors
+z-axis.
+Figure from \cite{Yang2011}
+}
+\label{fig:conebeam}
+\end{figure}
+%
+In \cref{fig:conebeam} it can be seen that there exists a certain line on the detector which will receive the same input as a fan beam detector.
+All other lines read projections through multiple planes of the object and are thus not rebinnable to parallel beam.
 

### 2017-07-17 seminar_cor/seminar_cor_haegele.tex
more content on reconstruction

@@
 \begin{equation}
+\label{eq:radon}
 R(r,\gamma) = \int_{-\infty}^{\infty} f(r\;cos(\gamma) + t\;sin(\gamma), r\;sin(\gamma)-t\;cos(\gamma))\;dt.
@@
 \subsection{Reconstruction}
+This subsection will explain the process of reconstructing an image from the projections obtained in CT.
+Only the 2D case will be explained as it is relevant for understanding the COR determination techniques later on.
+Also this subsection will only cover Fourier based reconstruction, especially filtered back projection.
+
+The problem of reconstruction can be stated as follows.\\
+Given the Radon transform from \cref{eq:radon} $R(r,\gamma)$ how do we calculate the original signal $f(x,y)$?
+The solution to this would be the inverse Radon transform $R(r,\gamma)^{-1}$.
+To solve this, the Fourier slice theorem can be leveraged.
+%
+\begin{align}
+p(x) &= R(x,0) = \int_{-\infty}^{\infty} f(x,y)\;dy. 
+\\
+\hat{f}(u,v) &= \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}f(x,y)e^{-2\pi i (xu+yv)}\;dx\;dy.
+\\
+s(u) &= \hat{f}(u,0) = \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}f(x,y)e^{-2\pi i xu}\;dx\;dy
+\label{eq:fourierslice}
+\\
+&= \int_{-\infty}^{\infty}\left[\int_{-\infty}^{\infty}f(x,y)\;dy\right]\;e^{-2\pi i xu}\;dx
+\nonumber
+\\
+&= \int_{-\infty}^{\infty}p(x)e^{-2\pi i xu}\;dx = \hat{p}(u).
+\nonumber
+\end{align}
+%
+The proof of the Fourier slice thoerem above (\cref{eq:fourierslice}) states that the Fourier transform $\hat{p}(u)$ of a projection $p(x)$ of $f(x,y)$, is the same as the slice $s(u)$ (parallel to the projection line and through the center) of the Fourrier transform $\hat{f}(u,v)$ of $f(x,y)$.
+In short $s(u) = \hat{p}(u)$.
+
+As th Radon transform is a set of projections the Fourier slice theorem can be used to build the Fourier tranform of the original signal and reconstruct it using the inverse Fourier transform.
+\Cref{fig:inverseradon} illustrates this method.
+%
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\linewidth]{img/inverseradon.png}
+\caption{
+The inverse Radon transform can be calculated using the Fourier slice theorem.
+First all projections are transformed to the Fourier space along the $\xi$-axis ($=r$-axis).
+Then the Fourier transform is built by placing all transformed projections according to their angle into a polar coordinate system.
+The coordinate system thus consists of slices of the Fourier transform of the original signal which can then
+be obtained using an inverse Fourier transform.
+Figure from \cite{Buzug2008_chap5}
+}
+\label{fig:inverseradon}
+\end{figure}
+%
+The problem of applying this algorithm in practise is, that due to the limited angular resolution (i.e.
+discretization) of the Radon transform the frequency information density in the Fourier space decreases with for higher frequencies \cite{Buzug2008_chap5}.
+In the process of cartesian regridding of the polar coordinates a lot of data for the high frequencies would have to be made up in order to perfrom an inverse fast fourier transform.
+Therefore a different algorithm, the filtered back projection, is used in practise.
 

### 2017-07-17 seminar_cor/seminar_cor_haegele.tex
finished prerequisites

@@
 The capability of the photons to penetrate matter, depends on the material, which means that the decrease in number of photons along the incident beam is different for different materials.
-The decrease in the number of photons, or attenuation, results from the photons being absorbed or scattered as they travel through the matters~\cite{Buzug2008_chap1}.
+The decrease in the number of photons, or attenuation, results from the photons being absorbed or scattered as they travel through the matter~\cite{Buzug2008_chap1}.
 Using an \xray detector, the photon count after matter penetration can be measured and thus the attenuation along the beam from the source to the detector.
@@
 The problem of applying this algorithm in practise is, that due to the limited angular resolution (i.e.
-discretization) of the Radon transform the frequency information density in the Fourier space decreases with for higher frequencies \cite{Buzug2008_chap5}.
+discretization) of the Radon transform the frequency information density in the Fourier space decreases for higher frequencies \cite{Buzug2008_chap5}.
 In the process of cartesian regridding of the polar coordinates a lot of data for the high frequencies would have to be made up in order to perfrom an inverse fast fourier transform.
@@
 \subsubsection{Filtered Back Projection}
+The back projection algorithm reconstructs the original image from a discrete radon transform (i.e. parallel beam sinogram).
+For the reconstruction of a single pixel, the information has to be gathered from the ray projections of the rays that intersect this pixel (see \cref{eq:backprojection}).
+\begin{equation}
+\label{eq:backprojection}
+g(x,y) = \int_0^\pi R(x\;cos(\gamma)+y\;sin(\gamma),\gamma)\;d\gamma. 
+\end{equation}
+
+This can be imagined as smearing the projections over a canvas in their respective directions.
+At first glance this may seem as the correct inversion of the projections but is not.
+This becomes clear when thinking about the original image being a dirac delta (image with all entreis zero except for the central pixel).
+The projections of this image will be 1D dirac deltas, but using the back projection algorithm will not
+restore the 2D dirac but instead smear the 1D diracs also into areas which are supposed to be zero.
+In fact the resulting image looks quite blurry which is due to the back projection equaling to the original image convolved with the point spread function of the imaging system (i.e.
+the Radon transform) $\frac{1}{|(x,y)|}$ \cite{Buzug2008_chap5}.
+To counter this, the projections first need to be filtered with a high pass filter.
+This introduces negative values in the projections which in the back projection process cancel out positive contributions of other projections.
+As amplifying high frequencies also aplifies noise, there are different filters that can be used in practise \cite{farquhar1998}.
 

### 2017-07-19 seminar_cor/seminar_cor_haegele.tex
started COR section

@@
 This introduces negative values in the projections which in the back projection process cancel out positive contributions of other projections.
+Figure \cref{fig:backprojection} shows the results of unfiltered and filtered back projection.
 As amplifying high frequencies also aplifies noise, there are different filters that can be used in practise \cite{farquhar1998}.
+%
+\begin{figure}[!h]
+\centering
+\subfloat{\includegraphics[width=0.18\linewidth]{img/bp_02}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/bp_04}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/bp_08}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/bp_16}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/bp_32}}
+\\
+\subfloat{\includegraphics[width=0.18\linewidth]{img/fbp_02}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/fbp_04}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/fbp_08}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/fbp_16}}
+\hfil
+\subfloat{\includegraphics[width=0.18\linewidth]{img/fbp_32}}
+\caption{
+On the top row, the results of unfiltered back projection are shown. 
+From left to right the number of projections used increases from 2 to 32.
+The resulting image looks blurry.
+On the bottom row the results of filtered back projection are shown.
+The resulting image looks quite crips and edges are clearly visible.
+The Shepp Logan phantom~\cite{shepplogan} was used as source image.  
+}
+\label{fig:backprojection}
+\end{figure}
+%
 
 \section{COR Determination}
+The problem of incorrect assumption of the location of the COR can occur in CT setups with misaligned source and detector.
+Especially in industrial CT with resolutions on the micrometer scale, calibrating a CT system correctly can be very challenging.
+In parallel beam geometry, a misaligned COR, i.e. the COR is projected onto $\tau$ on the $r$-axis, results in a shift of the sinogram along the $r$-axis by $\tau$.
+\begin{equation}
+R_\tau(r,\gamma) = R(r-\tau,\gamma)
+\end{equation}
+The actual problem arises when reconstructing the signal from $R_\tau$ as the algorithms, e.g. filtered back projection (FBP), assume the COR to be in the center.
+Thus, ray intersections do not lign up correctly and the reconstructed image suffers from streaking artifacts.
+\Cref{fig:shiftfbp} shows the effect of shift of COR in FBP.
+%
+\begin{figure}[!h]
+\centering
+\subfloat[$\tau=0$px]{\includegraphics[width=0.24\linewidth]{img/fbp_shift_0}}
+\hfil
+\subfloat[$\tau=2$px]{\includegraphics[width=0.24\linewidth]{img/fbp_shift_2}}
+\hfil
+\subfloat[$\tau=4$px]{\includegraphics[width=0.24\linewidth]{img/fbp_shift_4}}
+\hfil
+\subfloat[$\tau=8$px]{\includegraphics[width=0.24\linewidth]{img/fbp_shift_8}}
+\caption{
+FBP of the phantom from \cref{fig:sinogram_parallel} with different shifts ($\tau$) of COR.
+The resolution of the sinogram in $r$ direction was 256 pixels.
+With incresing shift, stronger vertical streaking artifacts are introduced into the reconstructed image.
+}
+\label{fig:shiftfbp}
+\end{figure}
+%
 

### 2017-07-19 seminar_cor/seminar_cor_haegele.tex
wrote center of mass section

@@
 \begin{equation}
+\label{eq:radonoffset}
 R_\tau(r,\gamma) = R(r-\tau,\gamma)
@@
 
+In order to correctly reconstruct the signal, the COR first need to be determined.
+Several approaches to this problem can be found in the literature, which will be introduced in the following subsections.
+The variety of approaches can subdivided in sinogram base methods, which analyze a sinogram in order to find the COR, and reconstruction based methods, which compare reconstruction results with different COR in an iterative fashion.
+
 \subsection{Sinogram Based Methods}
+In this subsection the approaches that are based on sinogram analysis are introduced.
+The approaches are presentred in chronical order.
 
 \subsubsection{High Density Feature}
+Crawford~et~al.~\cite{crawford1988} presented an approach for finding the COR in which they used two $180^\circ$ opposing projections of a distinct high density feature.
+When locating this feature in the two projections, the COR can be calculated as the sum of the detector coordinates of the feature divided by two.
+This becomes clear when thinking about the feature being offset from the COR by $\delta$ in projection $p_1$.
+The corresponding position of the feature in projection $p_2$ will then be on the opposite side of the COR, offset by $-\delta$ from the COR. \Cref{fig:highdenistyfeature} illustrates this concept.
+%
+\begin{figure}[!h]
+\centering
+\subfloat[phantom]{\includegraphics[width=0.32\linewidth]{img/highdensity_inv.png}}
+\hfil
+\subfloat[sinogram]{\includegraphics[width=0.65\linewidth]{img/highdensityradon_degrees_feature_centerline_horiz.png}}
+\caption{
+A phantom with a high density feature is shown on the left side.
+In corresponding sinogram on the right, the sinusodial trace of the feature can be observed.
+The dotted line indicates the COR in the sinogram.
+It can be seen that the feature locations in $180^\circ$ opposing projections have the same distance to the COR.  
+}
+\label{fig:highdenistyfeature}
+\end{figure}
+%
+
+This approach could also be expanded to use all projections in the sinogram.
+For each opposing pair of projections the COR could be calcualted and then the mean of all CORs will yield the final result.
+This approach can be automated if there esists a distinct global maximum in the projections of the sinogram, otherwise a location of the correct feature is not guaranteed.
 
 \subsubsection{Center of Mass}
+Azevedo~et~al.~\cite{azevedo90} proposed a more generally applicable algorithm, in which they replaced the high density feature from \cite{crawford1988} with the center of mass.
+The center of mass of the original signal is a point $(\overline{x},\overline{y}$) which's coordinates can be calculated as follows (calculation of $\overline{y}$ analogous to $\overline{x}$).
+%
+\begin{equation}
+\overline{x} = \frac{
+\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}x\;f(x,y)\;dx\;dy
+}
+{
+\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}f(x,y)\;dx\;dy
+}
+\end{equation}
+%
+As the center of mass of the original signal $f(x,y)$ is projected onto the projection center of mass in parallel beam geometry, it can be used as a reference feature.
+The projection center of mass is dependent on the projection angle and can be calculated as follows.
+%
+\begin{equation}
+\overline{r_\tau}(\gamma) = \frac{
+\int_{-\infty}^{\infty}r\;R_\tau(r,\gamma)\;dr
+}
+{
+\int_{-\infty}^{\infty}R_\tau(r,\gamma)\;dr
+}
+=
+\frac{
+\int_{-\infty}^{\infty}r\;R(r-\tau,\gamma)\;dr
+}
+{
+\int_{-\infty}^{\infty}R_(r-\tau,\gamma)\;dr
+}
+\end{equation}
+%
+where $R_\tau(r,\gamma)$ is the radon transform with COR offset by~$\tau$ (\cref{eq:radonoffset}).
+The previous equation can be simplified to the following.
+%
+\begin{equation}
+\label{eq:projection_centerofmass}
+\overline{r_\tau}(\gamma) = \tau + \overline{r}(\gamma) = \tau + \overline{x}\;cos(\gamma) + \overline{y}\;sin(\gamma)
+\end{equation}
+%
+Using linear regression (least squares) the parameters $\tau$, $\overline{x}$ and $\overline{y}$ can be approximated from the matrix notation of \cref{eq:projection_centerofmass}.
+%
+\begin{equation}
+\begin{pmatrix}
+1 & cos(\gamma_1) & sin(\gamma_1) \\
+\vdots & \vdots & \vdots \\
+1 & cos(\gamma_n) & sin(\gamma_n) 
+\end{pmatrix}
+\cdot
+\begin{pmatrix}
+\tau \\
+\overline{x} \\
+\overline{y}
+\end{pmatrix}
+=
+\begin{pmatrix}
+\overline{r_\tau}(\gamma_1) \\
+\vdots \\
+\overline{r_\tau}(\gamma_n)
+\end{pmatrix}
+\end{equation}
+%
+Even though $\overline{x}$ and $\overline{y}$ are not needed for a correct reconstruction, it is a neat bonus.
 

### 2017-07-21 seminar_cor/seminar_cor_haegele.tex
added todo just for me to remember
cross corr section

@@
 
-\section{COR Determination}
+\section{COR Determination}\label{sec:cordetermination}
 The problem of incorrect assumption of the location of the COR can occur in CT setups with misaligned source and detector.
@@
 
+Azevedo~et~al. showed that their algorithm is quite robust against Gaussian noise and isolated singularities~\cite{azevedo90}.
+Unfortuanately their algorithm is not applicable to fan beam geometry in general the objects center of mass is no longer projected onto the projections center of mass as was the case for parallel beam.
+However for an object function $f(x,y) = \delta(x-\overline{x}, y-\overline{y})$, that is a dirac delta function translated by $(\overline{x},\overline{y})$, the center of mass is projected onto the projections center of mass.
+Therefore this method can be used for calibration (i.e. finding the COR beforehand) using an object that approximates a dirac delta like a thin wire or pin phantom.
+
 \subsubsection{Cross-Correlation}
+Yang et al.~\cite{Yang2011} proposed another sinogram based method for COR determination using image cross-correllation.
+Given a sinogram in a prallel beam geometry over $360^\circ$ (which can also be obtained by rebinning a fan-beam sinogram), the sinogram is divided in half so that both halfs cover a $180^\circ$ range.
+As the information in both sub sinograms is redundant in that way that they are mirror images of each other,
+flipping one of them will result in two sinograms over $180^\circ$ with identical information.
+%
+\begin{equation}
+R(r,\gamma)^{[0,\pi]} = R(-r,\gamma)^{[\pi,2\pi]}
+\end{equation}
+%
+A shift of COR, as explained in the beginning of \cref{sec:cordetermination}, results in a shift of the sinogram along the $r$-axis.
+Thus, the two sub sinograms are identical under this shift.
+%
+\begin{equation}
+R_\tau(r+\tau,\gamma)^{[0,\pi]} = R_\tau(-(r+\tau),\gamma)^{[\pi,2\pi]}
+\end{equation}
+%
+Using the 1D the cross correlation of the two sinograms along the $r$-axis, the shift can be determined as follows.
+%
+\begin{equation}
+2\tau = \underset{r}{arg\,max}\quad R_\tau(r,\gamma)^{[0,\pi]} \star R_\tau(-r,\gamma)^{[\pi,2\pi]}
+\end{equation}
+%
 

### 2017-07-21 seminar_cor/seminar_cor_haegele.tex
finished sinogram based

@@
 
-\subsubsection{Cone Beam}
+\subsubsection{Cone Beam}\label{sssec:conebeam}
 The previously introduced geometries are both 2-dimensional (2D) as they described projections within an x-y-plane onto a one dimensional detector.
@@
 %
-A shift of COR, as explained in the beginning of \cref{sec:cordetermination}, results in a shift of the sinogram along the $r$-axis.
+A shift of COR, as explained in the beginning of \cref{sec:cordetermination}, results in a shift of the sinogram along the $r$-axis (see~\cref{eq:radonoffset}).
 Thus, the two sub sinograms are identical under this shift.
@@
 %
-Using the 1D the cross correlation of the two sinograms along the $r$-axis, the shift can be determined as follows.
+Using the 1D cross correlation of the two sinograms along the $r$-axis, the shift can be determined as follows.
 %
@@
 %
+The authors~\cite{Yang2011} provided evidence that their algorithm is robust against Gaussian noise even for large variance values.
+They also stated that their algorithm can be used for 3D CT with cone beam geometry when using the central slice (see \cref{sssec:conebeam}).
+As locating the central slice is not trivial and depends on calibration as well, it was shown that the algorithm still performs well for sinograms that slightly deviate from that of the central slice.
 

### 2017-07-21 seminar_cor/seminar_cor_haegele.tex
image metrics done

@@
 
-\subsubsection{Filtered Back Projection}
+\subsubsection{Filtered Back Projection}\label{sssec:fbp}
 The back projection algorithm reconstructs the original image from a discrete radon transform (i.e. parallel beam sinogram).
@@
 \subsection{Reconstruction Based Methods}
+Reconstruction based methods, as opposed to sinogram based methods, use an iterative approach to finding the COR.
+The individual steps of an iterative algorithm are:
+\begin{enumerate}
+\item reconstruct image with COR $\tau_i$
+\item generate a score for the reconstructed image that indicates the quality of the reconstruction, and thus the correctness of $\tau_i$
+\item repeat for different $\tau_i$
+\item select $\tau_i$ with best score as correct COR
+\end{enumerate}
+This process is of course very time consuming due to the repeating reconstruction step, but is not dependent on specific CT geometries.
+The crucial part of this algorithm is the generation of a score, which requires an image metric $Q(\tilde{f})$ to assess the quality of the reconstructed image $\tilde{f}$.
+In the following two different metrics are introduced.
 
 \subsubsection{Integral of Negativity}
+As proposed by Donath et~al.~\cite{donath2006} the integral of negativity can be leveraged to build an image metric $Q_{IN}$.
+%
+\begin{align}
+Q_{IN}(\tilde{f}) &= -\frac{1}{m_0}\;\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}\varepsilon(-\tilde{f}(x,y))\,\tilde{f}(x,y)\;dx\,dy.
+\\
+m_0 &= \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}f(x,y)\;dx\,dy = \int_{-\infty}^{\infty} R(r,\gamma_0)\;dr
+\\
+\varepsilon(a) &= \bigg\{
+\begin{matrix}
+1: & a \geq 0 \\
+0: & else
+\end{matrix}\quad \text{(Heaviside function)} 
+\end{align}
+%
+$Q_{IN}$ is calculated as the ratio of the integral of negativity of the reconstructed image to the zeroth order moment $m_0$ (total mass) of the original signal. 
+Fortunately the original image does not need to be known in order to calculate $m_0$, it is the integral of along the $r$-axis of any projection of the sinogram.
+
+Recalling the filtering step in FBP where negative projection values are introduced to cancel out positive contribution (see~\cref{sssec:fbp}), it makes sense to analyse the negativity of the reconstructed image.
+In a perfect reconstruction the negativity would be zero as the original image had no negative values, so minimizing this metric yields the correct COR.
+
+Brunetti and De~Carlo~\cite{brunetti2004} used a similar metric in their work based on the number of negative values in the reconstructed image.
 
 \subsubsection{Histogram Entropy}
+As pointed out by Donath et~al.~\cite{donath2006}, the metric $Q_{IN}$ can only be used for non-negative object functions.
+Therefore they provided another metric based on the histogram entropy that is also applicable when negative attenuation coefficients are present (which can occur when samples are recorded in a surrounding medium e.g.
+water \cite{donath2006}).
+They defined a metric $Q_H$ as follows.
+%
+\begin{align}
+Q_H(\tilde{f}) &= \frac{H}{H_{max}} 
+\\
+H &= -\int_{-\infty}^{\infty}p(g)\,log_2(h\,p(g))\;dg
+\end{align}
+%
+Here the metric is build by normalizing the entropy $H$ by the maximum entropy $H_{max}$.
+$H$ is a continous version of the discrete entropy for gray-level images, where $p(g)$ is an estimate probability density function for continous gray levels $g$ and $h$ describes the density resolution.
+
+The reasoning for the correctness of this metric is based on the assumption that the reconstructed image consists of homgenious regions, which is at least true for the outer regions that are zero.
+In a reconstruction with wrong COR, additional structures are introduced which smear out the histogram.
 

### 2017-07-22 seminar_cor/seminar_cor_haegele.tex
done with conclusion

@@
 
-\subsection{Image Acquisition in CT}
+\subsection{Image Acquisition in CT}\label{ssec:imageacquisition}
 This subsection will briefly explain how images are typically aquired in CT.
@@
 
-\subsubsection{Parallel Beam}
+\subsubsection{Parallel Beam}\label{sssec:parallelbeam}
 The simplest geometry in CT (from a mathematical point of view) is the parallel beam geometry.
@@
 
-\subsubsection{Fan Beam}
+\subsubsection{Fan Beam}\label{sssec:fanbeam}
 A more practical geometry is the fan beam, in which a source emits a flat beam that extends to the sides (like a fan blade). 
@@
 
-\subsection{Reconstruction}
+\subsection{Reconstruction}\label{ssec:reconstruction}
 This subsection will explain the process of reconstructing an image from the projections obtained in CT.
@@
 
-\subsubsection{High Density Feature}
+\subsubsection{High Density Feature}\label{sssec:highdensityfeature}
 Crawford~et~al.~\cite{crawford1988} presented an approach for finding the COR in which they used two $180^\circ$ opposing projections of a distinct high density feature.
@@
 
-\subsubsection{Center of Mass}
+\subsubsection{Center of Mass}\label{sssec:centerofmass}
 Azevedo~et~al.~\cite{azevedo90} proposed a more generally applicable algorithm, in which they replaced the high density feature from \cite{crawford1988} with the center of mass.
@@
 
-\subsubsection{Cross-Correlation}
+\subsubsection{Cross-Correlation}\label{sssec:crosscorrelation}
 Yang et al.~\cite{Yang2011} proposed another sinogram based method for COR determination using image cross-correllation.
@@
 
-\subsection{Reconstruction Based Methods}
+\subsection{Reconstruction Based Methods}\label{ssec:reconstructionbasedmethods}
 Reconstruction based methods, as opposed to sinogram based methods, use an iterative approach to finding the COR.
@@
 
-\subsubsection{Integral of Negativity}
+\subsubsection{Integral of Negativity}\label{sssec:qin}
 As proposed by Donath et~al.~\cite{donath2006} the integral of negativity can be leveraged to build an image metric $Q_{IN}$.
@@
 
-\subsubsection{Histogram Entropy}
+\subsubsection{Histogram Entropy}\label{sssec:qh}
 As pointed out by Donath et~al.~\cite{donath2006}, the metric $Q_{IN}$ can only be used for non-negative object functions.
@@
 \section{Conclusion}
-The conclusion goes here.
+In this paper about the determination of the COR in CT, a brief introduction was given about image acquisition in CT in the beginning~(\cref{ssec:imageacquisition}).
+Several CT geometries were introduced; 
+the parallel beam geometry~(\cref{sssec:parallelbeam}) which is mathematically described by the Radon transform and which most of the COR determination algorithms assume;
+the fan beam geometry (\cref{sssec:fanbeam}) which is more common and can be transormed to a parallel beam geometry;
+and the cone beam geometry (\cref{sssec:conebeam}) which, in contrast to the former, is a 3D geometry.
+In the last part of the introduction to CT imaging, the reconstruction was discussed (\cref{ssec:reconstruction}) in which the Fourier slice theorem was explained as well as the filtered back projection (\cref{sssec:fbp}) as a representative reconstruction algorithm.
+
+In the following section the key topic of this paper was adressed.
+Firstly the problem of inaccurate or wrong COR assumption for reconstruction was explained (\cref{sec:cordetermination}) which is the image degradation due to artifacts.
+Afterwards the different methdods for COR determination found in the literature were presented which can be classified into sinogram based methods and reconstruction based methods.
+
+The sinogram based methods comprise the high density feature tracing method (\cref{sssec:highdensityfeature}), in which a pin phantom can be used to create a distinct sinusodial trace in the sinogram from which the actual COR can be deduced;
+the center of mass method (\cref{sssec:centerofmass}) which is also applicable without the use of a special phantom as it uses the objects center of mass as a tracing point;
+and the cross-correlation method (\cref{sssec:crosscorrelation}) which leverages the redundant information in the upper and lower half of a sinogram over $360^\circ$.
+
+In the section of reconstruction based methods a basic iterative algorithm was introduced (\cref{ssec:reconstructionbasedmethods}) which uses uses an image metric to judge the quality of a reconstructed image.
+Following that, two image metrics were presented that could be used in the algorithm, which are the integral of negativity metric (\cref{sssec:qin}) and the histogram entropy metric (\cref{sssec:qh}).
 
@@
 
-\appendices
-\section{}
+%\appendices
+%\section{}
 

### 2017-07-22 seminar_cor/seminar_cor_haegele.tex
altered images, done abstract

@@
 \begin{abstract}
-The abstract goes here.
+This paper will discuss the aspect of the determination of center of rotation in computed tomography.
+An introduction to computed tomography imaging will be given in the beginning as a prerequisite to the actual topic. 
+This will cover different geometries (parallel beam, fan beam and cone beam) which are present in computed tomography and used to model the projection space of a CT-scan as well as basic knowledge about tomographic reconstruction, including the Fourier slice theorem and filtered back projection.
+The problem of wrong center of rotation assumption in tomographic reconstruction will be explained which has the effect of image degradation.
+Different solutions to the determination of the correct location of the center of rotation found in the literature will be presented.
+It is found that the approaches can be split into sinogram and reconstruction based methods, where sinogram based methods analyze the sinogram that is the product of a CT-scan, while reconstruction based methods analyze reconstructed images of different centers of rotation to find the correct one.
+Finally a conclusion summarizing the paper will be drawn.  
 \end{abstract}
@@
 \centering
-\includegraphics[width=\linewidth]{img/phantomandradon.png}
+\includegraphics[width=\linewidth]{img/phantomandradondarker.png}
 \caption{
@@
 \centering
-\subfloat[phantom]{\includegraphics[width=0.32\linewidth]{img/highdensity_inv.png}}
+\subfloat[phantom]{\includegraphics[width=0.32\linewidth]{img/highdensity_inv_darker.png}}
 \hfil

### 2017-07-26 seminar_cor/seminar_cor_haegele.tex
added presentation

@@
 % The paper headers
-\markboth{Seminar on Image Processing - Algorithms and Architectures - Univerity of Stuttgart - IPVS/PAS - SS17}%
+\markboth{Seminar on Image Processing - Algorithms and Architectures - University of Stuttgart - IPVS/PAS - SS17}%
 {H{\"a}gele: Determination of Center of Rotation in Computed Tomography}
@@
 An introduction to computed tomography imaging will be given in the beginning as a prerequisite to the actual topic. 
-This will cover different geometries (parallel beam, fan beam and cone beam) which are present in computed tomography and used to model the projection space of a CT-scan as well as basic knowledge about tomographic reconstruction, including the Fourier slice theorem and filtered back projection.
-The problem of wrong center of rotation assumption in tomographic reconstruction will be explained which has the effect of image degradation.
+This will cover different geometries (parallel beam, fan beam and cone beam), which are present in computed tomography and used to model the projection space of a CT-scan, as well as basic knowledge about tomographic reconstruction, including the Fourier slice theorem and filtered back projection.
+The problem of wrong center of rotation assumption in tomographic reconstruction will be explained, which has the effect of image degradation.
 Different solutions to the determination of the correct location of the center of rotation found in the literature will be presented.
@@
 The \xrays are emmited by a source on one side of the scanned object, intersect with the object, and hit a detector on the other side.
-The resulting meassurement of the detector is a projection of the object's internals.
+The resulting measurement of the detector is a projection of the object's internals.
 Rotating the object and repeating the process results in a set of projections, which can be used to reconstruct an image of the internal structure of the scanned object. 
@@
 \section{Prerequisites}
-In order to explain the problem of inaccurate COR assumption in tomographic reconstruction, some prerquisite knowledge about CT will be briefly explained in this section.
+In order to explain the problem of inaccurate COR assumption in tomographic reconstruction, some prerequisite knowledge about CT will be briefly explained in this section.
 
 \subsection{Image Acquisition in CT}\label{ssec:imageacquisition}
-This subsection will briefly explain how images are typically aquired in CT.
+This subsection will briefly explain how images are typically acquired in CT.
 
@@
 In parallel beam the source moving perpendicular to the ray direction for sampling a 1-dimensional image (graph). 
-In fan beam, rays are not perpendicular but instead emmited in fan blade form. 
+In fan beam, rays are not perpendicular but instead emitted in fan blade form. 
 $\gamma$ is the rotation angle at which the signal is obtained. 
@@
 
-Unfortuantely, parallel beam geometry is impractical as a pencil beam \xray source (source emmiting only one ray in a specific direction) would have to be moved parallel to the detector to collect all samples for a single rotation angle, which is very time consuming \cite{Buzug2008_chap7}.
+Unfortuantely, parallel beam geometry is impractical as a pencil beam \xray source (source emitting only one ray in a specific direction) would have to be moved parallel to the detector to collect all samples for a single rotation angle, which is very time consuming \cite{Buzug2008_chap7}.
 However, parallel beam geometry occurs at synchrotron facilities \cite{something}.
@@
 The geometry can be defined by the COR it rotates around, the offset of the source to the COR and its aperture angle.
-In clinical practise curved detector rays are used, as shown in \cref{subfig:fanbeam}, to get constant aperture angle steps for each detector cell.
+In clinical practice curved detector rays are used, as shown in \cref{subfig:fanbeam}, to get constant aperture angle steps for each detector cell.
 %
@@
 %
-The problem of applying this algorithm in practise is, that due to the limited angular resolution (i.e.
+The problem of applying this algorithm in practice is, that due to the limited angular resolution (i.e.
 discretization) of the Radon transform the frequency information density in the Fourier space decreases for higher frequencies \cite{Buzug2008_chap5}.
-In the process of cartesian regridding of the polar coordinates a lot of data for the high frequencies would have to be made up in order to perfrom an inverse fast fourier transform.
-Therefore a different algorithm, the filtered back projection, is used in practise.
+In the process of cartesian regridding of the polar coordinates a lot of data for the high frequencies would have to be made up in order to perform an inverse fast Fourier transform.
+Therefore a different algorithm, the filtered back projection, is used in practice.
 
@@
 At first glance this may seem as the correct inversion of the projections but is not.
-This becomes clear when thinking about the original image being a dirac delta (image with all entreis zero except for the central pixel).
-The projections of this image will be 1D dirac deltas, but using the back projection algorithm will not
-restore the 2D dirac but instead smear the 1D diracs also into areas which are supposed to be zero.
+This becomes clear when thinking about the original image being a Dirac delta (image with all entries zero except for the central pixel).
+The projections of this image will be 1D Dirac deltas, but using the back projection algorithm will not
+restore the 2D Dirac but instead smear the 1D Diracs also into areas which are supposed to be zero.
 In fact the resulting image looks quite blurry which is due to the back projection equaling to the original image convolved with the point spread function of the imaging system (i.e.
@@
 Figure \cref{fig:backprojection} shows the results of unfiltered and filtered back projection.
-As amplifying high frequencies also aplifies noise, there are different filters that can be used in practise \cite{farquhar1998}.
+As amplifying high frequencies also amplifies noise, there are different filters that can be used in practice \cite{farquhar1998}.
 %
@@
 In this subsection the approaches that are based on sinogram analysis are introduced.
-The approaches are presentred in chronical order.
+The approaches are presented in chronological order.
 
@@
 This approach could also be expanded to use all projections in the sinogram.
-For each opposing pair of projections the COR could be calcualted and then the mean of all CORs will yield the final result.
-This approach can be automated if there esists a distinct global maximum in the projections of the sinogram, otherwise a location of the correct feature is not guaranteed.
+For each opposing pair of projections the COR could be calculated and then the mean of all CORs will yield the final result.
+This approach can be automated if there exists a distinct global maximum in the projections of the sinogram, otherwise a location of the correct feature is not guaranteed.
 
@@
 Azevedo~et~al. showed that their algorithm is quite robust against Gaussian noise and isolated singularities~\cite{azevedo90}.
-Unfortuanately their algorithm is not applicable to fan beam geometry in general the objects center of mass is no longer projected onto the projections center of mass as was the case for parallel beam.
-However for an object function $f(x,y) = \delta(x-\overline{x}, y-\overline{y})$, that is a dirac delta function translated by $(\overline{x},\overline{y})$, the center of mass is projected onto the projections center of mass.
-Therefore this method can be used for calibration (i.e. finding the COR beforehand) using an object that approximates a dirac delta like a thin wire or pin phantom.
+Unfortunately their algorithm is not applicable to fan beam geometry in general the objects center of mass is no longer projected onto the projections center of mass as was the case for parallel beam.
+However for an object function $f(x,y) = \delta(x-\overline{x}, y-\overline{y})$, that is a Dirac delta function translated by $(\overline{x},\overline{y})$, the center of mass is projected onto the projections center of mass.
+Therefore this method can be used for calibration (i.e. finding the COR beforehand) using an object that approximates a Dirac delta like a thin wire or pin phantom.
 
 \subsubsection{Cross-Correlation}\label{sssec:crosscorrelation}
-Yang et al.~\cite{Yang2011} proposed another sinogram based method for COR determination using image cross-correllation.
-Given a sinogram in a prallel beam geometry over $360^\circ$ (which can also be obtained by rebinning a fan-beam sinogram), the sinogram is divided in half so that both halfs cover a $180^\circ$ range.
+Yang et al.~\cite{Yang2011} proposed another sinogram based method for COR determination using image cross-correlation.
+Given a sinogram in a parallel beam geometry over $360^\circ$ (which can also be obtained by rebinning a fan-beam sinogram), the sinogram is divided in half so that both halfs cover a $180^\circ$ range.
 As the information in both sub sinograms is redundant in that way that they are mirror images of each other,
@@
 
-Recalling the filtering step in FBP where negative projection values are introduced to cancel out positive contribution (see~\cref{sssec:fbp}), it makes sense to analyse the negativity of the reconstructed image.
+Recalling the filtering step in FBP where negative projection values are introduced to cancel out positive contribution (see~\cref{sssec:fbp}), it makes sense to analyze the negativity of the reconstructed image.
 In a perfect reconstruction the negativity would be zero as the original image had no negative values, so minimizing this metric yields the correct COR.
@@
 Here the metric is build by normalizing the entropy $H$ by the maximum entropy $H_{max}$.
-$H$ is a continous version of the discrete entropy for gray-level images, where $p(g)$ is an estimate probability density function for continous gray levels $g$ and $h$ describes the density resolution.
-
-The reasoning for the correctness of this metric is based on the assumption that the reconstructed image consists of homgenious regions, which is at least true for the outer regions that are zero.
+$H$ is a continuous version of the discrete entropy for gray-level images, where $p(g)$ is an estimate probability density function for continuous gray levels $g$ and $h$ describes the density resolution.
+
+The reasoning for the correctness of this metric is based on the assumption that the reconstructed image consists of homogenious regions, which is at least true for the outer regions that are zero.
 In a reconstruction with wrong COR, additional structures are introduced which smear out the histogram.
@@
 the parallel beam geometry~(\cref{sssec:parallelbeam}) which is mathematically described by the Radon transform and which most of the COR determination algorithms assume;
-the fan beam geometry (\cref{sssec:fanbeam}) which is more common and can be transormed to a parallel beam geometry;
+the fan beam geometry (\cref{sssec:fanbeam}) which is more common and can be transformed to a parallel beam geometry;
 and the cone beam geometry (\cref{sssec:conebeam}) which, in contrast to the former, is a 3D geometry.
@@
 
-In the following section the key topic of this paper was adressed.
+In the following section the key topic of this paper was addressed.
 Firstly the problem of inaccurate or wrong COR assumption for reconstruction was explained (\cref{sec:cordetermination}) which is the image degradation due to artifacts.
-Afterwards the different methdods for COR determination found in the literature were presented which can be classified into sinogram based methods and reconstruction based methods.
+Afterwards the different methods for COR determination found in the literature were presented which can be classified into sinogram based methods and reconstruction based methods.
 

### 2018-01-22 conclusion.tex
initial upload this is a short version of my bachelor thesis. had to do it in conjunction with the 'ringvorlesung' in master.

\section{Conclusion and Future Work}
A correlation based approach to the 6DOF image registration problem consisting of 3D translation and rotation was introduced.
The algorithm is capable of determining arbitrary rotations of an object between two voxel data sets.
It leverages the Fourier transform to decouple the rotational from translational information.
The rotation is calculated using sphere surfaces with different rotation axes of the data sets Fourier magnitude spectra, where the sphere surface with highest cross-correlation coefficient indicates the correct axis and the location of this coefficient indicates the angle of rotation around that axis.

The accuracy of the algorithm is dependent on the resolution of the data sets but could be improved when using a sub pixel accurate cross-correlation as proposed by Campbell and Wu~\cite{campbell08gradientcrosscorr}.

For high resolution data sets, a single sphere surface may contain insufficient data due to the small frequency range of a single radius.
More information could be obtained when integrating over a radius range in sphere surface extraction. 
B\"ulow and Birk~\cite{bulow13dof6registration} provided evidence for accuracy improvement in their approach when using multiple layers of several radii.

### 2018-01-22 introduction.tex
initial upload this is a short version of my bachelor thesis. had to do it in conjunction with the 'ringvorlesung' in master.

\section{Introduction}

In image registration, the process of transforming different sets of data to the same coordinate system \cite{wiki_imageregistration}, it is often required to determine the translation and rotation of objects between several images. 
This process is necessary to be able to compare or integrate data obtained under different conditions.
This paper presents an approach to automatically determine the translation and rotation of an object in two three-dimensional (3D) voxel data sets. 
As a real world use case, the recent work of Guhathakurta et~al.\ \cite{Guhathakurta} can be used. 
They presented a technique to reduce reconstruction and beam hardening artifacts from computed tomography (CT) by merging two CT scans where the scanned object was rotated for the second scan. 
In order to combine the voxel data sets it was important to know the exact orientation of the object after rotation, which required a supervised step for aligning the data sets which could be automated using the presented approach.

The solution to this image registration problem was adapted from the two-dimensional (2D) correlation-based approach towards calculation of rotation and translation of moving cells presented by Wilson and Theriot \cite{WilsonT06}.
The presented approach uses Fourier spectra for the decoupling of rotation from translation and cross-correlation techniques for shift determination.
The rotation determination process compares cross-correlations of spherical transformations of Fourier spectra to find the best matching 3D rotation.

### 2018-01-22 methodology.tex
initial upload this is a short version of my bachelor thesis. had to do it in conjunction with the 'ringvorlesung' in master.

\section{Methodology}
The starting point to the problem are two voxel data sets in regular grids, which were obtained using a CT scanner.
In both scans the same object was used, but scanned with a different orientation and position the second time.
The basic idea for determining the rotation and translation of the object between those two data sets, is to first determine the rotation, undo the rotation in the second data set, and then determine the translation.
This approach is based on the work of Wilson and Theriot~\cite{WilsonT06} who proposed a solution for the analog 2D problem with 3 degrees of Freedom (3DOF).
They used the well known method of decoupling the rotation from the translation problem using the Fourier transform, leveraging its 
rotation theorem, which states that the Fourier transform of a rotated 2D signal is equal to the rotated Fourier transform of the original signal.

$\mathcal{F}(f \circ R) = \mathcal{F}(f) \circ R \;\quad$
with $\mathcal{F}$ as the Fourier transform and $\circ$ as the operator that applies a coordinate transformation, in this case the rotation $R$.

The magnitude spectrum of the Fourier transform yields a translation independent representation of the original data set, leveraging the transform's translation property.
In 2D the rotation can be solved using a cross-correlation on the polar transformed magnitude spectra, as the angle determination turns into 1D shift determination problem in polar coordinates.
Similarly the translation can be determined using cross-correlation on the first and the back rotated second dataset.

In 3D all these steps can be used equivalently except for the crucial part of rotation determination from the magnitude spectra, as there is no 3D equivalent for the polar transform and the DOF increase from one in the 2D case to 3 in the 3D case.
Therefore the main contribution of this paper is the following algorithm for 3D rotation determination.
Given that under a certain rotation both magnitude spectra are identical, this means that the cross-correlation between the two spectra with correct rotation contains the global maximum correlation coefficient.
The problem's solution thus can be expressed as follows where $\star$ denotes the cross-correlation operator, $|\mathcal{F}(f)|$ is the Fourier magnitude spectrum of $f$ and $R_{\alpha,\beta,\gamma}$ is the rotation composed of the 3 Euler angles $\alpha,\beta,\gamma$.
\begin{equation}
\argmax_{\alpha,\beta,\gamma}\;[\;max\;(|\mathcal{F}(f_1)| \star (|\mathcal{F}(f_2)| \circ R_{\alpha,\beta,\gamma})\;)\;]
\end{equation}
As this expression is highly expensive to compute, the magnitude spectra will undergo a spherical coordinate transformation. 
When the rotation axis of this sphere transform can be arbitrarily chosen, the second magnitude spectrum's coordinate rotation ($R_{\alpha,\beta,\gamma}$) can be replaced by the sphere transform.
\begin{equation}\label{eq:rotatedspheretransform}
\Phi_{\alpha,\beta,\gamma}(r,\theta,\varphi) := f(x,y,z) \quad|\quad \text{with} 
\begin{pmatrix}
x \\ y \\ z
\end{pmatrix}
= R_{\alpha,\beta,\gamma} \; \cdot
\begin{pmatrix}
r \sin \theta \cos \varphi \\ 
r \sin \theta \sin \varphi \\ 
r \cos \theta
\end{pmatrix}
\end{equation}
Note that for choosing the rotation axis of the sphere transform, only two of the 3 Euler angles are needed as one of them only rotates the rotation axis around itself while translating the angle $\varphi$ to a different starting position. 
This redundancy is removed by choosing $\gamma=0$.
Now the correct rotation axis is found when the sphere transforms of both spectra are identical, which is the case when the 1-dimensional cross-correlation along the $\varphi$ axis contains the maximum correlation coefficient.
The solution for the first 2 Euler angles can be expressed as follows where $m = |\mathcal{F}(f)|$, thus $\Phi_{\alpha,\beta,\gamma}(m)$ denotes the sphere transform of $f$'s Fourier magnitude.
\begin{equation}
\argmax_{\alpha,\beta}\;[\;max\;(\Phi_{0,0,0}(m_1) \star \Phi_{\alpha,\beta,0}(m_2)\;)\;]
\end{equation}
While finding the maximum correlation coefficient of the cross-correlation, the third Euler angle can be directly determined from the maximum's position in direction of $\varphi$.
To actually benefit from the representation of magnitudes in spherical coordinates in terms of computation, the sphere transform is reduced to 2D using a fixed radius. 
The transform then turns into an extractor of a sphere surface, which still contains enough information if the radius is chosen to cover a low frequency of the Fourier magnitudes (high frequencies mostly correspond to noise and thus contain no usable information).

In practice solving for $\alpha$ and $\beta$ is still computationally intensive. 
Therefore a coarse to fine strategy can be leveraged to incrementally refine the solution for these two angles by trying a small number of different axis orientations at first and then refining the orientation in the area of highest maximum correlation coefficients.

### 2018-01-22 relatedwork.tex
initial upload this is a short version of my bachelor thesis. had to do it in conjunction with the 'ringvorlesung' in master.

\section{Related Work}

In the literature, several approaches to image registration problems have been proposed \cite{Brown92imageregistration}.
The particular problem with 6DOF for 3D rotation and 3D translation has been solved for example using feature detection \cite{wang1996_CT_markers}, a knowledge based technique where the features to be detected have to be known beforehand.
Other approaches leverage correlation metrics to determine the optimal solution.
Censi and Carpin~\cite{censi09houghradon} used projections of the data sets into the Hough/Radon domain and exploited its properties to solve for the rotation and translation via a series of cross-correlations.
Similar to the presented approach, the works of B\"ulow and Birk \cite{bulow13dof6registration} as well as Lucchese et~al.\ \cite{lucchese02rangedatareg} use the Fourrier transform to decouple rotation from translation.
In contrast to the presented approach Lucchese et~al.\ solve for an estimate rotation axis using a radial projection of the difference between the Fourrier magnitudes, whereas the presented approach cross-correlates a series of spherically transformed magnitude spectra to find the rotation with maximum correlation.
B\"ulow and Birk also use cross-correlation on the two spherically transformed magnitudes, but solve for the 3 rotation directly from this representation.
The downside of this technique is, that not all arbitrary rotations can be determined but are limited due to the nature of a spherical coordinate transformation.
The presented approach does not share this downside as it uses spherical transformations with different axes.

### 2018-05-08 seminar_vof/functionplot.m
new seminar stuff

@@ -1,0 +1,9 @@
+clear;
+clc;
+
+fn = @(x) sqrt(1+x^2);
+x = linspace(-10,10,100);
+y = arrayfun(fn,x);
+
+plot(x,y)
+

### 2018-05-08 seminar_vof/functionplot.m
more things

@@ -1,9 +1,23 @@
 clear;
 clc;
 
-fn = @(x) sqrt(1+x^2);
-x = linspace(-10,10,100);
-y = arrayfun(fn,x);
+chrb = @(x,eps) 2*(eps^2)*sqrt(1+(x^2)/(eps^2))-2*(eps^2)+1;
+fnc1 = @(x) chrb(x,1);
+fnp = @(x) log(1+x^2);
+fna = @(x) 2*abs(x)-1;
+fnq = @(x) x*x+1;
+x = linspace(-4,4,100);
+y1 = arrayfun(fnc1,x);
+y2 = arrayfun(fna,x);
+y3 = arrayfun(fnq,x);
 
-plot(x,y)
-
+hold on;
+p1 = plot(x,y1);
+p3 = plot(x,y2,'--');
+p4 = plot(x,y3,'--');
+axis([-inf inf 0 5]);
+axis equal
+set(p1, 'Linewidth', 2.5);
+set(p2, 'Linewidth', 2.5);
+set(p3, 'Linewidth', 1.3);
+set(p4, 'Linewidth', 1.3);

### 2018-05-10 seminar_vof/functionplot.m
more things

@@ -1,23 +1,51 @@
 clear;
 clc;
 
-chrb = @(x,eps) 2*(eps^2)*sqrt(1+(x^2)/(eps^2))-2*(eps^2)+1;
-fnc1 = @(x) chrb(x,1);
-fnp = @(x) log(1+x^2);
-fna = @(x) 2*abs(x)-1;
-fnq = @(x) x*x+1;
-x = linspace(-4,4,100);
-y1 = arrayfun(fnc1,x);
-y2 = arrayfun(fna,x);
-y3 = arrayfun(fnq,x);
+%chrb = @(x,eps) 2*(eps^2)*sqrt(1+(x^2)/(eps^2))-2*(eps^2)+1;
+%chrbp = @(x,eps) 1/sqrt(1+(x^2)/(eps^2));
+%prml = @(x,eps) eps^2*log(1+(x^2)/(eps^2));
+%prmlp = @(x,eps) 1/(1+(x^2)/(eps^2));
+%fnc1 = @(x) chrbp(x,1);
+%fnp = @(x) prml(x,1);
+%fna = @(x) log(abs(x)+1);
+%fnq = @(x) x*x;
+%x = linspace(-4,4,100);
+%y1 = arrayfun(fnc1,x);
+%y2 = arrayfun(fna,x);
+%y3 = arrayfun(fnq,x);
+%
+%hold on;
+%p1 = plot(x,y1);
+%%p3 = plot(x,y2,'--');
+%p4 = plot(x,y3,'--');
+%axis([-inf inf 0 5]);
+%axis equal
+%set(p1, 'Linewidth', 2.5);
+%%set(p2, 'Linewidth', 2.5);
+%%set(p3, 'Linewidth', 1.3);
+%set(p4, 'Linewidth', 1.3);
+
+
+T = 1;
+lam = 0.1;
+delta = @(benefit) T+benefit;
+c = @(benefit) 1/(1+e^(-delta(benefit)/lam)); 
+phi = @(x) lam*( log(1-c(x)) -c(x)*log(1/c(x) -1) ) -T*c(x);
+
+
+x = linspace(-2,2,100);
+y1 = arrayfun(phi,x);
 
 hold on;
 p1 = plot(x,y1);
-p3 = plot(x,y2,'--');
-p4 = plot(x,y3,'--');
-axis([-inf inf 0 5]);
+%p3 = plot(x,y2,'--');
+%p4 = plot(x,y3,'--');
[…]

### 2018-06-12 seminar_vof/paper/paper.tex
renamed template to paper, fixed name clash issue of /ifpdf

\documentclass[journal]{vgtc} 
\usepackage{hs-vis_ss10}

%% Please note that the use of figures other than the optional teaser
%% is not permitted on the first page of the journal version.  Figures
%% should begin on the second page and be in CMYK or Grey scale
%% format, otherwise, colour shifting may occur during the printing
%% process.  Papers submitted with figures other than the optional
%% teaser on the first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for
%% Type 1 typefaces, ``graphicx'' for inclusion of EPS figures. and
%% ``times'' for proper handling of the times font family.

\usepackage{mathptmx} 
\usepackage{graphicx}
\usepackage{times}

%% allow for this line if you want the electronic option to work
%% properly
\vgtcinsertpkg

%% author name
\author{Roy G. Biv}

%% paper title
\title{Visualisierung von Regenbogenfarben}

%% short title for header
\shorttitle{Visualisierung von Regenbogenfarben}

%% Abstract section.
\abstract{%
  Hier sollte eine kurze Zusammenfassung der Ausarbeitung stehen. Der
  Umfang sollte 150 bis 250 Worte betragen. Duis autem vel eum iriure
  dolor in hendrerit in vulputate velit esse molestie consequat, vel
  illum dolore eu feugiat nulla facilisis at vero eros et accumsan et
  iusto odio dignissim qui blandit praesent luptatum zzril delenit
  augue duis dolore te feugait nulla facilisi. Lorem ipsum dolor sit
  amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod
  tincidunt ut laoreet dolore magna aliquam erat volutpat.

  Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
  suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis
  autem vel eum iriure dolor in hendrerit in vulputate velit esse
  molestie consequat, vel illum dolore eu feugiat nulla facilisis at
  vero eros et accumsan et iusto odio dignissim qui blandit praesent
  luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
} % end of abstract

%% Uncomment below to include a (optional) teaser figure.
%\teaser{ \centering
%  \includegraphics[width=16cm]{images/CypressView}
%  \caption{In den Wolken: Vancouver von Cypress Mountain. Auf der
%    ersten Seite d"urfen keine Grafiken au"ser dieser optionalen
%    Aufmachgrafik (Teaser) abgebildet sein.}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title
%%   block.

%%   the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

Die Ausarbeitung darf eine L"ange von 6 Seiten nicht unterschreiten
und 8 Seiten nicht "uberschreiten. Dies beinhaltet Text, Abbildungen
und Referenzen. Bei Ausarbeitungen mit mehr als 8 Seiten werden nur
die ersten 8 Seiten zur Bewertung herangezogen. F"ur die Form der
Ausarbeitung steht auf der Seminarwebseite diese \LaTeX{}-Vorlage
bereit. Diese Vorlage ist bindend und darf nicht abge"andert
werden. Abgaben in anderer Form werden nicht akzeptiert.

\section{Instruktionen}

Bitte nachfolgende Abschnitte sorgf"altig durchlesen.

\subsection{Sprache}

Die Ausarbeitung kann entweder auf Deutsch oder Englisch geschrieben
werden.

\subsection{Abbildungen und Tabellen}

Alle Abbildungen (siehe Abb.\ \ref{fig:sample}) und Tabellen (Tabelle\
\ref{tab:vis_accept}) sollten zentriert sein
(\verb|\centering|). Abbildungen "uber beide Textspalten
(Abb. \ref{fig:multicolumn}) k"onnen mit
\verb|\begin{figure*}|\ldots\verb|\end{figure*}| eingef"ugt werden.

\subsection{Referenzen}

Literaturangaben wie beispielsweise Levoy \cite{levoy:1989:DSV} werden
mit Hilfe von BibTeX erzeugt. Dazu werden die Referenzen in die
Literaturliste (hier \emph{literatur.bib}) eingetragen und
entsprechend mit \verb|\cite| referenziert.

\subsection{\LaTeX-"Ubersetzung}

Die \LaTeX-Datei kann mit \emph{latex} oder \emph{pdflatex} "ubersetzt
werden. Dabei ist zu beachten, dass f"ur die "Ubersetzung mit
\emph{latex} die Grafiken in Postscript (eps) vorliegen, f"ur
\emph{pdflatex} entsprechend als jpg, png oder pdf.  Der Ablauf ist
dabei der folgende:
\begin{enumerate}
\item \verb|pdflatex <quelldatei.tex>|
\item \verb|bibtex <quelldatei>|
\item \verb|pdflatex <quelldatei.tex>| (evtl. mehrfach)
\end{enumerate}
Alternativ kann auch das mitgelieferte Makefile verwendet werden.

\section{Rendering}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam
voluptua~\cite{kitware2003,Max:1995:OMF}. At vero eos et accusam et
justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi~\cite{notes2002}. Lorem ipsum dolor sit amet, consectetuer
adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet
dolore magna aliquam erat volutpat.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

\section{Exposition}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum~\cite{ware:2004:IVP}. Stet clita kasd gubergren, no sea takimata
sanctus est Lorem ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat~\cite{kindlmann:1999:SAG}.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat~\cite{levoy:1989:DSV}. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua.

\begin{figure}[tb]
  \centering
  \includegraphics[width=1.5in]{sample}
  \caption{\label{fig:sample} Beispielillustration.}
\end{figure}

\begin{equation}
  \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

\begin{table}
  %% Table captions on top in journal version
  \caption{\label{tab:vis_accept} Vis Paper Acceptance Rate}
  \scriptsize
  \begin{center}
    \begin{tabular}{cccc}
      Year & Submitted & Accepted & Accepted (\%)\\
      \hline
      1994 &  91 & 41 & 45.1\\
      1995 & 102 & 41 & 40.2\\
      1996 & 101 & 43 & 42.6\\
      1997 & 117 & 44 & 37.6\\
      1998 & 118 & 50 & 42.4\\
      1999 & 129 & 47 & 36.4\\
      2000 & 151 & 52 & 34.4\\
      2001 & 152 & 51 & 33.6\\
      2002 & 172 & 58 & 33.7\\
      2003 & 192 & 63 & 32.8\\
      2004 & 167 & 46 & 27.6\\
      2005 & 268 & 88 & 32.8\\
      2006 & 228 & 63 & 27.6
    \end{tabular}
  \end{center}
\end{table}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

\subsection{Mezcal Head}

Duis autem~\cite{Lorensen:1987:MCA} vel eum iriure dolor in hendrerit
in vulputate velit esse molestie consequat, vel illum dolore eu
feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
ut laoreet dolore magna aliquam erat volutpat%
\footnote{Fu"snoten erscheinen an der Unterkante der Spalte. Sie
  sollten jedoch vermieden werden, da der Lesefluss gest"ort wird.}.

\subsubsection{Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo
consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
hendrerit in vulputate velit esse molestie consequat, vel illum dolore
eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
dignissim qui blandit praesent luptatum zzril delenit augue duis
dolore te feugait nulla facilisi.

\paragraph{Rejected Ejector Seat Reservation}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie

\section{Conclusion}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum.

\bibliographystyle{abbrv}
%% use following if all content of bibtex file should be shown
% \nocite{*}
\bibliography{literatur}
\end{document}

### 2018-07-05 seminar_vof/paper/paper.tex
writing variational flow recap

@@
 \usepackage{times}
+%\usepackage[utf8]{inputenc}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{mathtools}
+\usepackage{bigints}
+\usepackage{xfrac}
+\usepackage{bm}
+
+\newcommand{\lam}{\lambda}
+\newcommand{\trans}{\quad\bigg|\;}
+\newcommand{\Trans}{\quad\Bigg|\;}
+% matrix shortcut
+\newcommand{\mat}[1]{\begin{pmatrix} #1 \end{pmatrix}}
+% matrix shortcut with increased vertical spacing
+\newcommand{\Mat}[1]{\begingroup
+\renewcommand*{\arraystretch}{1.5}
+\mat{#1}
+\endgroup}
+% matrix format but without braces (secretly a matrix)
+\newcommand{\secretMat}[2][r]{\begin{matrix*}[#1] #2 \end{matrix*}}
+% matrix format but without braces (secretly a matrix) with increased vertical spacing
+\newcommand{\SecretMat}[2][r]{\begingroup
+\renewcommand*{\arraystretch}{1.7}
+\secretMat[#1]{#2}
+\endgroup}
+% shortcuts for df/dx stuff in derivative
+\newcommand{\der}{\partial}
+\newcommand{\deriv}[2]{\frac{\der #1}{\der #2}}
+% make star '*' become \cdot (normal multiplication sign)
+\mathcode`\*="8000
+{\catcode`\*\active\gdef*{\cdot}}
+\newcommand{\Real}{\mathbb{R}}
+\newcommand{\flow}{\vec{u}}
+\newcommand{\x}{\vec{x}}
+\newcommand{\y}{\vec{y}}
+\newcommand{\ev}{\vec{r}}
+\newcommand{\aaux}{\vec{a}}
+\newcommand{\baux}{\vec{b}}
+\newcommand{\T}{^\top}
+\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
 
@@
 
-\section{Rendering}
+\section{Introduction}
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam
-voluptua~\cite{kitware2003,Max:1995:OMF}. At vero eos et accusam et
-justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
+In computer vision, the field of optic flow deals with the estimation of a 2-dimensional flow field based on two consecutive frames of an image sequence.
+\\(mehr dazu warum optic flow geil ist)\\
+Different methods for flow estimation exist, such as ... ... and variational optic flow which is based on the calculus of variations.
+In the following the concept of variational flow estimation is recapitulated and the approach by \verb|\cite{daspaper}| with order adaptive regulalarization is presented.
 
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi~\cite{notes2002}. Lorem ipsum dolor sit amet, consectetuer
-adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet
-dolore magna aliquam erat volutpat.
+\section{Variational Optic Flow}
+In this section the general framework for flow estimation using variational methods is recapitulated.
+\\\\
+Given two subsequent images
+$I_0,I_1 ~:~ \Omega \subset \Real^2 \to \Real$ \\
+of an image sequence, we seek the displacement vector field 
+\\$\flow = (u~v)\T ~:~ \Omega \subset \Real^2 \to \Real^2$ 
+that maps 
+$I_0$ onto $I_1$ 
+such that\\
+$I_0(\x) = I_1(\x+\flow(\x))$.
+\\\\
+To estimate the flow function $\flow$ an energy functional is minimized. 
+Typically this functional consists of two terms, a data term $D(\flow)$ and a regularization term $R(\flow)$.
+\begin{align*}
+\flow' &= \argmin_{\flow} E(\flow)
+\\
+E(\flow) &= D(\flow) + \alpha R(\flow)
+\end{align*}
+In this framework, the data term is used to penalize flow functions which violate certain constancy assumptions on image features, such as the brightness constancy assumtion 
+$D_{brght}(\flow) = \int_\Omega (I_0(\x)-I_1(\x+\flow(\x))^2\;d\x$.
+As the data term is insufficient for solving this problem and would yield multiple ambiguous solutions as illustrated in \ref{fig:ambigdataterm}, the regularization term is used to overcome this inconvenience and penalize flow functions that are unlikely.
+This is done by setting up constraints for the flow, such as the first order smoothness assumption
+$R_{smooth}(\flow) = \int_\Omega || \nabla u ||_2^2 + || \nabla v ||_2^2\;d\x$.
+This assumption enforces constant flow due to penalizing non vanishing gradients, which amount to changes in the vector field.
 
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=1.5in]{images/ambiguousflow.png}
+  \caption{\label{fig:ambigdataterm} Beispielillustration.}
+\end{figure}
 

### 2018-07-10 seminar_vof/paper/paper.tex
some changes to template and mooore content

@@
 \usepackage{times}
-%\usepackage[utf8]{inputenc}
 \usepackage{amssymb}
@@
 \usepackage{bm}
+\usepackage{cleveref}
+\usepackage{csquotes}
+\usepackage{subcaption}
+\usepackage[english]{babel}
 
@@
 %% paper title
-\title{Visualisierung von Regenbogenfarben}
+\title{Order-Adaptive Regularisation for Variational Optic Flow:\\Global, Local and in Between}
 
 %% short title for header
-\shorttitle{Visualisierung von Regenbogenfarben}
+\shorttitle{Order-Adaptive Regularisation}
 
@@
 
-Die Ausarbeitung darf eine L"ange von 6 Seiten nicht unterschreiten
-und 8 Seiten nicht "uberschreiten. Dies beinhaltet Text, Abbildungen
-und Referenzen. Bei Ausarbeitungen mit mehr als 8 Seiten werden nur
-die ersten 8 Seiten zur Bewertung herangezogen. F"ur die Form der
-Ausarbeitung steht auf der Seminarwebseite diese \LaTeX{}-Vorlage
-bereit. Diese Vorlage ist bindend und darf nicht abge"andert
-werden. Abgaben in anderer Form werden nicht akzeptiert.
-
-\section{Instruktionen}
-
-Bitte nachfolgende Abschnitte sorgf"altig durchlesen.
-
-\subsection{Sprache}
-
-Die Ausarbeitung kann entweder auf Deutsch oder Englisch geschrieben
-werden.
-
-\subsection{Abbildungen und Tabellen}
-
-Alle Abbildungen (siehe Abb.\ \ref{fig:sample}) und Tabellen (Tabelle\
-\ref{tab:vis_accept}) sollten zentriert sein
-(\verb|\centering|). Abbildungen "uber beide Textspalten
-(Abb. \ref{fig:multicolumn}) k"onnen mit
-\verb|\begin{figure*}|\ldots\verb|\end{figure*}| eingef"ugt werden.
-
-\subsection{Referenzen}
-
-Literaturangaben wie beispielsweise Levoy \cite{levoy:1989:DSV} werden
-mit Hilfe von BibTeX erzeugt. Dazu werden die Referenzen in die
-Literaturliste (hier \emph{literatur.bib}) eingetragen und
-entsprechend mit \verb|\cite| referenziert.
-
-\subsection{\LaTeX-"Ubersetzung}
-
-Die \LaTeX-Datei kann mit \emph{latex} oder \emph{pdflatex} "ubersetzt
-werden. Dabei ist zu beachten, dass f"ur die "Ubersetzung mit
-\emph{latex} die Grafiken in Postscript (eps) vorliegen, f"ur
-\emph{pdflatex} entsprechend als jpg, png oder pdf.  Der Ablauf ist
-dabei der folgende:
-\begin{enumerate}
-\item \verb|pdflatex <quelldatei.tex>|
-\item \verb|bibtex <quelldatei>|
-\item \verb|pdflatex <quelldatei.tex>| (evtl. mehrfach)
-\end{enumerate}
-Alternativ kann auch das mitgelieferte Makefile verwendet werden.
-
-\section{Introduction}
-
 In computer vision, the field of optic flow deals with the estimation of a 2-dimensional flow field based on two consecutive frames of an image sequence.
@@
 
-\section{Variational Optic Flow}
+\section{Variational Optic Flow}\label{sec:variationalopticflow}
 In this section the general framework for flow estimation using variational methods is recapitulated.
@@
 Typically this functional consists of two terms, a data term $D(\flow)$ and a regularization term $R(\flow)$.
-\begin{align*}
-\flow' &= \argmin_{\flow} E(\flow)
+\begin{flalign*}
+\flow' &= \argmin_{\flow} E(\flow)&&
 \\
-E(\flow) &= D(\flow) + \alpha R(\flow)
-\end{align*}
-In this framework, the data term is used to penalize flow functions which violate certain constancy assumptions on image features, such as the brightness constancy assumtion 
-$D_{brght}(\flow) = \int_\Omega (I_0(\x)-I_1(\x+\flow(\x))^2\;d\x$.
-As the data term is insufficient for solving this problem and would yield multiple ambiguous solutions as illustrated in \ref{fig:ambigdataterm}, the regularization term is used to overcome this inconvenience and penalize flow functions that are unlikely.
-This is done by setting up constraints for the flow, such as the first order smoothness assumption
-$R_{smooth}(\flow) = \int_\Omega || \nabla u ||_2^2 + || \nabla v ||_2^2\;d\x$.
+E(\flow) &= D(\flow) + \alpha R(\flow)&&
+\end{flalign*}
+In this framework, the data term is used to penalize flow functions which violate certain constancy assumptions on image features, such as the brightness constancy assumtion. 
+\begin{flalign*}
+D_{brght}(\flow) = \int_\Omega (I_0(\x)-I_1(\x+\flow(\x))^2\;d\x&&
+\end{flalign*}
+As the data term is insufficient for solving this problem and would yield multiple ambiguous solutions as illustrated in \cref{fig:ambigdataterm}, the regularization term is used to overcome this inconvenience and penalize flow functions that are unlikely.
+This is done by setting up constraints for the flow, such as the first order smoothness assumption.
+\begin{flalign*}
+R_{smooth}(\flow) = \int_\Omega || \nabla u ||_2^2 + || \nabla v ||_2^2\;d\x&&
+\end{flalign*}
 This assumption enforces constant flow due to penalizing non vanishing gradients, which amount to changes in the vector field.
+\begin{figure}[htb]
+\centering
+\includegraphics[width=\linewidth]{images/ambiguousflow.png}
+\caption{Images show a rectangle that moved from bottom left to top right. The red arrows sketch a flow field explaining where the pixels moved.
+Even though both fields perfectly fulfill a data term with brightness and gradient constancy assumption, only the right field is reasonable and would minimize a first order smoothness regularization.}
+\label{fig:ambigdataterm}
+\end{figure}
 
-\begin{figure}[tb]
+In order to minimize the energy functional, an iterative update sheme is developed using the euler lagrange equation \cite{eulerlagrange}.
+
+\section{Motivation}
+As explained in \cref{sec:variationalopticflow} the regularization term can be used to constrain the \enquote{type} of the estimated flow function.
+In practice either first or second order smoothness assumptions are made in which a flow function minimizes the regularization term when its first or second derivative vanishes.
+Therefore, first order methods are suitable for motion that is constant, i.e. parallel to the image plane.
+Second order methods can be leveraged for estimating affine flow fields as the class of linear functions implies vanishing second derivative.
+\Cref{fig:flowfields} shows a constant and linear flow field for comparison.
+\begin{figure}[htb]
 \centering
-  \includegraphics[width=1.5in]{images/ambiguousflow.png}
-  \caption{\label{fig:ambigdataterm} Beispielillustration.}
+\begin{subfigure}{0.45\linewidth}
+\centering
+\includegraphics[width=\linewidth]{images/constantflow.eps}
+\end{subfigure}
+\begin{subfigure}{0.45\linewidth}
+\centering
+\includegraphics[width=\linewidth]{images/linearflow.eps}
+\end{subfigure}
+\caption{
+Graphs show vector fields of different behaviour, left is constant with $f(x,y) = (1\;\, 1)\T$, right is linear with $f(x,y) = (x \;\, y)\T$.
+Constant flow is typical for motion parallel to the image plane, linear flow for motion perpenicular to the image plane. 
+}
+\label{fig:flowfields}
 \end{figure}
+Depending on the situation in the image sequence, a constant flow field may be more reasonable than a linear one.
+Imagine for example a sequence captured from the view point of a car driver who waits at a crossing while other cars pass the crossing from left to right in front of him.
+The optic flow in this situation would be constant as all cars move in the same direction parallel to the image plane.
+Now imagine the car driver moving forwards as the crosisng is now free, the whole scene around him will appear to move towards him.
+The optic flow is now linear as the surroundings move perpendicular to the image plane.
+However, the flow behaviour is not exclusive, both constant and linear flow can be present in the same sequence, for example when another car heading in the same direction switches lanes.
 

### 2018-07-12 seminar_vof/paper/paper.tex
moar content aka lazy commitmsg

@@
 \newcommand{\argmin}{\mathop{\mathrm{argmin}}}
+\newcommand{\charbonnier}{\Psi_{\text{Ch}}}
+\newcommand{\peronamalik}{\Psi_{\text{PM}}}
 
@@
 \end{figure}
+
 Depending on the situation in the image sequence, a constant flow field may be more reasonable than a linear one.
@@
 However, the flow behaviour is not exclusive, both constant and linear flow can be present in the same sequence, for example when another car heading in the same direction switches lanes.
+To account for such situations, a variational model has to make more sophisticated assumptions than simple first or second order smoothness.
+
+Fortunately, the presented approach manages to combine first and second order regularisation and even adaptively choose optimal order for a given sequence.
+Moreover it provides different adaption shemes to enable a global, local, non-local and region based selection of regularisation.
+
+\section{Order Adaptive Variational Model}
+In this section the variational model by \verb|\citeauthor{daspaper}| introduced in \cite{daspaper} will be discussed in detail.
+
+Lets start with the data term which uses a combination of brightness and gradient constancy assumption.
+\begin{flalign*}
+D(\flow) &= \int_\Omega \charbonnier(p_{\text{brght}}(\flow)) + \gamma*\charbonnier(p_{\text{grad}}(\flow))\;d\x &&
+\\
+p_{\text{brght}}(\flow) &= (\;I_1(\x+\flow(\x))-I_0(\x)\;)^2 &&
+\\
+p_{\text{grad}}(\flow) &= ||\;\nabla I_1(\x+\flow(\x))-\nabla I_0(\x)\;||_2^2 &&
+\end{flalign*}
+In this setup, $\gamma$ is used to weight the terms against each other.
+As can be seen $p_{\text{brght}}$ and $p_{\text{grad}}$ are quadratic penalizers, which are sensitive to outliers.
+To remove the strong influence of outliers on the minimization, the subquadratic charbonnier penalizer $\charbonnier$ is used to 
+robustify the constancy assumptions.
+\begin{flalign*}
+\charbonnier(s^2) &= 2\epsilon^2\sqrt{1+s^2/\epsilon^2} &&
+\end{flalign*}
 

### 2018-07-12 seminar_vof/paper/paper.tex
moar

@@
 \newcommand{\argmin}{\mathop{\mathrm{argmin}}}
-\newcommand{\charbonnier}{\Psi_{\text{Ch}}}
-\newcommand{\peronamalik}{\Psi_{\text{PM}}}
+\newcommand{\charbonnier}{\psi_{\text{Ch}}}
+\newcommand{\peronamalik}{\psi_{\text{PM}}}
 
@@
 \section{Order Adaptive Variational Model}
-In this section the variational model by \verb|\citeauthor{daspaper}| introduced in \cite{daspaper} will be discussed in detail.
+In this section the variational model by Maurer, Stoll and Bruhn introduced in their work \cite{daspaper} will be discussed in detail.
 
-Lets start with the data term which uses a combination of brightness and gradient constancy assumption.
+\subsection{Dataterm}
+The employed data term in the variational model uses a combination of brightness and gradient constancy assumption.
 \begin{flalign*}
@@
 \end{flalign*}
-In this setup, $\gamma$ is used to weight the terms against each other.
-As can be seen $p_{\text{brght}}$ and $p_{\text{grad}}$ are quadratic penalizers, which are sensitive to outliers.
-To remove the strong influence of outliers on the minimization, the subquadratic charbonnier penalizer $\charbonnier$ is used to 
-robustify the constancy assumptions.
+In this setup, $p_{\text{brght}}$ is the penalizer for the brightness constancy assumption and $p_{\text{grad}}$ is the penalizer for the gradient constancy assumption which can be understood as a check for
+edges being mapped correctly.
+The parameter $\gamma$ is used to weight the terms against each other.
+As can be seen $p_{\text{brght}}$ and $p_{\text{grad}}$ are quadratic penalizers, and are thus sensitive to outliers.
+To remove the strong influence of outliers on the minimization, the subquadratic charbonnier penalizer \cite{charbonnier} $\charbonnier$ is used to 
+robustify both constancy assumptions.
 \begin{flalign*}
@@
 \end{flalign*}
+Note that the charbonnier penalizer is a function of a squared argument.
+When plotting the function with respect to $s$ instead of $s^2$ its behaviour becomes clearer, as shown in \cref{fig:penalizers}.
+For values close to zero, the function penalizes quadratically and approaches linear behaviour in the limit $s\to\pm\infty$, resulting in outliers being less fatal.
+
+\subsection{Regulalarisation}
+Hello
 

### 2018-07-12 seminar_vof/functionplot.m

@@ -1,50 +1,50 @@
 clear;
 clc;
 
-%chrb = @(x,eps) 2*(eps^2)*sqrt(1+(x^2)/(eps^2))-2*(eps^2)+1;
-%chrbp = @(x,eps) 1/sqrt(1+(x^2)/(eps^2));
-%prml = @(x,eps) eps^2*log(1+(x^2)/(eps^2));
-%prmlp = @(x,eps) 1/(1+(x^2)/(eps^2));
-%fnc1 = @(x) chrbp(x,1);
-%fnp = @(x) prml(x,1);
-%fna = @(x) log(abs(x)+1);
-%fnq = @(x) x*x;
-%x = linspace(-4,4,100);
-%y1 = arrayfun(fnc1,x);
-%y2 = arrayfun(fna,x);
-%y3 = arrayfun(fnq,x);
-%
-%hold on;
-%p1 = plot(x,y1);
-%%p3 = plot(x,y2,'--');
-%p4 = plot(x,y3,'--');
-%axis([-inf inf 0 5]);
-%axis equal
-%set(p1, 'Linewidth', 2.5);
-%%set(p2, 'Linewidth', 2.5);
-%%set(p3, 'Linewidth', 1.3);
-%set(p4, 'Linewidth', 1.3);
-
-
-T = 1;
-lam = 0.1;
-delta = @(benefit) T+benefit;
-c = @(benefit) 1/(1+e^(-delta(benefit)/lam)); 
-phi = @(x) lam*( log(1-c(x)) -c(x)*log(1/c(x) -1) ) -T*c(x);
-
-
-x = linspace(-2,2,100);
-y1 = arrayfun(phi,x);
+chrb = @(x,eps) 2*(eps^2)*sqrt(1+(x^2)/(eps^2))-2*(eps^2)+1;
+chrbp = @(x,eps) 1/sqrt(1+(x^2)/(eps^2));
+prml = @(x,eps) eps^2*log(1+(x^2)/(eps^2));
+prmlp = @(x,eps) 1/(1+(x^2)/(eps^2));
+fnc1 = @(x) chrb(x,1);
+fnp = @(x) prml(x,1);
+fna = @(x) log(abs(x)+1);
+fnq = @(x) x*x;
+x = linspace(-4,4,100);
+y1 = arrayfun(fnc1,x);
+y2 = arrayfun(fna,x);
+y3 = arrayfun(fnq,x);
 
 hold on;
 p1 = plot(x,y1);
[…]

### 2018-07-16 seminar_vof/paper/paper.tex
penalizer function plots

@@
 \centering
-\begin{subfigure}{0.45\linewidth}
+\begin{subfigure}{0.49\linewidth}
 \centering
@@
 \end{subfigure}
-\begin{subfigure}{0.45\linewidth}
+\begin{subfigure}{0.49\linewidth}
 \centering
@@
 robustify both constancy assumptions.
-\begin{flalign*}
+\begin{flalign}\label{eq:charbonnier}
 \charbonnier(s^2) &= 2\epsilon^2\sqrt{1+s^2/\epsilon^2} &&
-\end{flalign*}
-Note that the charbonnier penalizer is a function of a squared argument.
+\end{flalign}
+Note that the charbonnier penalizer is a function of a squared argument so that an arbitrary quadratically penalized assumption can be plugged in.
 When plotting the function with respect to $s$ instead of $s^2$ its behaviour becomes clearer, as shown in \cref{fig:penalizers}.
@@
 Hello
+
+\begin{figure}[htb]
+\centering
+\begin{subfigure}{0.49\linewidth}
+\centering
+\includegraphics[width=\linewidth]{images/charbonnier.eps}
+\end{subfigure}
+\begin{subfigure}{0.49\linewidth}
+\centering
+\includegraphics[width=\linewidth]{images/peronamalik.eps}
+\end{subfigure}
+\caption{
+Graphs show penalizer functions and their asymptotic equivalents as dashed lines.\\
+Left is charbonnier \cref{eq:charbonnier} with $\epsilon=1$ and asymptotic equivalents $\lim_{s\to 0}\charbonnier(s^2)=s^2+2$ and $\lim_{s\to\infty}\charbonnier(s^2)=2s$.\\
+Right is Perona Malik \cref{eq:peronamalik} with $\epsilon=1$ and asymptotic equivalents $\lim_{s\to 0}\peronamalik(s^2)=s^2$ and $\lim_{s\to\infty}\peronamalik(s^2)=2\ln(s)$.
+}
+\label{fig:penalizers}
+\end{figure}
 

### 2018-07-16 seminar_vof/functionplot.m

@@ -1,29 +1,41 @@
 clear;
 clc;
 
-chrb = @(x,eps) 2*(eps^2)*sqrt(1+(x^2)/(eps^2))-2*(eps^2)+1;
+chrb = @(x,eps) 2*(eps^2)*sqrt(1+(x^2)/(eps^2));%-2*(eps^2)+1;
 chrbp = @(x,eps) 1/sqrt(1+(x^2)/(eps^2));
 prml = @(x,eps) eps^2*log(1+(x^2)/(eps^2));
 prmlp = @(x,eps) 1/(1+(x^2)/(eps^2));
 fnc1 = @(x) chrb(x,1);
 fnp = @(x) prml(x,1);
 fna = @(x) log(abs(x)+1);
+fnl = @(x) 2*abs(x)
 fnq = @(x) x*x;
+fnlog = @(x) 2*log(x)
 x = linspace(-4,4,100);
-y1 = arrayfun(fnc1,x);
-y2 = arrayfun(fna,x);
+y1 = arrayfun(fnp,x);
+y2 = arrayfun(fnlog,x);
 y3 = arrayfun(fnq,x);
 
 hold on;
 p1 = plot(x,y1);
-%p3 = plot(x,y2,'--');
+p3 = plot(x,y2,'--');
 p4 = plot(x,y3,'--');
-axis([-inf inf 0 5]);
+axis([0 3 0 4]);
+%axis([0 3 2 6]);
 axis equal
 set(p1, 'Linewidth', 2.5);
 %set(p2, 'Linewidth', 2.5);
-%set(p3, 'Linewidth', 1.3);
+set(p3, 'Linewidth', 1.3);
 set(p4, 'Linewidth', 1.3);
+
+title("PeronaMalik")
+%title("Charbonnier")
+ylabel('y')
+set(get(gca,'ylabel'),'rotation',0)
+xlabel("s")
+print '-S300,400' -deps peronamalik.eps
+
+
 
 
 %T = 1;

### 2018-07-16 seminar_vof/paper/images/fig2eps.m

@@ -1,0 +1,6 @@
+ylim([-2 2])
+axis equal
+xlim([-2 2])
+ylabel("y")
+xlabel("x")
+print "-S300,300" -deps filename.eps

### 2018-07-17 seminar_vof/paper/paper.tex
sum formulas

@@
 \newcommand{\peronamalik}{\psi_{\text{PM}}}
+\newcommand{\Rfirst}{R_{1\text{st}}}
+\newcommand{\Rsecond}{R_{2\text{nd}}}
 
@@
 
-\section{Motivation}
+\section{Motivation}\label{sec:motivation}
 As explained in \cref{sec:variationalopticflow} the regularization term can be used to constrain the \enquote{type} of the estimated flow function.
@@
 The employed data term in the variational model uses a combination of brightness and gradient constancy assumption.
-\begin{flalign*}
+\begin{flalign}
+\label{eq:dataterm}
 D(\flow) &= \int_\Omega \charbonnier(p_{\text{brght}}(\flow)) + \gamma*\charbonnier(p_{\text{grad}}(\flow))\;d\x &&
-\\
+\\\label{eq:brightnessconstancy}
 p_{\text{brght}}(\flow) &= (\;I_1(\x+\flow(\x))-I_0(\x)\;)^2 &&
-\\
+\\\label{eq:gradientconstancy}
 p_{\text{grad}}(\flow) &= ||\;\nabla I_1(\x+\flow(\x))-\nabla I_0(\x)\;||_2^2 &&
-\end{flalign*}
+\end{flalign}
 In this setup, $p_{\text{brght}}$ is the penalizer for the brightness constancy assumption and $p_{\text{grad}}$ is the penalizer for the gradient constancy assumption which can be understood as a check for
@@
 \subsection{Regulalarisation}
-Hello
+As said in \cref{sec:motivation} the variational model is able to adaptively choose regularization order. 
+Therfore the regularisation term is a combination of a first and a second order part.
+Lets first have a look at the first oreder regularizer $\Rfirst$. 
+\begin{flalign}
+\Rfirst(\flow) &= \int_\Omega S_1(\flow)\;d\x&&
+\\
+S_1(\flow) = \sum_{l=1}^2
+\end{flalign}
 

### 2018-07-17 seminar_vof/paper/paper.tex
1st order reg intuition

@@
 
-\subsection{Dataterm}
+\subsection{Dataterm}\label{sec:dataterm}
 The employed data term in the variational model uses a combination of brightness and gradient constancy assumption.
@@
 \Rfirst(\flow) &= \int_\Omega S_1(\flow)\;d\x&&
-\\
-S_1(\flow) = \sum_{l=1}^2
+\\\label{eq:S1}
+S_1(\flow) &= \peronamalik\Big((\ev_1\T\nabla u)^2+(\ev_1\T\nabla v)^2\Big)
++ \charbonnier\Big((\ev_2\T\nabla u)^2+(\ev_2\T\nabla v)^2\Big)&&
 \end{flalign} 
+This is the anisotropic complementary regulariser by Zimmer~et~al.~\cite{opticflowinharmony} which makes use of the eigenvectors of the regularisation tensor \cite{opticflowinharmony}.
+The regularization tensor is a generalization of the structure tensor \cite{maxanisotropy} for arbitrary image features.
+In short, it summarizes gradient information within a neighborhood of a pixel.
+The eigenvector $\ev_1$ points in the direction of steepest change (i.e. over an edge) while $\ev_2$ is orthogonal to $\ev_1$ (i.e. points along an edge).
 
+Lets postpone the discussion of the regularization tensor for a second and get some intuition on the way $S_1$ works.
+As can be seen from \cref{eq:S1}, the term uses two penalizer functions, the Charbonnier penalizer which was introduced in \cref{sec:dataterm} in \cref{eq:charbonnier} and the Perona~Malik penalizer $\peronamalik$.
+\begin{flalign}\label{eq:peronamalik}
+\peronamalik(s^2) &= \epsilon^2\ln(1+s^2/\epsilon^2)&&
+\end{flalign}
+As arguments to the penalizer functions, projections of the flow gradients onto the eigenvectors are used.
+These projections are minimal if the flow gradient vanishes as in standard first order smoothness terms.
+Special about this term are the different penalizations of projections on the first and second eigenvector respectively.
+The graphs of the two penalizer functions are shown in \cref{fig:penalizers}, where it can be seen that values close
+to zero are penalized quadratically, values further away are penalized subquadratically in both functions.
+In the limit however, the Perona-Malik penalizer takes on smaller values than the charbonnier penalizer.
+This has the effect of flow changes in direction of the first eigenvector to be less costly than in direction of the second eigenvector.
+From the perspective of image features and the regularization tensor this is like getting a \enquote{discount} on flow change over an edge.
 \begin{figure}[htb]
@@
 \end{figure}
+
+To continue the discussion of the regularization tensor, lets state its definition first.
+\begin{flalign}
+S_\Phi(\x) = \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
+\end{flalign}
 

### 2018-07-19 seminar_vof/paper/paper.tex
refactored seubsections

@@
 
-\section{Order Adaptive Variational Model}
+\section{Variational Model}
 In this section the variational model by Maurer, Stoll and Bruhn introduced in their work \cite{daspaper} will be discussed in detail.
+As said in \cref{sec:motivation} the variational model is able to adaptively choose regularization order. 
+Therfore the regularisation term is a combination of a first and a second order part.
+First the models data term is discussed followed by the first and second order part of the regularization and their underlying concepts.
 
@@
 
-\subsection{Regulalarisation}
-As said in \cref{sec:motivation} the variational model is able to adaptively choose regularization order. 
-Therfore the regularisation term is a combination of a first and a second order part.
-Lets first have a look at the first oreder regularizer $\Rfirst$. 
+\subsection{First Order Regularizer}
+The first order term of the regularization is the anisotropic complementary regulariser by Zimmer~et~al.~\cite{opticflowinharmony}.
 \begin{flalign}
@@
 \end{flalign} 
-This is the anisotropic complementary regulariser by Zimmer~et~al.~\cite{opticflowinharmony} which makes use of the eigenvectors of the regularisation tensor \cite{opticflowinharmony}.
+This term makes use of the eigenvectors of the regularisation tensor \cite{opticflowinharmony}.
 The regularization tensor is a generalization of the structure tensor \cite{maxanisotropy} for arbitrary image features.
@@
 To continue the discussion of the regularization tensor, lets state its definition first.
+\begin{flalign}\label{eq:structuretensor}
+S_\Phi(\x) &= \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
+&&\\\label{eq:eigendecomp}
+S_\Phi &= Q \Lambda Q\T = \mat{\ev_1&\ev_2}\mat{\lam_1&0\\0&\lam_2}\mat{\ev_1\T\\ \ev_2\T}&&
+\end{flalign}
+The tensor is a convolution of a weighting function $\omega$ with a matrix.
+The matrix is the outer product of the gradient of the selected image feature $\Phi$ which is calculated as $\nabla\Phi(\x)\nabla\Phi(\x)\T$.
+For the structure tensor the image feature would be brightness $\Phi=I$, but as this is the generalization, $\Phi$ can be something else like .
+Using the Eigen value decomposition, the orthonormal eigenvectors $\ev_1$ and $\ev_2$ are obtained which summarize the gradient distribution in the neighborhood defined by the weighting function $\omega$.
+If the weighting function was a Dirac delta $\omega = \delta$ the Eigen decomposition of the tensor would yield the normalized gradient as first eigenvector.
+\begin{flalign*}
+&\int \delta(\tau) \nabla\Phi(\x-\tau)\nabla\Phi(\x-\tau)\T \;d\tau = \nabla\Phi(\x)\nabla\Phi(\x)\T&&
+\\
+&\nabla\Phi\nabla\Phi\T = \lambda_1*\ev_1\ev_1\T = \mat{\ev_1&\ev_2}\mat{\lambda_1&0\\0&0}\mat{\ev_1\T\\ \ev_2\T}&&
+\end{flalign*}
+Here the first eigenvalue is the squared length of the gradient $\lambda_1=|\nabla\Phi|^2$.
+Of course this discussion only holds for Dirac delta as weighting function, for others like Gaussian, the eigenvectors cannot be factored in this way.
+
+\subsection{Second Order Regularizer}
+The second order term uses ...
 \begin{flalign}
-S_\Phi(\x) = \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
+\Rsecond(\flow) = \inf_{\aaux,\baux}[]
 \end{flalign}

### 2018-07-19 seminar_vof/paper/paper.tex
...

@@
 \begin{flalign}\label{eq:structuretensor}
-S_\Phi(\x) &= \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
+\mathcal{S}_\Phi(\x) &= \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
 &&\\\label{eq:eigendecomp}
-S_\Phi &= Q \Lambda Q\T = \mat{\ev_1&\ev_2}\mat{\lam_1&0\\0&\lam_2}\mat{\ev_1\T\\ \ev_2\T}&&
+\mathcal{S}_\Phi &= Q \Lambda Q\T = \mat{\ev_1&\ev_2}\mat{\lam_1&0\\0&\lam_2}\mat{\ev_1\T\\ \ev_2\T}&&
 \end{flalign}
@@
 The matrix is the outer product of the gradient of the selected image feature $\Phi$ which is calculated as $\nabla\Phi(\x)\nabla\Phi(\x)\T$.
-For the structure tensor the image feature would be brightness $\Phi=I$, but as this is the generalization, $\Phi$ can be something else like .
-Using the Eigen value decomposition, the orthonormal eigenvectors $\ev_1$ and $\ev_2$ are obtained which summarize the gradient distribution in the neighborhood defined by the weighting function $\omega$.
-If the weighting function was a Dirac delta $\omega = \delta$ the Eigen decomposition of the tensor would yield the normalized gradient as first eigenvector.
+For the structure tensor the image feature would be brightness $\Phi=I$, but as this is the generalization, $\Phi$ can be something else like the saturation value in HSV color space or luminance of CIE L*a*b*.
+Using the eigendecomposition, the orthonormal eigenvectors $\ev_1$ and $\ev_2$ are obtained which summarize the gradient distribution in the neighborhood defined by the weighting function $\omega$.
+If the weighting function was a Dirac delta $\omega = \delta$ the eigendecomposition of the tensor would yield the normalized gradient as first eigenvector.
 \begin{flalign*}
@@
 \begin{flalign}
-\Rsecond(\flow) = \inf_{\aaux,\baux}[]
+\Rsecond(\flow) =\;&\int_\Omega\; \inf_{\aaux,\baux}\big[\;S_2(\flow,\aaux,\baux) +\beta*S_\text{aux}(\aaux,\baux)\;\big] \;d\x
+&&\\
+S_2(\flow,\aaux,\baux) =\;& \peronamalik\Big( (\ev_1\T(\nabla u-\aaux))^2 + (\ev_1\T(\nabla v-\baux))^2 \Big)
+&&\\
++\;&\charbonnier\Big( (\ev_2\T(\nabla u-\aaux))^2 + (\ev_2\T(\nabla v-\baux))^2 \Big)
+&&\\
+S_\text{aux}(\aaux,\baux) =\;&
+&&
 \end{flalign}

### 2018-07-19 seminar_vof/paper/paper.tex
second order term mathized

@@
 \usepackage[english]{babel}
+
+\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
 
@@
 \newcommand{\Rsecond}{R_{2\text{nd}}}
+\newcommand{\jacob}{\mathcal{J}}
 
@@
 
-In order to minimize the energy functional, an iterative update sheme is developed using the euler lagrange equation \cite{eulerlagrange}.
+In order to minimize the energy functional, an iterative update sheme is developed using the Euler-Lagrange equation \cite{mathforphysicists}.
 
@@
 \begin{flalign}\label{eq:structuretensor}
-\mathcal{S}_\Phi(\x) &= \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
+\mathcal{S}_{\Phi,\omega}(\x) &= \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
 &&\\\label{eq:eigendecomp}
-\mathcal{S}_\Phi &= Q \Lambda Q\T = \mat{\ev_1&\ev_2}\mat{\lam_1&0\\0&\lam_2}\mat{\ev_1\T\\ \ev_2\T}&&
+\mathcal{S}_{\Phi,\omega}&= Q \Lambda Q\T = \mat{\ev_1&\ev_2}\mat{\lam_1&0\\0&\lam_2}\mat{\ev_1\T\\ \ev_2\T}&&
 \end{flalign}
@@
 For the structure tensor the image feature would be brightness $\Phi=I$, but as this is the generalization, $\Phi$ can be something else like the saturation value in HSV color space or luminance of CIE L*a*b*.
-Using the eigendecomposition, the orthonormal eigenvectors $\ev_1$ and $\ev_2$ are obtained which summarize the gradient distribution in the neighborhood defined by the weighting function $\omega$.
+Using the eigendecomposition (\cref{eq:eigendecomp}), the orthonormal eigenvectors $\ev_1$ and $\ev_2$ are obtained which summarize the gradient distribution in the neighborhood defined by the weighting function $\omega$.
 If the weighting function was a Dirac delta $\omega = \delta$ the eigendecomposition of the tensor would yield the normalized gradient as first eigenvector.
 \begin{flalign*}
-&\int \delta(\tau) \nabla\Phi(\x-\tau)\nabla\Phi(\x-\tau)\T \;d\tau = \nabla\Phi(\x)\nabla\Phi(\x)\T&&
-\\
-&\nabla\Phi\nabla\Phi\T = \lambda_1*\ev_1\ev_1\T = \mat{\ev_1&\ev_2}\mat{\lambda_1&0\\0&0}\mat{\ev_1\T\\ \ev_2\T}&&
+\mathcal{S}_{\Phi,\delta}= \nabla\Phi\nabla\Phi\T = \lambda_1*\ev_1\ev_1\T = \mat{\ev_1&\ev_2}\mat{\lambda_1&0\\0&0}\mat{\ev_1\T\\ \ev_2\T}&&
 \end{flalign*}
@@
 +\;&\charbonnier\Big( (\ev_2\T(\nabla u-\aaux))^2 + (\ev_2\T(\nabla v-\baux))^2 \Big)
+\nonumber&&\\
+S_\text{aux}(\aaux,\baux) =\;& \peronamalik\Big(\sum\nolimits_{k=1}^2\;(\ev_k\T \jacob_{\aaux}\;\ev_1)^2 + (\ev_k\T \jacob_{\baux}\;\ev_1)^2  \Big)
 &&\\
-S_\text{aux}(\aaux,\baux) =\;&
-&&
++\;&\charbonnier\Big(\sum\nolimits_{k=1}^2\;(\ev_k\T \jacob_{\aaux}\;\ev_2)^2 + (\ev_k\T \jacob_{\baux}\;\ev_2)^2  \Big)
+\nonumber&&
 \end{flalign}

### 2018-07-21 seminar_vof/paper/paper.tex
more second order term expl

@@
 In the limit however, the Perona-Malik penalizer takes on smaller values than the charbonnier penalizer.
-This has the effect of flow changes in direction of the first eigenvector to be less costly than in direction of the second eigenvector.
+This has the effect of flow changes tangent to the first eigenvector to be less costly than tangent to the second eigenvector.
 From the perspective of image features and the regularization tensor this is like getting a \enquote{discount} on flow change over an edge.
@@
 \subsection{Second Order Regularizer}
-The second order term uses ...
+The second order regularizer makes use of two auxiliary functions $\aaux(\x)$ and $\baux(\x)$ which are connected
+to the flow $\flow$ using a coupling term $S_2$. 
+These auxiliary functions are as well unknown and are controlled by an auxiliary smoothness term $S_\text{aux}$.
 \begin{flalign}
@@
 \end{flalign}
+This regularization term can be interpreted as the integral over the infimal convolution of $S_2$ and $S_\text{aux}$ with respect to the vector fields $\aaux$ and $\baux$.
+
+The coupling term $S_2$ is similar to the first order term $S_1$ in that it uses the two penalizer functions $\peronamalik$ and $\charbonnier$ to penalize non zero projections with respect to the eigenvectors of the structure tensor (\cref{eq:structuretensor}).
+This time however, not the flow gradients themselves are projected, but the differece of the gradients to the auxiliary functions.
+The coupling term therefore becomes minimal if the gradients are euqal to the auxiliary functions instead of equal to zero as in the first order term.
+Again, the prejection on the first eigenvector is penalized less due to the behavior of the perona malik penalizer, but is not intuitive in the form written above.
+If instead we reformulate the projection as $(\ev_1\T(\nabla u-\aaux))^2 = (\ev_1\T\nabla u-\ev_1\T\aaux)^2$ it becomes clear that the flow gradient and the auxiliary function are allowed to differ as long as they are roughly tangent to the first eigenvector.
+
+Due to the coupling term connecting the first derivative of the flow to the auxiliary functions, the
 

### 2018-07-21 seminar_vof/paper/paper.tex
variational model complete, next up: order adaption

@@
 Different methods for flow estimation exist, such as ... ... and variational optic flow which is based on the calculus of variations.
-In the following the concept of variational flow estimation is recapitulated and the approach by \verb|\cite{daspaper}| with order adaptive regulalarization is presented.
+In the following the concept of variational flow estimation is recapitulated and the approach by Maurer, Stoll and Bruhn \cite{daspaper} with order adaptive regulalarization is presented.
 
@@
 \\\\
-Given two subsequent images
-$I_0,I_1 ~:~ \Omega \subset \Real^2 \to \Real$ \\
-of an image sequence, we seek the displacement vector field 
-\\$\flow = (u~v)\T ~:~ \Omega \subset \Real^2 \to \Real^2$ 
-that maps 
-$I_0$ onto $I_1$ 
-such that\\
-$I_0(\x) = I_1(\x+\flow(\x))$.
+Given two subsequent images of an image sequence\\
+$I_0,I_1 ~:~ \Omega \subset \Real^2 \to \Real$,\\
+we seek the displacement vector field 
+\\$\flow = (u~v)\T ~:~ \Omega \subset \Real^2 \to \Real^2$\\
+that maps $I_0$ onto $I_1$ such that
+\\$I_0(\x) = I_1(\x+\flow(\x))$.
 \\\\
@@
 Graphs show vector fields of different behaviour, left is constant with $f(x,y) = (1\;\, 1)\T$, right is linear with $f(x,y) = (x \;\, y)\T$.
-Constant flow is typical for motion parallel to the image plane, linear flow for motion perpenicular to the image plane. 
+Constant flow is typical for motion parallel to the image plane, linear flow for motion orthogonal to the image plane. 
 }
@@
 \begin{flalign}\label{eq:structuretensor}
-\mathcal{S}_{\Phi,\omega}(\x) &= \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 && \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) && \Phi_y(\x-\tau)^2} \;d\tau
+\mathcal{S}_{\Phi,\omega}(\x) &= \bigintsss \omega(\tau)\Mat{\Phi_x(\x-\tau)^2 & \Phi_x(\x-\tau)\Phi_y(\x-\tau)\\\Phi_x(\x-\tau)\Phi_y(\x-\tau) & \Phi_y(\x-\tau)^2} \;d\tau
 &&\\\label{eq:eigendecomp}
@@
 For the structure tensor the image feature would be brightness $\Phi=I$, but as this is the generalization, $\Phi$ can be something else like the saturation value in HSV color space or luminance of CIE L*a*b*.
+Typically, the feature is chosen to be consistent with the feature used in the data term, which would be a combination of brightness and gradient norm.
 Using the eigendecomposition (\cref{eq:eigendecomp}), the orthonormal eigenvectors $\ev_1$ and $\ev_2$ are obtained which summarize the gradient distribution in the neighborhood defined by the weighting function $\omega$.
@@
 
-Due to the coupling term connecting the first derivative of the flow to the auxiliary functions, the
+In the auxiliary term $S_\text{aux}$ the Jacobians of the auxiliary functions $\jacob_{\aaux}$ and $\jacob_{\baux}$ are anisotropically penalized similar to the first order penalization of the flow gradients.
+Obviously, the auxiliary term is minimal if the derivatives of the auxiliary functions vanish, allowing for constant auxiliary functions with changes at edges which are indicated by the regularization tensor.
+Due to the connection of the auxiliary functions to the flow gradients in the coupling term, the flow is indirectly constrained to be linear.
+This is because the nullspace of the second order term are functions with derivatives equal to the auxiliary functions which are constrained to be constant.
 

### 2018-07-23 seminar_vof/paper/paper.tex
order adaption

@@
 
-\section{Variational Model}
+\section{Variational Model}\label{sec:variationalmodel}
 In this section the variational model by Maurer, Stoll and Bruhn introduced in their work \cite{daspaper} will be discussed in detail.
@@
 
-\subsection{Second Order Regularizer}
+\subsection{Second Order Regularizer}\label{subsec:secondorderreg}
 The second order regularizer makes use of two auxiliary functions $\aaux(\x)$ and $\baux(\x)$ which are connected
@@
 This is because the nullspace of the second order term are functions with derivatives equal to the auxiliary functions which are constrained to be constant.
+
+\section{Order Adaption}
+In \cref{sec:variationalmodel}, the data and regularization terms were discussed and it was explained that the first order regularizer enforces constant flow whereas the second order regularizer allows for linear flow.
+As illustrated in \cref{fig:flowfields} constant vector fields are suitable for modelling motion parallel to the image plane, motion orthogonal to the image plane can be modelled using a linear flow field.
+So what is needed to apply the most suitable regularizer depending on the current sequence, is an adaptive scheme which will be explained in the following.
+
+\subsection{Global Scheme}
+The global adaptive scheme will be choosing order once for the whole image domain.
+To do so a general regularization term consisting of both, first and second order, will be extended.
+The general regularization term is a simple convex combination of the terms with weighting parameter $c\in [0,1]$ and selecton term $\phi_\lambda(c)$ of the following form.
+\begin{flalign}
+R(\flow,c) = \int_\Omega \inf_{\aaux,\baux}\big[\; c*S_1(\flow) 
+& + (1-c)*S_2(\flow,\aaux,\baux) 
+&&\\
+& + \beta*S_\text{aux}(\aaux,\baux)
++ \phi_\lambda(c)\;\big] \;d\x \nonumber&&
+\end{flalign}
+It may be counter intuitive to only include $S_1$ and $S_2$ in the convex combination and not $S_\text{aux}$, but $S_\text{aux}$ is independent of $\flow$ and merely used to constrain the auxiliary functions.
+$S_1$ and $S_2$ on the other hand are the energies with respect to the flow and are comparable to each other as they only differ in the projected quantity, which was already discussed in \cref{subsec:secondorderreg}.
+The goal thus, is to embed some information into the the weighting parameter $c$ such that it equals to $1$ when first order regularization has less energy than second order regularization and equals to $0$ in the contrary case.
+However, since the class of linear functions which can be modelled using second order regularization includes constant functions, it is only desirable to use second order regularization when it yields a significant benefit over first order.
+To model this behavour of $c$ a sigmoid function of the benefit $\Delta$ of first over second order regularization is used.
+\begin{flalign}
+c = \frac{1}{1+e^{-\Delta/\lambda}} \quad\text{with}\quad\Delta=T+\frac{1}{|\Omega|}\int_\Omega S_2 - S_1 \; d\x &&
+\end{flalign}
+Here $\lambda$ is used to control the slope of the sigmoid function, and $T$ is the threshold by which the second order regularizer has to be better than first order, so that second order will be used.
+
+To get to this behavior of $c$ from the euler lagrange equation's point of view, the corresponding selection term $\phi_\lambda(c)$ has to be found.
+By examining the regularizers partial derivative $\frac{\partial R}{\partial c} = 0$ we get to the following equation.
+\begin{flalign}
+&\frac{\partial R}{\partial c} = \int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) + \phi_\lambda'(c) \;d\x = 0
+&&\nonumber\\
+&-|\Omega|*\phi_\lambda'(c) = \int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux)
+&&
+\end{flalign}
 

### 2018-07-23 seminar_vof/paper/paper.tex
order adaption moore

@@
 To model this behavour of $c$ a sigmoid function of the benefit $\Delta$ of first over second order regularization is used.
-\begin{flalign}
+\begin{flalign}\label{eq:c}
 c = \frac{1}{1+e^{-\Delta/\lambda}} \quad\text{with}\quad\Delta=T+\frac{1}{|\Omega|}\int_\Omega S_2 - S_1 \; d\x &&
@@
 \begin{flalign}
-&\frac{\partial R}{\partial c} = \int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) + \phi_\lambda'(c) \;d\x = 0
+\frac{\partial R(\flow,c)}{\partial c} &= \int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) + \phi_\lambda'(c) \;d\x \;=\; 0
 &&\nonumber\\
-&-|\Omega|*\phi_\lambda'(c) = \int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux)
+-\phi_\lambda'(c) &= \frac{1}{|\Omega|}\int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) \;d\x
+&&\nonumber\\
+\label{eq:deriv_r_to_c}
+\phi_\lambda'(c) &= \frac{1}{|\Omega|}\int_\Omega S_2(\flow,\aaux,\baux) - S_1(\flow) \;d\x
 &&
+\end{flalign}
+The right hand side of \cref{eq:deriv_r_to_c} also occurs in \cref{eq:c} which can be reformulated as follows in order to be plugged in.
+\begin{flalign}
+&\frac{1}{|\Omega|}\int_\Omega S_2 - S_1\;d\x = \Delta-T
+%&&\nonumber\\
+\quad\text{with}\quad
+\Delta = -\lambda*\ln(\frac{1}{c}-1)
+&&\nonumber\\\label{eq:phi'}
+&\phi_\lambda'(c) = -\lambda*\ln\Big(\frac{1}{c}-1\Big) - T
+\end{flalign}
+To obtain $\phi_\lambda$ we integrate both sides of \cref{eq:phi'} and get the following selection term with integration constant $C$.
+\begin{flalign}
+\phi_\lambda(c) = \lambda \Bigg( \ln(1-c)-c*\ln\Big(\frac{1}{c}-1\Big) \Bigg) -Tc + C &&
+\end{flalign}
+When chosing $C=T$ we can get a more intuitive formulation of the regularizer, using a cleverly tailored replacement $\phi$ for $\phi_\lambda$.
+\begin{flalign}
+\phi(c) &= \frac{\phi_\lambda(c) + T*(c-1)}{\lambda} = \ln(1-c)-c*\ln\Big(\frac{1}{c}-1\Big)
+&&\\
+R(\flow,c) = R
 \end{flalign}

### 2018-07-24 seminar_vof/paper/paper.tex
more schemes

@@
 %% author name
-\author{Roy G. Biv}
+\author{David H\"agele}
 
 %% paper title
-\title{Order-Adaptive Regularisation for Variational Optic Flow:\\Global, Local and in Between}
+\title{An Elaboration on\\Order-Adaptive Regularisation for Variational Optical Flow:\\Global, Local and in Between}
 
@@
 
+\subsubsection{Regularization Tensor}
 To continue the discussion of the regularization tensor, lets state its definition first.
@@
 As illustrated in \cref{fig:flowfields} constant vector fields are suitable for modelling motion parallel to the image plane, motion orthogonal to the image plane can be modelled using a linear flow field.
-So what is needed to apply the most suitable regularizer depending on the current sequence, is an adaptive scheme which will be explained in the following.
+So what is needed to apply the most suitable regularizer depending on the current sequence, is an adaptive scheme.
+In the following four different schemes will be discussed, where the global scheme will be used to introduce the underlying concept of the schemes.
 
@@
 However, since the class of linear functions which can be modelled using second order regularization includes constant functions, it is only desirable to use second order regularization when it yields a significant benefit over first order.
-To model this behavour of $c$ a sigmoid function of the benefit $\Delta$ of first over second order regularization is used.
+To model this behavour of $c$ a sigmoid function of the benefit $\Delta$ of first over second order regularization is used as illustrated in \cref{fig:c_phi}.
 \begin{flalign}\label{eq:c}
@@
 \begin{flalign}
-\frac{\partial R(\flow,c)}{\partial c} &= \int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) + \phi_\lambda'(c) \;d\x \;=\; 0
+\frac{\partial R(\flow,c)}{\partial c} = &\int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) + \phi_\lambda'(c) \;d\x \;=\; 0
 &&\nonumber\\
--\phi_\lambda'(c) &= \frac{1}{|\Omega|}\int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) \;d\x
+-\phi_\lambda'(c)*|\Omega| = &\int_\Omega S_1(\flow) - S_2(\flow,\aaux,\baux) \;d\x
 &&\nonumber\\
 \label{eq:deriv_r_to_c}
-\phi_\lambda'(c) &= \frac{1}{|\Omega|}\int_\Omega S_2(\flow,\aaux,\baux) - S_1(\flow) \;d\x
+\phi_\lambda'(c) = \frac{1}{|\Omega|}&\int_\Omega S_2(\flow,\aaux,\baux) - S_1(\flow) \;d\x
 &&
@@
 When chosing $C=T$ we can get a more intuitive formulation of the regularizer, using a cleverly tailored replacement $\phi$ for $\phi_\lambda$.
+\begin{flalign}\label{eq:phi}
+\phi(c) = \frac{\phi_\lambda(c) + T*(c-1)}{\lambda} &= \ln(1-c)-c*\ln\Big(\frac{1}{c}-1\Big)
+&&\\
+%= \ln(1-c)-c*\ln\Big(\frac{1-c}{c}\Big) 
+& = c*\ln(c) + (1-c)*\ln(1-c)
+%-c ln(1/c -1 ) = -c ln((1-c)/c) = -c ln(1-c) + c ln(c)
+&&\nonumber\\
+R(\flow,c) = \int_\Omega \inf_{\aaux,\baux}\big[\; c*S_1(\flow) 
+& + (1-c)*\big(S_2(\flow,\aaux,\baux) +T\big)
+&&\\
+& + \beta*S_\text{aux}(\aaux,\baux)
++ \lambda*\phi(c)\;\big] \;d\x \nonumber&&
+\end{flalign} 
+In this representation the threshold parameter $T$ appears in the convex combination together with $S_2$ and can be interpreted as an extra cost to be paid in order to use second order regularization.
+The sigmoid slope parameter $\lambda$ now works as a weight for the selection term $\phi$ which has become independent of $\lambda$.
+In fact the selection term has simplified quite a bit and it can be seen that it is a symmetric function which is also illustarted in \cref{fig:c_phi}.
+From the figure it can also be seen that $\phi$ is minimal for $c=\frac{1}{2}$ which can be interpreted as a prior on $c$ for when the model is indecisive on regularization order.
+So by definition of $\phi$ neither of the regularizers is preferred, but due to the threshold $T$ the descision is biased towards the first order regularization.
+
+\begin{figure}[htb]
+\centering
+\begin{subfigure}{0.49\linewidth}
+\centering
+\includegraphics[width=\linewidth]{images/cofdelta.eps}
+\end{subfigure}
+\begin{subfigure}{0.49\linewidth}
+\centering
+\includegraphics[width=\linewidth]{images/selectionterm.eps}
+\end{subfigure}
+\caption{
+The left graph shows the weighting parameter $c$ (\cref{eq:c}) with respect to $\Delta$, where $\Delta$ is the benefit of first over second order regularization biased by the threshold $T$.
+The solid line corresponds to $\lambda=0.2$, the dotted line to $\lambda=0.8$.
+\\
+The right graph shows the selection term $\phi(c)$ (\cref{eq:phi}) which is axially symmetric to $c=\frac{1}{2}$.
+}
+\label{fig:c_phi}
+\end{figure}
+
+\subsection{Local Scheme}
+The previously introduced global scheme chooses regularization for the whole image domain, however, different parts of the scene may move differently so that it makes sense to
+allow for different regularization depending on the location.
+So instead of using a global weighting parameter $c$, it is replaced by a weighting function $c_{\text{local}}(\x)$.
 \begin{flalign}
-\phi(c) &= \frac{\phi_\lambda(c) + T*(c-1)}{\lambda} = \ln(1-c)-c*\ln\Big(\frac{1}{c}-1\Big)
-&&\\
-R(\flow,c) = R
+c_{\text{local}}(\x) = \frac{1}{1+e^{-\Delta(\x)/\lambda}} \quad\text{with}\quad\Delta(\x)=T + S_2 - S_1&&
 \end{flalign}
+As a consequence $\Delta$ is now a spacially varying function as well, mapping to the benefit of first order regularization in a single pixel at $\x$, which allows the model to decide on order on a per pixel basis.
+
+\subsection{Non-Local Scheme}
+Since the local scheme allows for different regularization order per pixel based on a decision which only takes that pixels location into account, regularization order may fluctuate a lot in this model.
 

### 2018-07-24 seminar_vof/paper/images/selectionterm.m

@@ -1,0 +1,48 @@
+clear;
+clc;
+
+phi = @(c) c*log(c) + (1-c)*log(1-c);
+dphi = @(c) log(1/c-1);
+cofdelta = @(delta) 1/(1+exp(-delta/0.2));
+cofdelta2 = @(delta) 1/(1+exp(-delta/0.8));
+
+x = linspace(0,1,100);
+y = arrayfun(phi,x);
+y1 = arrayfun(cofdelta,x);
+y2 = arrayfun(cofdelta2,x);
+
+p1 = plot(x,y);
+set(p1, 'Linewidth', 2.5);
+axis([0 1 -0.75 0])
+axis equal;
+title(" ")
+ylabel('')
+xlabel('c')
+set(get(gca,'ylabel'),'rotation',0)
+%print '-S300,225' -deps selectionterm.eps
+
+
+
+x = linspace(-3,3,100);
+y = arrayfun(phi,x);
+y1 = arrayfun(cofdelta,x);
+y2 = arrayfun(cofdelta2,x);
+
+p1 = plot(x,y1);
+hold on;
+p2 = plot(x,y2, "--");
+set(p1, 'Linewidth', 2.5);
+set(p2, 'Linewidth', 2.5);
+axis([-2 2 -1 2])
+axis equal;
+title(" ")
+ylabel('c')
+xlabel('delta')
+set(gca,'YTick',[-1 0 1 2])
+set(get(gca,'ylabel'),'rotation',0)
+%print '-S300,225' -deps cofdelta.eps
+hold off;
+
+phiofdelta = @(d) 0.8*phi(cofdelta2(d));
+y = arrayfun(phiofdelta,x);
+plot(x,y)

### 2018-07-27 seminar_vof/paper/paper.tex
more schemes

@@
 \newcommand{\jacob}{\mathcal{J}}
+\newcommand{\patch}{\mathcal{N}}
+\newcommand{\cnl}{c_{\text{nl}}}
+\newcommand{\cNl}{\overset{\_}{c}_{\text{nl}}}
 
@@
 Since the local scheme allows for different regularization order per pixel based on a decision which only takes that pixels location into account, regularization order may fluctuate a lot in this model.
+To make the descision more consistent with the pixels neighorhood, the weighting function $c_{\text{local}}(\x)$ is altered to
+take a rectangular shaped patch of pixels around it into account.
+\begin{flalign}
+\cNl(\x) = \frac{1}{|\patch(\x)|}\int_{\patch(\x)}\cnl(\y)\;d\y 
+&&\\
+R_\text{nl}(\flow,\cnl) = \int_\Omega \inf_{\aaux,\baux}\big[\; \cNl*S_1(\flow) 
+& + (1-\cNl)*\big(S_2(\flow,\aaux,\baux) +T\big)
+&&\\
+& + \beta*S_\text{aux}(\aaux,\baux)
++ \lambda*\phi(\cnl)\;\big] \;d\x \nonumber&&
+\end{flalign}
+Here $\mathcal{N}(\x)$ denotes the patch of $\x$ and $|\mathcal{N}(\x)|$ the patch size which is used for normalization.
+To find $\cnl$, $R_\text{nl}(\flow,\cnl)$ is minimized with respect to it.
+\begin{flalign}
+\frac{\partial R(\flow,\cnl)}{\partial \cnl} = \cdots
+&&\\
+\text{derivation of~} \cnl \dots
+&&\\
+\cnl = \cdots
+&&
+\end{flalign}
+
+\subsection{Region-Based Scheme}
+The non-local scheme takes a neighborhood around a pixel into account for the descision on regularization order.
+To push it yet a little further, the region based scheme decides on order leveraging a level set function $z(\x):\Omega\to\mathbb{R}$ which devides the image space into first and second order regularized regions.
+The corresponding regularizer reads as follows.
+\begin{flalign}
+&R(\flow,z) = \int_\Omega c_\text{rgn}(z)*S_1 + (1-c_\text{rgn}(z))*(S_2 + T) + \theta*|\nabla c_\text{rgn}(z)| \;d\x
+&&\\
+&c_\text{rgn}(z) = \frac{1}{1+e^{z/\lambda}}
+&&
+\end{flalign}
+Here $c_\text{rgn}$ is as well a sigmoid function, but choosen to approximate the heaviside funtion ($\lambda\to 0$) in order to obtain hard boundaries for the regions defined by $z$.
+The selection term from the previous schemes has been replaced by a smoothness term on $c_\text{rgn}$ which penalizes level set functions with lots of sign changes.
+Due to this penalization short paths for $z(\x)=0$ are favored resulting in smooth region boundaries.
 

### 2018-07-28 mlr_summary.tex
started mlrsummary

% !TeX spellcheck = en_GB
\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{esvect}
\setlength{\parindent}{0pt}
\setlength{\parskip}{.36cm plus0.09cm minus0.09cm}
\usepackage[left=2.25cm, right=2.25cm, top=2.25cm, bottom=2.25cm]{geometry}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{esvect}
\usepackage{csquotes}

\newcommand{\lam}{\lambda}
\newcommand{\trans}{\quad\bigg|\;}
\newcommand{\Trans}{\quad\Bigg|\;}
% matrix shortcut
\newcommand{\mat}[1]{\begin{pmatrix} #1 \end{pmatrix}}
% matrix shortcut with increased vertical spacing
\newcommand{\Mat}[1]{\begingroup
\renewcommand*{\arraystretch}{1.5}
\mat{#1}
\endgroup}
% matrix format but without braces (secretly a matrix)
\newcommand{\secretMat}[2][r]{\begin{matrix*}[#1] #2 \end{matrix*}}
% matrix format but without braces (secretly a matrix) with increased vertical spacing
\newcommand{\SecretMat}[2][r]{\begingroup
\renewcommand*{\arraystretch}{1.7}
\secretMat[#1]{#2}
\endgroup}
% shortcuts for df/dx stuff in derivative
\newcommand{\der}{\partial}
\newcommand{\deriv}[2]{\frac{\der #1}{\der #2}}
\newcommand{\eqns}[1]{\begin{flalign} #1 \end{flalign}}
\newcommand{\eqnsnn}[1]{\begin{flalign*} #1 \end{flalign*}}
\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\dom}[1]{\mathbb{#1}}
\newcommand{\T}{^\top}
\newcommand{\equivalent}{\Leftrightarrow}
% make star '*' become \cdot (normal multiplication sign)
\mathcode`\*="8000
{\catcode`\*\active\gdef*{\cdot}}

\begin{document}
  \begin{center}
    \Large\textbf{Summary - Machine Learning}\\
    \vspace{0.25cm}
    \large{David H\"agele}, \normalsize{\today}
   \end{center}

\section{Machine Learning Briefly}
The goal of machine learning is to estimate or \enquote{learn} a mapping $f$ from one domain $\dom{X}$ to another $\dom{Y}$ by using a set of data points.
\eqnsnn{
f : \dom{X} \to \dom{Y} &&
}
In the field of supervised learning a data set $D$ usually contains observed data points of domain $\dom{X}$ and corresponding points of domain $\dom{Y}$. 
The objective is to find a mapping $f_\text{optimal}$ that minimizes an error $err$ between observed and estimated $y$.
\eqnsnn{&
D = \{ x_i , y_i \}\quad|\quad x_i\in\dom{X},~ y_i \in \dom{Y},~ i\in\dom{N}
&&\\&
err : \dom{Y}\times\dom{Y} \to \dom{R}
&&\\&
f_\text{optimal} = \argmin_f\;err(y,f(y))
&&}
In the field of unsupervised learning there are no corresponding observed points of domain $\dom{Y}$.
Optimality is not measured using an error to the observation, instead other objectives are to be fulfilled.

\section{Linear Regression}
One of the most basic algorithms for estimating a real valued function is linear regression.
Given a set of multivariate inputs $x_i \in \dom{R}^m$ and corresponding real valued outputs $y_i \in \dom{R}$ the following \enquote{loss} $L$ is to be minimized by a linear function $f_\beta:\dom{R}^m\to\dom{R}$.
\eqns{&
L(\beta) = \sum_{i=1}^n (\;y_i - f_\beta(x_i)\;)^2 
&&\\&
f_\beta(x) = x\T\beta \quad\text{with}\quad\beta\in\dom{R}^m
&&}
Lets call the matrix of transposed $x_i$'s $X$ and the vector of $y_i$'s $Y$.
This allows for rewriting the loss as follows.
\eqnsnn{&
L(\beta) = (Y-X\beta)\T(Y-X\beta)
&&\\&
X = \mat{x_1\T\\\vdots\\x_n\T} \quad Y = \mat{y_1\\\vdots\\y_n}
&&}
When minimizing we seek the roots of the derivative of the loss, by doing so we can derive the optimal solution for beta.
\eqnsnn{
\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ (Y-X\beta)\T X + (Y-X\beta)\T X = 2*(Y-X\beta)\T X = 0
&&\\&
\equivalent (Y-X\beta)\T X = X\T Y - X\T X\beta = 0
&&\\&
\equivalent X\T Y = X\T X\beta
&&\\&
\equivalent (X\T X)^{-1}X\T Y = \beta
&&
}
\subsection{Non-Linear Features}
Sadly the very easy linear regression algorithm only allows for linear functions to be estimated. 
Fortunately we can make things non-linear by defining a feature function which maps an input $x$ to feature space like so:
\eqnsnn{
\phi(x) \in \dom{R}^l \quad\text{with}\quad l\in\dom{N}
&&}
For real valued $x\in\dom{R}$ we can create the quadratic feature vector using the following.
\eqnsnn{
\phi(x) = \mat{1&x&x^2}\T
&&}
When using linear regression on these features we end up with a function that is quadratic in $x$ but linear in features $\phi(x)$.
\eqnsnn{&
X = \mat{\phi(x_1)\T\\\vdots\\\phi(x_n)\T}
&&\\&
f_\beta(x) = \phi(x)\T\beta = \phi(x)_1*\beta_1 + \phi(x)_2*\beta_2 + \phi(x)_3*\beta_3 = \beta_1 + x\beta_2 + x^2\beta_3
&&
}

\subsection{Regularization}
Maybe you may have noticed that the linear regression algorithm may fail in case that $(X\T X)$ is not invertible.
To ensure invertibility a tiny quantity can be added to the diagonal of the matrix $(X\T X + \lambda I)$.
Interestingly, the optimal solution to the following loss has this exact form.
\eqnsnn{&
L(\beta) = (Y-X\beta)\T(Y-X\beta) + \lambda*\beta\T\beta
&&\\&
\deriv{L(\beta)}{\beta} = 0 \equivalent 2(Y-X\beta)X\T + 2\lambda*\beta\T = 0
\equivalent (Y-X\beta)X\T + \lambda*\beta\T = 0
&&}

\end{document}

### 2018-07-28 mlr_summary.tex
linreg done, also done features,regularization and cross validation

@@
 \usepackage{csquotes}
+\usepackage{algpseudocode}
 
@@
 &&}
-Lets call the matrix of transposed $x_i$'s $X$ and the vector of $y_i$'s $Y$.
+This is the mean squared error and we want to minimize it using a linear function.
+Lets call the matrix of transposed $x_i$'s $X$ and the vector of $y_i$'s $y$.
 This allows for rewriting the loss as follows.
 \eqnsnn{&
-L(\beta) = (Y-X\beta)\T(Y-X\beta)
+L(\beta) = (y-X\beta)\T(y-X\beta)
 &&\\&
-X = \mat{x_1\T\\\vdots\\x_n\T} \quad Y = \mat{y_1\\\vdots\\y_n}
+X = \mat{x_1\T\\\vdots\\x_n\T} \quad y = \mat{y_1\\\vdots\\y_n}
 &&}
@@
 \eqnsnn{
-\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ (Y-X\beta)\T X + (Y-X\beta)\T X = 2*(Y-X\beta)\T X = 0
+\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ (y-X\beta)\T X + (y-X\beta)\T X = 2*(y-X\beta)\T X = 0
 &&\\&
-\equivalent (Y-X\beta)\T X = X\T Y - X\T X\beta = 0
+\equivalent (y-X\beta)\T X = X\T y - X\T X\beta = 0
 &&\\&
-\equivalent X\T Y = X\T X\beta
+\equivalent X\T y = X\T X\beta
 &&\\&
-\equivalent (X\T X)^{-1}X\T Y = \beta
+\equivalent (X\T X)^{-1}X\T y = \beta
 &&
@@
 \subsection{Non-Linear Features}
-Sadly the very easy linear regression algorithm only allows for linear functions to be estimated. 
+Sadly the very simple linear regression algorithm only allows for linear functions to be estimated. 
 Fortunately we can make things non-linear by defining a feature function which maps an input $x$ to feature space like so:
@@
 }
+Even more sophisticated features could be using gaussian functions at the datapoints which then amounts to a radial basis function.
+\eqnsnn{&
+\phi(x) = \mat{1 & \phi_1(x) & \cdots & \phi_N(x)}\T \quad\text{with}\quad \phi_i(x) = e^{-\frac{1}{2}||x_i-x||_2^2}
+&&}
 
@@
 Maybe you may have noticed that the linear regression algorithm may fail in case that $(X\T X)$ is not invertible.
-To ensure invertibility a tiny quantity can be added to the diagonal of the matrix $(X\T X + \lambda I)$.
+To ensure invertibility a tiny quantity can be added to or subtracted from the diagonal of the matrix like so: $(X\T X - \lambda I)$.
 Interestingly, the optimal solution to the following loss has this exact form.
-\eqnsnn{&
-L(\beta) = (Y-X\beta)\T(Y-X\beta) + \lambda*\beta\T\beta
-&&\\&
-\deriv{L(\beta)}{\beta} = 0 \equivalent 2(Y-X\beta)X\T + 2\lambda*\beta\T = 0
-\equivalent (Y-X\beta)X\T + \lambda*\beta\T = 0
+\eqnsnn{
+L(\beta) = (y&-X\beta)\T(y-X\beta) + \lambda*\beta\T\beta
+&&\\
+\deriv{L(\beta)}{\beta} = 0 &\equivalent 2(y-X\beta)\T X + 2\lambda*\beta\T = 0
+\equivalent (y-X\beta)\T X + \lambda*\beta\T = 0
+&&\\
+&\equivalent X\T y - X\T X\beta + \lambda*\beta = 0 
+\equivalent X\T y = X\T X\beta - \lambda*\beta = (X\T X - \lambda I) \beta
+&&\\
+&\equivalent (X\T X - \lambda I)^{-1}X\T y = \beta
 &&} 
+The loss function can be viewed as a combination of penalties for $\beta$.
+The first penalty is the mean squared error from before which is large when $\beta$ cannot reproduce the samples $y$ well.
+The second penalty is for $\beta$ of long lenghts, which means that short $\beta$ with small coefficients are preferred.
+These two penalties probably are contradictive and the optimal beta will be somewhere in between the individual minima for the penalties.
+This kind of regularization using $\lambda\beta\T\beta$ is called $L^2$ regularization and is a special case of the general Tikhonov regularization formulating constraints on $\beta$ as $\beta\T\Gamma\T\Gamma\beta$ with Tikhonov matrix $\Gamma$.
+$\Gamma\beta$ is a linear mapping which can be chosen to also formulate special constraints on beta like differences between coefficients, or projections to different vector spaces.
+
+When we use only half of the data set for \enquote{training} our model i.e.
+solving for $\beta$ we can check how well it reproduces the other half of the data set, which indicates how well the model generalizes.
+While the model may be very accurate for the data used for training, it may be inaccurate for the data used for \enquote{validation} (the other half), which means that the model is \enquote{over fitted} to the training data and does not approximate the actually underlying function well.
+In our case we can use regularization to allow the model to have a less perfect fit to the data when it instead fulfills the other objective of being small.
+The degree to which the model can violate the mean squared error depends on $\lambda$.
+
+\subsection{Cross Validation}
+In order to find the perfect regularization parameter $\lambda$ in order for our model to generalize well we need to try some out.
+This can be done using cross validation.
+In cross validation, we divide our data set into $K$ euqually sized parts (randomly), then we train our model using K-1 parts of the data and validate the model by computing the mean suared error using the missing part.
+This is done k times so that each part has been used for validation once and the errors are averaged.
+The following pseudo code should make the algorithm clear.
+\begin{algorithmic}[1]
+\State Split data $D$ into $K$ equally sized parts $\{D_1,\dots,D_K\}$ randomly
+\For{$k=1\dots K$}
+\State train on data $D_{\backslash D_k}$ yielding $\beta_i$
+\State calculate MSE using $D_k$ ~ $err_k = L^{MSE}(\beta_k,D_k)$
+\EndFor
+\State calculate total MSE as $err=\frac{1}{K}\sum_{k=1}^K \frac{err_k}{|D_k|}$ 
+\end{algorithmic}
+The total error will be a measure for generalization of the model, using this measure an optimal $\lambda$ can be found by trying different $\lambda$ and adaptively tightening the search space.
+For the start it may be useful to try lambdas of different orders of magnitude first.
 

### 2018-07-29 mlr_summary.tex
supportvectormachine done, now logistic regression

@@
 \newcommand{\argmin}{\mathop{\mathrm{argmin}}}
+\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
 \newcommand{\dom}[1]{\mathbb{#1}}
@@
 \newcommand{\equivalent}{\Leftrightarrow}
+\newcommand{\mathtext}[1]{\quad\text{#1}\quad}
+\newcommand{\with}{\mathtext{with}}
 % make star '*' become \cdot (normal multiplication sign)
@@
 &&\\&
-f_\beta(x) = x\T\beta \quad\text{with}\quad\beta\in\dom{R}^m
+f_\beta(x) = x\T\beta \with\beta\in\dom{R}^m
 &&}
@@
 \eqnsnn{
-\phi(x) \in \dom{R}^l \quad\text{with}\quad l\in\dom{N}
+\phi(x) \in \dom{R}^l \with l\in\dom{N}
 &&}
@@
 \eqnsnn{&
-\phi(x) = \mat{1 & \phi_1(x) & \cdots & \phi_N(x)}\T \quad\text{with}\quad \phi_i(x) = e^{-\frac{1}{2}||x_i-x||_2^2}
+\phi(x) = \mat{1 & \phi_1(x) & \cdots & \phi_N(x)}\T \with \phi_i(x) = e^{-\frac{1}{2}||x_i-x||_2^2}
 &&}
@@
 
+\section{Support Vector Machine}
+Imagine a binary classification problem where the data set consists of multivariate inputs $x_i \in \dom{R}^M$ and discrete outputs $y_i \in \{-1,1\}$ ($i = \{1,\dots,N\}$) which means that a point $x$ may either belong to class $1$ or $-1$.
+For this kind of problem we want to find a function that can separate the points of one class from another.
+In the 2 dimensional case, a simple separator would be a straight line and a point can either be on one or the other side of it.
+For higher dimensional $x$ the separator would be a hyperplane.
+
+To get used to the plane stuff again, lets review the point-normal form of a plane.
+In the simplest case where the plane intersects the origin, we can check if a point is in the plane by projecting its position vector $x$ onto the plane normal $n$.
+When $x$ is orthogonal to $n$ the projection $x\T n = 0$ and $x$ is in the nullspace of $n$ that is the plane.
+When the plane does not intersect the origin we first need to translate the coordinatesystem so that it does.
+We can do this by subtarcting a support vector of the plane $p_0$ (a position vector of a point that is part of the plane) from $x$, resulting in the check $(x-p_0)\T n = 0$ which can be rewritten as follows.
+\eqnsnn{&
+(x-p_0)\T n = 0 \equivalent x\T n + d = 0 \quad\text{with constant}\quad d = -p_0\T n
+&&}
+If $n$ is of unit length, then the distance of $x$ to the plane is given by the projection onto the normal $dist=x\T n + d$.
+When not normalized then the distance is given by the following.
+\eqnsnn{&
+dist = (x-p_0)\T \frac{n}{|n|} = \frac{x\T n}{|n|} - \frac{p_0\T n}{|n|} = \frac{x\T n + d}{|n|}
+&&}
+This means that the projection on the normal is the distance to the plane scaled by the normals length $x\T n + d = dist*|n|$.
+For a fixed projection value, this means, the smaller the normal the greater the distance.
+In fact for a projection of value $1$ the distance is $\frac{1}{|n|}$.
+We want the margin, that is the \enquote{width} of the separator, to be maximal.
+If we require projections to be at least $1$ (or $-1$ depending on the side) the margin is $\frac{1}{2*|n|}$.
+
+Lets formulate the separation objective as a constrained optimization problem with an inequality constraint, from which we can set up a Lagrangian.
+\eqnsnn{&
+\hat{n} = \argmin_{n} \frac{1}{2}|n|^2 \quad\text{subject to}\quad y_i(x_i\T n + d) \geq 1
+&&\\&
+\mathcal{L}(n,d,\lambda) = \frac{1}{2}\sum_{j=1}^M (n_j)^2 - \sum_{i=1}^N \lambda_i *(y_i(x_i\T n + d)-1)
+&&}
+When the data does not allow for a linear separation of the datapoints, then this problem cannot be solved because there is no feasible region.
+But we can modify the problem in order to allow points to be on the wrong side of the plane at a cost.
+For this a new variable $\xi_i\geq 0$ is introduced which measures by how much the inequality constrained has been violated and sanitizes the constraint so that it becomes feasible.
+Of course, we want the amount of violations to be minimal, so we introduce a penalty on large $\xi$ which we control by a weighting parameter $C$.
+\eqnsnn{&
+\hat{n} = \argmin_{n} \frac{1}{2}|n|^2 + C*\sum_{i=1}^N\xi_i \quad\text{subject to}\quad y_i(x_i\T n + d) \geq 1-\xi_i ~\text{and}~ \xi_i \geq 0 
+&&\\&
+\mathcal{L}(n,d,\lambda,\mu) = \frac{1}{2}\sum_{j=1}^M (n_j)^2  + C*\sum_{i=1}^N \xi_i - \sum_{i=1}^N \lambda_i *(y_i(x_i\T n + d)-1+\xi_i) - \sum_{i=1}^n \mu_i*\xi_i
+&&}
+When solving the dual problem (which is always convex), we get the optimal normal and classifier as
+\eqnsnn{&
+\hat{n} = \sum_{i=1}^N \hat{\lambda}_iy_ix_i
+&&\\&
+f(x) = \text{sign}(x\T \hat{n} + d)
+&&\\&
+\hat{\lambda} = \argmax_\lambda \sum_{i=1}^N\lambda_i - \frac{1}{2}\sum_{i=1}^N \sum_{k=1}^N \lambda_i \lambda_k y_i y_k x_i\T x_k
+\quad\text{subject to}\quad \sum_{i=1}^N \lambda_i y_i = 0 ~\text{and}~ C \geq \lambda_i \geq 0
+&&}
+Since $\lambda_i$ are the lagrangian multipliers of the inequality constraint they are either inactive ($=0$) if the corresponding point is correctly classified and outside the margin, or active ($\neq 0$) if the corresponding point violates the constraint.
+So all active $x_i$ contribute to the solution and are called the support vectors.
+
+\section{Logistic Regression}
+We already discussed one classifier, the support vector machine, that is a binary classifier.
+In this section we will take a look at a different approach to classification.
+As usual we have multivariate $x_i \in \dom{R}^M$ as inputs which are labeled with a discrete variable $y_i \in \dom{Y}=\{Y_1,\dots,Y_l\}$ ($i = {1,\dots,N}$).
+This time we want to know the probability of a class $y$ given that we observed some $x$ which is denoted by $P(y|x)$.
+
+To estimate the probabilities we make use of a discriminative function $f(x,y):\dom{R}^M \times \dom{Y} \to \dom{R}$ that we have to learn. The probabilities are then given as:
+\eqnsnn{&
+P(y|x) = \frac{e^{f(x,y)}}{Z} \mathtext{with normalization} Z = \sum_{y\in\dom{Y}} e^{f(x,y)}
+&&}
+The discriminative function has to have a high value when x is likely to be of class y, and a low value when it's not.
+This way of modeling probabilities is related to the Boltzmann distribution $p(x) = e^{-E(x)}$ where $E(x)$ is an energy and x has a high probability if its energy is low.
+
+Now that we set up the basic framework, we need to define the discriminative function.
+Lets formulate its as a linear function in features (similar to linear regression).
+\eqnsnn{&
+f(x,y) = \phi(x,y)\T\beta \mathtext{with for example} \phi(x,y) = \mat{\phi(x)[y=Y_1] \\ \phi(x)[y=Y_2] \\ \vdots \\ \phi(x)[y=Y_l]}\T
+&&}
+The $[cond]$ operator is $1$ if the condition is true, otherwise $0$. 
+The objective of our model is to maximize the likelihood $p(y_i|x_i)$ for each $(x_i,y_i)$, but instead it could also maximize the log likelihood $\ln(p(y_i|x_i))$ since the logarithm is a monotonic increasing function.
+Also, in machine learning we rather like to minimize stuff, so lets minimize the neg-log-likelihood by minimizing the following loss.
+\eqnsnn{&
+L(\beta) = \sum_{i=1}^N - \ln(\;p(y_i|x_i)\;) = - \sum_{i=1}^N\Big( f(x_i,y_i) - \sum_{y \in \dom{Y}} f(x_i,y) \Big)
+&&}
+
 \end{document}

### 2018-07-30 mlr_summary.tex
logistic regression derivative

@@
 \eqnsnn{&
-f(x,y) = \phi(x,y)\T\beta \mathtext{with for example} \phi(x,y) = \mat{\phi(x)[y=Y_1] \\ \phi(x)[y=Y_2] \\ \vdots \\ \phi(x)[y=Y_l]}\T
+f(x,y) = \phi(x,y)\T\beta \with \phi(x,y) = \mat{\phi(x)[y=Y_1] \\ \phi(x)[y=Y_2] \\ \vdots \\ \phi(x)[y=Y_l]}
 &&}
@@
 Also, in machine learning we rather like to minimize stuff, so lets minimize the neg-log-likelihood by minimizing the following loss.
-\eqnsnn{&
-L(\beta) = \sum_{i=1}^N - \ln(\;p(y_i|x_i)\;) = - \sum_{i=1}^N\Big( f(x_i,y_i) - \sum_{y \in \dom{Y}} f(x_i,y) \Big)
+\eqnsnn{
+L(\beta) &= -\sum_{i=1}^N \ln(\;p(y_i|x_i)\;) 
+= - \sum_{i=1}^N\Big( \ln\big( e^{f(x_i,y_i)}\big) - \ln\big(\sum_{y' \in \dom{Y}} e^{f(x_i,y')}\big) \Big)
+&&\\
+&= - \sum_{i=1}^N\Big( \ln\big( e^{f(x_i,y_i)}\big)\Big) +\sum_{i=1}^N\Big( \ln\big(\sum_{y' \in \dom{Y}} e^{f(x_i,y')}\big) \Big) = -A(\beta) + B(\beta)
+&&
+}
+Lets assume for simplicity that $\phi(x)$ is scalar, so that there is one $\beta_c$ corresponding to $\phi(x,y)_c$ with $c \in \{1,\dots,l\}$ indicating the class $Y_c$. 
+\eqnsnn{
+\deriv{L(\beta)}{\beta_c} &= -\deriv{A(\beta)}{\beta_c} + \deriv{B(\beta)}{\beta_c}
+&&\\
+\deriv{A(\beta)}{\beta_c} &= \sum_{i=1}^N \deriv{f(x_i,y_i)}{\beta_c} = \sum_{i=1}^N \phi(x_i,y_i)_c\T
+&&\\
+\deriv{B(\beta)}{\beta_c} &= \sum_{i=1}^N \deriv{}{\beta_c} \ln\big(\sum_{y' \in \dom{Y}} e^{f(x_i,y')}\big) 
+&&\\
+&= \sum_{i=1}^N \frac{1}{\sum_{y' \in \dom{Y}} e^{f(x_i,y')}} * \deriv{}{\beta_c} \sum_{y' \in \dom{Y}} e^{f(x_i,y')}
+&&\\
+&=\sum_{i=1}^N \frac{1}{\sum_{y' \in \dom{Y}} e^{f(x_i,y')}} * \sum_{y' \in \dom{Y}} e^{f(x_i,y')}*\phi(x_i,y')_c\T
+&&\\
+&=\sum_{i=1}^N \frac{e^{f(x_i,Y_c)}}{\sum_{y' \in \dom{Y}} e^{f(x_i,y')}} * \phi(x_i,Y_c)_c\T
+&&\\
+&=\sum_{i=1}^N p(Y_c|x_i) * \phi(x_i,Y_c)_c\T
+&&\\
+\deriv{L(\beta)}{\beta_c} &= \sum_{i=1}^N p(Y_c|x_i) * \phi(x_i,Y_c)_c\T - \phi(x_i,y_i)_c\T
 &&}

### 2018-07-31 mlr_summary.tex
newton method

@@
 
-To estimate the probabilities we make use of a discriminative function $f(x,y):\dom{R}^M \times \dom{Y} \to \dom{R}$ that we have to learn. The probabilities are then given as:
+To estimate the probabilities we make use of a discriminative function $f(x,y):\dom{R}^M \times \dom{Y} \to \dom{R}$ that we have to learn.
+The discriminative function has to have a high value when $x$ is likely to be of class $y$, and a low value when it's not so that $y(x) = \argmax_y f(x,y)$ where $y(x)$ is the correct class for $x$.
+
+Similar to the Boltzmann distribution $p(x) = e^{-E(x)}$ where $E(x)$ is an energy (or error), the probabilities are calculated as:
 \eqnsnn{&
-P(y|x) = \frac{e^{f(x,y)}}{Z} \mathtext{with normalization} Z = \sum_{y\in\dom{Y}} e^{f(x,y)}
+P(y|x) = \frac{e^{f(x,y)}}{Z} \mathtext{with normalization} Z = \sum_{y'\in\dom{Y}} e^{f(x,y')}
 &&}
-The discriminative function has to have a high value when x is likely to be of class y, and a low value when it's not.
-This way of modeling probabilities is related to the Boltzmann distribution $p(x) = e^{-E(x)}$ where $E(x)$ is an energy and x has a high probability if its energy is low.
 
 Now that we set up the basic framework, we need to define the discriminative function.
-Lets formulate its as a linear function in features (similar to linear regression).
+Lets formulate it as a linear function in features (similar to linear regression).
 \eqnsnn{&
@@
 \eqnsnn{
-\deriv{L(\beta)}{\beta_c} &= -\deriv{A(\beta)}{\beta_c} + \deriv{B(\beta)}{\beta_c}
+\deriv{L(\beta)}{\beta} &= -\deriv{A(\beta)}{\beta} + \deriv{B(\beta)}{\beta}
 &&\\
-\deriv{A(\beta)}{\beta_c} &= \sum_{i=1}^N \deriv{f(x_i,y_i)}{\beta_c} = \sum_{i=1}^N \phi(x_i,y_i)_c\T
+\deriv{A(\beta)}{\beta} &= \sum_{i=1}^N \deriv{f(x_i,y_i)}{\beta} = \sum_{i=1}^N \phi(x_i,y_i)\T
 &&\\
-\deriv{B(\beta)}{\beta_c} &= \sum_{i=1}^N \deriv{}{\beta_c} \ln\big(\sum_{y' \in \dom{Y}} e^{f(x_i,y')}\big) 
+\deriv{B(\beta)}{\beta} &= \sum_{i=1}^N \deriv{}{\beta} \ln\big(\sum_{y' \in \dom{Y}} e^{f(x_i,y')}\big) 
 &&\\
-&= \sum_{i=1}^N \frac{1}{\sum_{y' \in \dom{Y}} e^{f(x_i,y')}} * \deriv{}{\beta_c} \sum_{y' \in \dom{Y}} e^{f(x_i,y')}
+&= \sum_{i=1}^N \frac{1}{\sum_{y' \in \dom{Y}} e^{f(x_i,y')}} * \deriv{}{\beta} \sum_{y' \in \dom{Y}} e^{f(x_i,y')}
 &&\\
-&=\sum_{i=1}^N \frac{1}{\sum_{y' \in \dom{Y}} e^{f(x_i,y')}} * \sum_{y' \in \dom{Y}} e^{f(x_i,y')}*\phi(x_i,y')_c\T
+&=\sum_{i=1}^N \frac{1}{Z} * \sum_{y' \in \dom{Y}} e^{f(x_i,y')}*\phi(x_i,y')\T
 &&\\
-&=\sum_{i=1}^N \frac{e^{f(x_i,Y_c)}}{\sum_{y' \in \dom{Y}} e^{f(x_i,y')}} * \phi(x_i,Y_c)_c\T
+&=\sum_{i=1}^N \sum_{y' \in \dom{Y}} 
+\frac{e^{f(x_i,y')}}{Z} * \phi(x_i,y')\T
 &&\\
-&=\sum_{i=1}^N p(Y_c|x_i) * \phi(x_i,Y_c)_c\T
+&=\sum_{i=1}^N \sum_{y' \in \dom{Y}} p(y'|x_i) * \phi(x_i,y')\T
 &&\\
-\deriv{L(\beta)}{\beta_c} &= \sum_{i=1}^N p(Y_c|x_i) * \phi(x_i,Y_c)_c\T - \phi(x_i,y_i)_c\T
+\Big(\deriv{L(\beta)}{\beta}\Big)\T = \nabla_\beta L(\beta) &= \sum_{i=1}^N \Big( \sum_{y' \in \dom{Y}} p(y'|x_i) * \phi(x_i,y') \Big) - \phi(x_i,y_i)
 &&}
 
+As can be seen, the gradient of the loss is not linear in $\beta$ because $\beta$ sits inside of the non-linear calculation of $p(y'|x_i)$.
+So unfortunately we cannot analytically derive the optimal $\beta$, but instead we can use the Newton algorithm to iteratively get there.
+
+\subsection{Newton Algorithm}
+The newton algortihm can be derived from a taylor expansion of a function.
+Lets review the taylor expanson first, by writing down a univariate $(n-1)$ order taylor of $f$ around $x$ with step size $h$.
+\eqnsnn{&
+f(x+h) \approx f(x) + \frac{f'(x)*h}{1!} + \frac{f''(x)*h^2}{2!} + \frac{f^3(x)*h^3}{3!} + \dots + \mathcal{O}(h^n)
+&&}
+From this, the forward difference can be derived (which is not what we want, but shows how neat Taylor is).
+\eqnsnn{&
+f(x+h) \approx f(x) + f'(x)*h + \mathcal{O}(h^2) \;\equivalent\; \frac{f(x+h)-f(x)}{h} \approx f'(x) + \mathcal{O}(h^2)
+&&}
+
+Lets do something else, lets set a first order Taylor to zero (root finding):
+\eqnsnn{
+f(x_{n+1}) = f(x_n+h) &\approx f(x_n) + f'(x_n)*h + \mathcal{O}(h^2)
+&&\\
+0 &\approx f(x_n) + f'(x_n)*h
+&&\\
+0 &\approx \frac{f(x_n)}{f'(x_n)} + h
+&&\\
+h &\approx - \frac{f(x_n)}{f'(x_n)}
+&&}
+Here we derived the stepsize $h$ of a newton step, which we can use to update the position $x_n$ in order to get $x_{n+1} = x_n+h$. 
+When iterating this newton step, we approach a root of $f(x)$.
+
+In our case, we want to find a root of the gradient $\nabla L$, lets again formulate a Taylor expansion, this time multivariate second order with step $\delta$ and Hessian $\mathcal{H}_L$.
+\eqnsnn{
+L(\beta_{n+1}) = L(\beta_n + \delta) &\approx L(\beta_n) + \nabla L(\beta_n)\T \delta + \frac1 2 \delta\T \mathcal{H}_L(\beta_n)\delta + \mathcal{O}(|\delta|^3)
+&&}
+The derivative of this taylor with respect to the step $\delta$ yields an approximation of the derivative of $L$ around $\beta_n$. To find a root of it, we set it to zero.
+\eqnsnn{
+0 = \deriv{L(\beta_n+\delta)}{\delta} &\approx \nabla L(\beta_n)\T + \delta\T \mathcal{H}_L(\beta_n)
+&&\\
+0 &\approx \nabla L(\beta_n)\T[\mathcal{H}_L(\beta_n)]^{-1} + \delta\T \mathcal{H}_L(\beta_n)[\mathcal{H}_L(\beta_n)]^{-1} = \nabla L(\beta_n)\T[\mathcal{H}_L(\beta_n)]^{-1} + \delta\T
+&&\\
+\delta &\approx -[\mathcal{H}_L(\beta_n)]^{-1}\nabla L(\beta_n)
+&&}
+This is the newton step we can use to iteratively work our way to the optimal $\beta$ using 
+\\$\beta_{n+1} = \beta_n+\delta$.
+What is missing in order to optimize $\beta$ for logistic regression is the Hessian of the loss $\mathcal{H}_L(\beta) = \deriv{}{\beta}\nabla L(\beta)$, which is left to the reader to derive.
+
 \end{document}

### 2018-07-31 mlr_summary.tex
NN and PCA

@@
 
+\tableofcontents
+
 \section{Machine Learning Briefly}
@@
 \eqnsnn{
-\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ (y-X\beta)\T X + (y-X\beta)\T X = 2*(y-X\beta)\T X = 0
+\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ -(y-X\beta)\T X - (y-X\beta)\T X = -2*(y-X\beta)\T X = 0
 &&\\&
@@
 &&\\
-\deriv{L(\beta)}{\beta} = 0 &\equivalent 2(y-X\beta)\T X + 2\lambda*\beta\T = 0
-\equivalent (y-X\beta)\T X + \lambda*\beta\T = 0
+\deriv{L(\beta)}{\beta} = 0 &\equivalent -2(y-X\beta)\T X + 2\lambda*\beta\T = 0
 &&\\
-&\equivalent X\T y - X\T X\beta + \lambda*\beta = 0 
-\equivalent X\T y = X\T X\beta - \lambda*\beta = (X\T X - \lambda I) \beta
+&\equivalent \lambda*\beta\T - (y-X\beta)\T X = 0
+&&\\
+&\equivalent \lambda*\beta - X\T y + X\T X\beta = 0 
+&&\\
+&\equivalent X\T y = X\T X\beta - \lambda*\beta = (X\T X - \lambda I) \beta
 &&\\
@@
 As usual we have multivariate $x_i \in \dom{R}^M$ as inputs which are labeled with a discrete variable $y_i \in \dom{Y}=\{Y_1,\dots,Y_l\}$ ($i = {1,\dots,N}$).
-This time we want to know the probability of a class $y$ given that we observed some $x$ which is denoted by $P(y|x)$.
+This time we want to know the probability of a class $y$ given that we observed some $x$ which is denoted by $p(y|x)$.
 
@@
 \eqnsnn{&
-P(y|x) = \frac{e^{f(x,y)}}{Z} \mathtext{with normalization} Z = \sum_{y'\in\dom{Y}} e^{f(x,y')}
+p(y|x) = \frac{e^{f(x,y)}}{Z} \mathtext{with normalization} Z = \sum_{y'\in\dom{Y}} e^{f(x,y')}
 &&}
@@
 The $[cond]$ operator is $1$ if the condition is true, otherwise $0$.
+
 The objective of our model is to maximize the likelihood $p(y_i|x_i)$ for each $(x_i,y_i)$, but instead it could also maximize the log likelihood $\ln(p(y_i|x_i))$ since the logarithm is a monotonic increasing function.
@@
 }
-Lets assume for simplicity that $\phi(x)$ is scalar, so that there is one $\beta_c$ corresponding to $\phi(x,y)_c$ with $c \in \{1,\dots,l\}$ indicating the class $Y_c$. 
+The derivative of the loss is then calculated from the two separated sums $A(\beta)$ and $B(\beta)$:
 \eqnsnn{
@@
 &&\\
-\Big(\deriv{L(\beta)}{\beta}\Big)\T = \nabla_\beta L(\beta) &= \sum_{i=1}^N \Big( \sum_{y' \in \dom{Y}} p(y'|x_i) * \phi(x_i,y') \Big) - \phi(x_i,y_i)
+\Big(\deriv{L(\beta)}{\beta}\Big)\T &= \nabla_\beta L(\beta) = \sum_{i=1}^N \Big( \sum_{y' \in \dom{Y}} p(y'|x_i) * \phi(x_i,y') \Big) - \phi(x_i,y_i)
 &&}
@@
 
+\section{Neural Networks}
+Neural networks currently are the very hype in many fields of inteligent systems, probably because they are so flexible.
+But lets not get carried away and start by defining a regression problem.
+Given data $D=\{x_i,y_i\}_{i=1}^N$ with multivariate $x_i \in \dom{R}^M$ and scalar $y\in\dom{R}$ we want to estimate again as in linear regression $y=f(x)$.
+For this, we model the function $f$ as follows:
+\eqnsnn{&
+f(x) = z_{L+1} = W_L*x_L \with x_l = \sigma(z_l) \;\;l\in\{2,\dots,L\} \mathtext{,} \sigma(z) = \frac{1}{1+e^{-z}} \mathtext{and} x_1 = x
+&&}
+In this equation we have $l$ matrices $W_1 \dots W_l$ with dimensions $W_1^{d(2)\times M}\; W_2^{d(3)\times d(2)}\dots W_l^{1\times d(l)}$. 
+The sigmoid function $\sigma$ takes the role of the \enquote{activation function} which is an analogy to neurons in a brain which can propagate an action potential based on the magnitude of the signal.
+
+Lets get more intuition on this from a simple example.
+Suppose $x \in \dom{R}^3$ and $y \in \dom{R}^1$ and we have three matrices $W_1\;W_2\;W_3$.
+To be able to input our $x$ into the neural net we have to choose the first dimension of $W_1$ to be $3$ and the last dimension of $W_3$ to be $1$ in order to map to $\dom{R}^1$.
+The other dimensions are left to be choosen freely.
+Lets choose $W_2^{10\times 10}$ which implies $W_1^{10 \times 3}$ and $W_3^{1 \times 10}$.
+Our function is then:
+\eqnsnn{&
+f(x) = W_3\;\sigma\Big(W_2\;\sigma(W_1x)\Big)
+&&}
+This algorithm is called the forward propagation of $x$ through the neural network.
+The unknowns in this equation are the entries of the matrices, the so called weights.
+In order to optimize the weights we need to define a cost function $c(f,y)$ (a loss) for the prediction $f$.
+In the literature, we can find many different cost functions that can be used, we'll stick to the well known mean squared error.
+\eqnsnn{&
+c(f,y) = (f-y)^2 \mathtext{so that} C_\text{MSE} = \frac1n \sum_{i=1}^N c(f(x_i),y_i)
+&&\\&
+\deriv{\;c(f,y)}{f} = 2(f-y)\T \mathtext{if $f$ and $y$ were vectors, in our case we can ommit the transpose}
+&&}
+Lets call $\deriv{\;c(f,y)}{f} = \delta_{L+1}$ as it is the derivative of the loss with repect to $z_{L+1}$ (in our case $\delta_4$ the derivative of $f(x) = z_4=W_3\;x_3$).
+Due to the chain rule the derivatives for other layers are recursively calculated like this:
+\eqnsnn{&
+\delta_l = \deriv{c}{z_l} = \deriv{c}{z_{l+1}} * \deriv{z_{l+1}}{x_l}*\deriv{x_l}{z_l}
+&&}
+The derivative of the of the loss with respect to a specific weight of a matrix $W_{l,ij}$ is given by:
+\eqnsnn{&
+\deriv{c}{W_{l,ij}} = \deriv{c}{z_{l+1,i}}*\deriv{z_{l+1,i}}{W_{l,ij}}
+&&}
+
+To optimize the weights of the matrices, we feed the neural net all of our data samples $x_i$ and collect the losses $c(f(x_i),y_i)$.
+We then backpropagate the losses through the net, which means we collect the derivatives of the losss for all weights and sum corresponding derivatives up forming a total gradient for a specific weight.
+The weights are then updated using a gradient descend step which reads $W_{l,ij} = W_{l,ij} - \alpha \deriv{c}{W_{l,ij}}$ with learning rate $\alpha$ (probably $\alpha=0.1$).
+And then we iterate forward and backward propagation.
+
+\section{PCA}
+Principle Component Analysis (PCA) is a method for dimensionality reduction.
+This may be useful when you have very high dimensional data, but want to reduce the amount of data you want to process in order to learn a function.
+In this case you may want to get a few of the \enquote{most important} parts of your highdimensional data point, or lets say a low dimensional datapoint which encodes the most important aspects of yout high dimensional point.
+
+A way to do that, is to find a projection of our data $x_i \in \dom R^M$ onto $z_i \in \dom R^P$ where $P \leq M$.
+Lets look at the projection from the other side, lets say we transform $z$ to $x$ using the following, where $V_P$ is the projection matrix and $\mu$ is a vector accounting for a translation in $\dom R^M$.
+\eqnsnn{&
+x \approx V_P z + \mu
+&&\\&
+err = \sum_{i=1}^N || V_P z_i + \mu - x_i ||^2
+&&}
+The error $err$ measures the mean squared error of the transformation of $z$ to $x$.
+Using this, we can derive the optimal transformation of $x$ to $z$ when minimizing with respect to $z$.
+\eqnsnn{
+\hat{z}_{1:N} &= \argmin_{z_{1:N}} \sum_{i=1}^N || V_P z_i + \mu - x_i ||^2
+&&\\
+0 &= \deriv{}{z_i}\sum_{i=1}^N || V_P z_i + \mu - x_i ||^2 = \deriv{}{z_i} (V_P z_i + \mu - x_i)\T(V_P z_i + \mu - x_i)
+&&\\
+ &= 2*(V_P z_i + \mu - x_i)\T V_P
+&&\\
+\equivalent 0 &= V_P\T V_P z_i + V_P\T(\mu-x_i)
+&&}
+If we constrain $V_p$ to be orthonormal (composed of orthonormal vectors), then $V_P\T V_P = I$ and this simplifies to the follownig:
+\eqnsnn{&
+z_i = V_P\T(x_i-\mu)
+&&}
+Doing the same for $\mu$ yields the following:
+\eqnsnn{
+\hat{\mu} &= \argmin_{\mu} \sum_{i=1}^N || V_P z_i + \mu - x_i ||^2
+&&\\
+0 &= \deriv{}{\mu} \sum_{i=1}^N (V_P z_i + \mu - x_i)\T(V_P z_i + \mu - x_i)
+ = 2* \sum_{i=1}^N (V_P z_i + \mu - x_i)\T 
+&&\\
+\equivalent 0 &= N*\mu*\sum_{i=1}^N (V_P z_i - x_i)
+&&\\
+\equivalent \mu &= \frac1N\sum_{i=1}^N V_Pz_i - \frac1N\sum_{i=1}^N x_i
+&&}
+So $\mu$ is the difference of the expeted value of the projected $z_i$ and the expected value $x_i$.
+Lets assume that $E[z] = 0$ because that would be nice, and we don't care about the expected value of $z$ as it is unrelated to what we try to achieve. 
+Due to the linearity of the expected value the projection of $z$ also has $0$ as expected value.
+\eqnsnn{&
+\mu = \frac1N\sum_{i=1}^N x_i = E[x]
+&&}
+
+So the projection $z_i = V_P\T(x_i-\mu)$ is the projecting the \enquote{centered} $x_i$.
+For the derivation of $V_P$ lets assume that $x_i$ are already centered ($E(x)=0$).
+What we want to minimize is the projection error, which we can estimate by back projecting the projection of $x_i$ and comparing it with the initial $x_i$.
+\eqnsnn {&
+\hat{V}_P = \argmin_{V_P} \sum_{i=1}^N ||x_i- V_PV_P\T x_i||^2 
+&&}
+
 \end{document}

### 2018-07-31 mlr_summary.tex
PCA finished

@@
 This may be useful when you have very high dimensional data, but want to reduce the amount of data you want to process in order to learn a function.
-In this case you may want to get a few of the \enquote{most important} parts of your highdimensional data point, or lets say a low dimensional datapoint which encodes the most important aspects of yout high dimensional point.
+In this case you may want to get a few of the \enquote{most important} parts of your highdimensional data point, or lets say a low dimensional datapoint which encodes the most important aspects of your high dimensional point.
 
@@
 &&}
+This time we will derive the optimum by observing the properties of the equation instead of root finding in the derivative.
+Lets reformulate the loss a bit.
+\eqnsnn {
+L(V_P) &= \sum_{i=1}^N (x_i- V_PV_P\T x_i)\T(x_i- V_PV_P\T x_i)
+&&\\
+&= \sum_{i=1}^N x_i\T x_i - 2 x_i\T V_P V_P\T x_i 
++ x_i\T V_P V_P\T V_P V_P\T x_i
+&&\\
+& = \sum_{i=1}^N x_i\T x_i - 2x_i\T V_P V_P\T x_i  + x_i\T V_P V_P\T x_i
+&&\\
+&= \sum_{i=1}^N x_i\T x_i - x_i\T V_P V_P\T x_i
+&&\\
+&= C - \sum_{i=1}^N x_i\T V_P V_P\T x_i \with C = \sum_{i=1}^N x_i\T x_i
+&&\\
+&= C - \sum_{i=1}^N || V_P\T x_i ||^2 \mathtext{($V_P\T$ composed of row vectors $v_j\T$)}
+&&\\
+&= C - \sum_{i=1}^N \sum_{j=1}^P (v_j\T x_i)*(v_j\T x_i)
+ \;=\; C - \sum_{j=1}^P \sum_{i=1}^N  (x_i\T v_j)*(x_i\T v_j)
+&&\\
+&= C - \sum_{j=1}^P ||X v_j||^2 \mathtext{($X$ composed of row vectors $x_i\T$)}
+&&\\
+&= C - \sum_{j=1}^P v_j\T X\T X v_j
+&&}
+So whatwe can see from this, is that the loss will be minimal if we maximize $\sum_{j=1}^P v_j\T X\T X v_j$.
+We could again find the root of its derivative but, it will be more fun with some clever factorization of $X\T X$, which is the Eigendecomposition. 
+Since $X\T X$ is a symmetric matrix, its Eigendecomposition is
+\eqnsnn{&
+X\T X = Q\Lambda Q\T \mathtext{with orthonormal} Q \mathtext{and diagonal} \Lambda
+&&\\
+& \argmax_{v_{1:P}} \sum_{j=1}^P v_j\T Q\Lambda Q\T v_j
+&&} 
+Suppose the eigenvlus $\lambda_1,\dots,\lambda_M$ are sorted in value descending order on the diagonal of $\Lambda$.
+To maximize the the above expression, we need to choose $v_{1:P}$ to be the first $P$ eigenvectors $q_i$ corresponding to the $P$ largest eigenvalues. 
+If we could, we would choose all $v_j$ to be the first eigenvector, as this would truly yield the maximum, but due to our constraint of $v_j$s being orthonormal to each other we can't do that.
+
+One last thing to be mentioned before moving on is, that instead of using the Eigendecomposition of $X\T X$ to get the optimal $V_P$ we can as well use the singular value decomposition (SVD) of $X$ to get there:
+\eqnsnn{&
+X = USW\T \mathtext{with orthonormal} U,\;W \mathtext{and diagonal} S
+&&\\&
+X\T X = (USW\T)\T (USW\T) = W S\T U\T USW\T = WS^2W\T = Q\Lambda Q\T \implies W = Q \quad S^2=\Lambda
+&&}
 

### 2018-08-01 mlr_summary.tex
sum fixes

@@
 \eqnsnn{
-\deriv{L(\beta)}{\beta} = 0 ~&\equivalent~ -(y-X\beta)\T X - (y-X\beta)\T X = -2*(y-X\beta)\T X = 0
+\deriv{L(\beta)}{\beta} = 0 ~&\equivalent -(y-X\beta)\T X - (y-X\beta)\T X = -2*(y-X\beta)\T X = 0
 &&\\&
@@
 Maybe you may have noticed that the linear regression algorithm may fail in case that $(X\T X)$ is not invertible.
-To ensure invertibility a tiny quantity can be added to or subtracted from the diagonal of the matrix like so: $(X\T X - \lambda I)$.
+To ensure invertibility a tiny quantity can be added to the diagonal of the matrix like so: $(X\T X + \lambda I)$.
 Interestingly, the optimal solution to the following loss has this exact form.
@@
 &&\\
-&\equivalent X\T y = X\T X\beta - \lambda*\beta = (X\T X - \lambda I) \beta
+&\equivalent X\T y = X\T X\beta + \lambda*\beta = (X\T X + \lambda I) \beta
 &&\\
-&\equivalent (X\T X - \lambda I)^{-1}X\T y = \beta
+&\equivalent (X\T X + \lambda I)^{-1}X\T y = \beta
 &&} 
@@
 \subsection{Cross Validation}
-In order to find the perfect regularization parameter $\lambda$ in order for our model to generalize well we need to try some out.
+In order to find the perfect regularization parameter $\lambda$ for our model to generalize well, we need to try some $\lambda$s out.
 This can be done using cross validation.
@@
 We want the margin, that is the \enquote{width} of the separator, to be maximal.
-If we require projections to be at least $1$ (or $-1$ depending on the side) the margin is $\frac{1}{2*|n|}$.
+If we require projections to be at least $1$ (or $-1$ depending on the side) the margin is $\frac{2}{|n|}$.
 
@@
 &&}
-So whatwe can see from this, is that the loss will be minimal if we maximize $\sum_{j=1}^P v_j\T X\T X v_j$.
+So what we can see from this, is that the loss will be minimal if we maximize $\sum_{j=1}^P v_j\T X\T X v_j$.
 We could again find the root of its derivative but, it will be more fun with some clever factorization of $X\T X$, which is the Eigendecomposition. 
@@
 &&} 
-Suppose the eigenvlus $\lambda_1,\dots,\lambda_M$ are sorted in value descending order on the diagonal of $\Lambda$.
-To maximize the the above expression, we need to choose $v_{1:P}$ to be the first $P$ eigenvectors $q_i$ corresponding to the $P$ largest eigenvalues. 
+Suppose the eigenvalues $\lambda_1,\dots,\lambda_M$ are sorted in value descending order on the diagonal of $\Lambda$.
+To maximize the above expression, we need to choose $v_{1:P}$ to be the first $P$ eigenvectors $q_i$ corresponding to the $P$ largest eigenvalues. 
 If we could, we would choose all $v_j$ to be the first eigenvector, as this would truly yield the maximum, but due to our constraint of $v_j$s being orthonormal to each other we can't do that.

### 2018-08-01 mlr_summary.tex
kernel trick

@@
 
-So the projection $z_i = V_P\T(x_i-\mu)$ is the projecting the \enquote{centered} $x_i$.
+So the projection $z_i = V_P\T(x_i-\mu)$ is projecting the \enquote{centered} $x_i$.
 For the derivation of $V_P$ lets assume that $x_i$ are already centered ($E(x)=0$).
@@
 
+\section{The Kernel Trick}
+Remember when we introduced features to make things non-linear in linear regression.
+We use features to calculate new quantities from a given datapoint such as 
+\\$\phi(x) = \mat{1 & x & x^2 & e^x & e^{-x} & \sin(x) & \cos(x)}\T$.
+This is a very powerful way to allow for non-linearity but on the downside of it is the question:
+How to choose features so that some property of the original function or distribution becomes linear?
+This is not easy and sometimes coming up with the right features is impossible.
+
+The kernel trick is a way to get rid of explicitly defining features.
+Before diggin in, lets define what a kernel in this sense even is.
+A kernel is the dot product of the feature vectors of two inputs $x$ and $x'$.
+\eqnsnn{&
+k(x,x') = \phi(x)\T\phi(x')
+&&}
+This implies that a kernel is symmtric with respect to argument order $k(x,x') = k(x',x)$.
+Of course when we difine features, we imply a kernel function.
+If on the other hand we define a kernel function, we imply features.
+Due to the kernel being a dot product, it is a measure of similarity of the feature vectors and thus also a measure of similarity of $x$ and $x'$.
+If we keep this in mind, we can define kernels with respect to the distance of two points, for example a gaussian kernel:
+\eqnsnn{&
+k_\mathcal{N}(x,x') = \mathcal{N}(||x-x'||, \mu, \sigma^2) = \frac1{\sqrt{2\pi\sigma^2}}e^{\frac{(||x-x'||-\mu)^2}{2\sigma^2}}
+&&}
+
+The \enquote{trick} in the kernel trick is now to get rid of the features $\phi(x)$ by reformulating the problem to only include dot products $\phi(x)\T\phi(x')$, which are then replaced by kernels $k(x,x')$.
+Lets illustrate this process on various examples in the following.
+
+\subsection{Kernel PCA}
+In PCA we have following eigenvalue problem.
+\eqnsnn{&
+X\T X  v_j = \lambda v_j \mathtext{where $v_j$ is one of the eigenvectors used for the projection}
+&&\\
+&
+\lambda v_j = X\T X v_j = \sum_{i=1}^N \phi(x_i) \phi(x_i)\T v_j \mathtext{(sum of outer products of $\phi(x_i)$)}
+&&}
+From this we can extract a calculation for $v_j$ by replacing the dot product of $v_j$ and $\phi(x_i)$ by a new variable $b_{ij}$:
+\eqnsnn{&
+\sum_{i=1}^N \phi(x_i) \phi(x_i)\T v_j = \sum_{i=1}^N \phi(x_i) b_{ij} = \lambda v_j
+&&\\&
+\sum_{i=1}^N \phi(x_i) \alpha_{ij} = v_j \with \alpha_{ij} = \frac{b_{ij}}{\lambda}
+&&\\&
+\sum_{i=1}^N \phi(x_i) \alpha_{ij} = X\T\alpha_j = v_j
+&&} 
+When we substitue this into the initial eigenvalue problem we get the following we get to the following.
+\eqnsnn{
+X\T X  v_j = \lambda v_j \quad&=\quad X\T X X\T \alpha_j = \lambda X\T\alpha_j
+&&\\
+&\equivalent\quad XX\T XX\T \alpha_j = \lambda XX\T\alpha_j
+&&}
+This is pretty much what we have been waiting for. 
+We generated an expression using matrices of dot products of features, which we can replace by a kernel matrix \\$XX\T_{ik} = \phi(x_i)\T\phi(x_k) = k(x_i,x_k) = K_{ik}$.
+This brings us to a new eigenvalue problem which only contains kernel expressions.
+\eqnsnn{&
+K K \alpha_j = \lambda K\alpha_j \quad\equivalent\quad K\alpha_j = \lambda\alpha_j
+&&}
+
+We probably can get to this a little bit more straight forward since we know we want $XX\T$ in our equation:
+\eqnsnn{
+X\T X  v_j = \lambda v_j \quad\equivalent\quad X\T XX\T  v_j &= \lambda X\T v_j 
+&&\\ \quad\equivalent\quad XX\T XX\T  v_j &= \lambda XX\T v_j
+&&\\ \quad\equivalent\quad K v_j &= \lambda v_j
+&&}
+
 \end{document}

### 2018-08-05 mlr_summary.tex
so far so good, exam was written, no more need to further this

@@
 The \enquote{trick} in the kernel trick is now to get rid of the features $\phi(x)$ by reformulating the problem to only include dot products $\phi(x)\T\phi(x')$, which are then replaced by kernels $k(x,x')$.
-Lets illustrate this process on various examples in the following.
+Lets illustrate this process by deriving kernel PCA in the following.
 
@@
 \eqnsnn{
-X\T X  v_j = \lambda v_j \quad\equivalent\quad X\T XX\T  v_j &= \lambda X\T v_j 
-&&\\ \quad\equivalent\quad XX\T XX\T  v_j &= \lambda XX\T v_j
-&&\\ \quad\equivalent\quad K v_j &= \lambda v_j
+X\T X  v_j = \lambda v_j \quad\quad\equivalent\quad XX\T X v_j &= \lambda Xv_j
+&&\\\equivalent\quad XX\T XX\T\alpha_j &= \lambda XX\T \alpha_j \with v_j = X\T \alpha_j
+&&\\\equivalent\quad KK\alpha_j &= \lambda K \alpha_j
+&&\\\equivalent\quad K\alpha_k &= \lambda \alpha_j
 &&}
+Alright we can solve for the eigenvectors in kernel form, but how do we project $x$ to $z$?
+\\In feature representation we previously had
+\eqnsnn{&
+z = V_P\T \phi(x) \with V_P ~\text{composed of eigenvectors $v_j$} 
+&&}
+We now have $A_P$ composed of eigenvectors $\alpha_j$ where $v_j = X\T\alpha_j$, which means that $V_P = 
+X\T A_P$. When substituting this we get:
+\eqnsnn{&
+z = (X\T A_P)\T \phi(x) = A_P\T\; X\;\phi(x) = A_P\T\;\kappa(x) \with \kappa(x) = X \phi(x) = \mat{k(x,x_1)\\\vdots\\k(x,x_N)}
+&&}
+Mission completed, we got rid of all occurences of features $\phi$ and instead use kernel values $k(x,x')$.
+
+\section{Clustering}
+Clustering is used to guess classes for data points (unlabeled so there is no ground truth $\neq$ classification).
+So the goal is to find $K$ classes and assign the data points to them in a fashion that similar data is in the same cluster.
+
+Two well known clustering algorithm are K-Means and the Gaussian Mixture Model.
+Lets briefly discuss K-Means before diving into GMM.
+
+\subsection{K-Means}
+In K-Means, we define clusters by centroids and cluster membership by minimal distant centroid.
+So a point belongs to the cluster which has the closest centroid to it.
+The loss that is minimized is:
+\eqnsnn{
+L(\mu) = \sum_{i=1}^N \sum_{k=1}^K C_{ik}*||x_i-\mu_k||^2
+&&}
+With data points $x_i \in \dom{R}^M$ centroids $\mu_k \in dom{R}^M$ and membership matrix $C$ (one entry of a row is $1$ others are $0$).
+The algorithm works like this:
+\begin{enumerate}
+\item initialize centeroids $\mu$ randomly
+\item calculate C (set matrix entries $C_{ik}$ to one where corresponding $\mu_k$ is clostest to $x_i$).
+\item calculate new centroids as $\mu_k = \frac1{\sum_{i=1}^N C_{ik}} \sum_{i=1}^N C_{ik} x_i$
+\item repeat from (2) until centroids are no longer changing.
+\end{enumerate}
+
+This algorithm will only find local minima, which is why we usually do several runs to hopefully find the global minimum or at least a quite good local minimum.
+The main problem tough is to find a reasonable $K$.
+Of course, the loss decreases with higher $K$, when $K=N$ then every point is its own cluster and the loss is $0$.
+Usually we examine the change of loss for increasing $K$ and stop increasing when the change is not substantial anymore.
+
+\subsection{Gaussian Mixture Model}
+In GMM instead of $K$ centroids we have $K$ gaussians from which we want to describe the whole distribution of points.
+Each gaussian has a weight called \enquote{mixing value} $\pi_k$, its mean $\mu_k$ and covariance matix $\Sigma_k$.
+The whole distribution of all points is given by:
+\eqnsnn{&
+p(x) = \sum_{k=1}^K \pi_k * \mathcal{N}(x|\mu_k,\Sigma_k) \with 0 \leq\pi_k \leq 1 \mathtext{and} \sum_{k=1}^K \pi_k = 1
+&&}
+The loss that is minimized is the neg-log-likelihood of the data (equivalent to maximizing likelihood):
+\eqnsnn{&
+L(\mu,\Sigma,\pi) = -\sum_{i=1}^N\ln\Big(\sum_{k=1}^K * \pi_k * \mathcal{N}(x_i|\mu_k,\Sigma_k)\Big)
+&&}
+The algorithm works similar to K-Means but more parameters need to be updated:
+\begin{enumerate}
+\item Initialize mixing values uniformly $\pi_{1:K} = 1/K$, covariances to identity $\Sigma_{1:K}=I$ and means $\mu_{1_K}$ randomly.
+\item Calculate the posterior probabilities that point $x_i$ belongs to cluster $k$, $P_{ik} = \mathcal{N}(x_i|\mu_k,\Sigma_k)*\pi_k/Z_i$ with normalization $Z_i=\sum_{k'=1}^N\mathcal{N}(x_i|\mu_{k'},\Sigma_{k'})*\pi_{k'}$.
+\item Sum up the posteriors for each cluster $N_k = \sum_{1=1}^N P_{ik}$.
+\item Update the mixing values $\pi_k = N_k/N$.
+\item Update the means $\mu_k = (1/N_k) * \sum_{i=1}^N P_{ik} x_i$.
+\item Update the covariances $\Sigma_k = (1/N_k) \sum_{i=1}^N P_{ik} * (x_i-\mu_k)(x_i-\mu_k)\T$.
+\item Repeat from (2) until convergence.
+\end{enumerate}
+The main difference to K-Means is, that we do not have exclusive memberhips $C_{ik}$ but membership probabilities $P_{ik}$ which is like a soft assignment of points to clusters.
 

### 2018-08-06 seminar_vof/paper/paper.tex
sum more stuff
adaptive schemes completed

@@
 The sigmoid slope parameter $\lambda$ now works as a weight for the selection term $\phi$ which has become independent of $\lambda$.
-In fact the selection term has simplified quite a bit and it can be seen that it is a symmetric function which is also illustarted in \cref{fig:c_phi}.
+In fact the selection term has simplified quite a bit and it can be seen that it is a symmetric function which is illustarted in \cref{fig:c_phi}.
 From the figure it can also be seen that $\phi$ is minimal for $c=\frac{1}{2}$ which can be interpreted as a prior on $c$ for when the model is indecisive on regularization order.
@@
 Here $\mathcal{N}(\x)$ denotes the patch of $\x$ and $|\mathcal{N}(\x)|$ the patch size which is used for normalization.
-To find $\cnl$, $R_\text{nl}(\flow,\cnl)$ is minimized with respect to it.
+To find $\cnl$, $R_\text{nl}(\flow,\cnl)$ is minimized with respect to it and yields:
 \begin{flalign}
-\frac{\partial R(\flow,\cnl)}{\partial \cnl} = \cdots
-&&\\
-\text{derivation of~} \cnl \dots
-&&\\
-\cnl = \cdots
+\cnl(\y) = \frac{1}{1+e^{\Delta(\y)/\lambda}} \quad\text{with}\quad \Delta(\y) = \int_{\mathcal{N}(\y)} \frac{ T + S_2 - S_1 }{|\mathcal{N}(\x)|} \;d\x
 &&
 \end{flalign}
+It may be surprising at first that the integrated $\cnl$ itself contains a nested integral, but it is easy to convince yourself when thinking about a 
+discrete version $c_{\x}$.
+When taking the derivative with respect to a single $c_{\x}$, then due to the integral in $\cNl$ we need to identify all terms which contain $c_{\x}$ (which is defined by the patch $\mathcal{N}(\x)$).
+The reason for normalization $\frac{1}{|\mathcal N(\x) |}$ being inside the integral is that patches at the boundaries vary in size (become smaller as less pixels are defined there).
 
@@
 The non-local scheme takes a neighborhood around a pixel into account for the descision on regularization order.
-To push it yet a little further, the region based scheme decides on order leveraging a level set function $z(\x):\Omega\to\mathbb{R}$ which devides the image space into first and second order regularized regions.
+To push it yet a little further, the region based scheme decides on order leveraging a level-set function $z(\x):\Omega\to\mathbb{R}$ which devides the image space into first and second order regularized regions.
+Recall that a level-set function defines a closed curve (or curves) by $\{\x \;|\; z(\x)=0\}$ that defines the descision boundary.
+The regions are thus defined by $\{\x \;|\; z(\x) < 0\}$ and $\{\x \;|\; z(\x) > 0\}$ which will be leveraged to regionally decide on regularization order.
 The corresponding regularizer reads as follows.
@@
 Here $c_\text{rgn}$ is as well a sigmoid function, but choosen to approximate the heaviside funtion ($\lambda\to 0$) in order to obtain hard boundaries for the regions defined by $z$.
-The selection term from the previous schemes has been replaced by a smoothness term on $c_\text{rgn}$ which penalizes level set functions with lots of sign changes.
-Due to this penalization short paths for $z(\x)=0$ are favored resulting in smooth region boundaries.
+The selection term from the previous schemes has been replaced by a smoothness term on $c_\text{rgn}$ which penalizes non-smooth level set functions which results in less sign changes.
+Due to this penalization short paths for $\{\x \;|\; z(\x)=0\}$ are favored resulting in smooth region boundaries.
+
+Unfortunately this sophisticated scheme comes at a larger computational cost as $z$ has to be evolved due to its smoothenss term \cite{someone}.
+
+\section{Experiments}
+To evaluate this new approach, the authors conducted a few experiments which will be presented in this section.
 
@@
 \end{document}
+\grid

### 2018-08-07 seminar_vof/paper/paper.tex
experiments section started
more experiments stuff

@@
 
+To get an insight into the order selection of the algorithm with the different selection schemes, four synthetic image sequences were created from which the actual flow fields could be extracted as ground truth.
+\Cref{fig:classroom} shows the 4 image pairs and their corresponding flow fields.
+
+\begin{figure*}[htb]
+\centering
+\includegraphics[width=0.9\linewidth]{images/classroom.png}
+\caption{\textbf{Classroom sequences}~\cite{daspaper}. 
+First and second rows contains first frames and second frames.
+Third row contains the corresponding flow fields with flow direction encoded by hue, and flow magnitude by brightness.
+Each column corresponds to another sequence.
+Sequences 1,2 and 4 result from fronto-parallel motion as can be seen from the flow directions being the same for each pixel.
+Sequence 3 results from \enquote{fronto-orthogonal} or affine motion similar to the linear flow field depicted in \cref{fig:flowfields}.}
+\label{fig:classroom}
+\end{figure*}
+
+By looking at the gradient magnitude $\big|\deriv{\flow(\x)}{\x}\big|$ of the ground truth flow field, areas of vanishing gradients can be identified.
+Small values of the gradient magnitude thus indicate fronto-parallel motion whereas large values indicate affine motion.
+When these are compared to the order deciding $c$ term of the different schemes, we can get a good impression on how the algorithm decided.
+
+\begin{figure*}[htb]
+\centering
+\includegraphics[width=0.9\linewidth]{images/classroom_c.png}
+\caption{Adaption behavior of the different schemes~\cite{daspaper}. 
+Each column corresponds to the respective sequence introduced in \cref{fig:classroom}.
+First row shows the gradient magnitude of the ground truth flow fields, dark pixels indicate small magnitudes,
+bright pixels high magnitudes.
+The following rows correspond to the regularization order selection parameters $c$, dark pixels indicate first order regularization, bright pixels indicate second order regularization.
+The second row corresponds to the global scheme, third to local, fourth to non-local and fifth to region based. 
+}
+\label{fig:classroom_c}
+\end{figure*}
+
+\Cref{fig:classroom_c} shows a comparison of the adaption behavior of the four different adaption schemes with respect to the ground truth gradient magnitude which works as an indicator for regularization order.
+For the global scheme (2nd row) the images are either completely white or completely black as the selection is global and thus the same for each pixel.
+For the global scheme, the algorithm surprisingly picked second order regularization for all sequences except for the fourth sequence.
+On the other hand, the fourth sequence has almost no depth to it and is thus truly fronto-parallel, so all schemes decided for first order for the majority of the pixels.
+What can also be seen is that for the local scheme (third row) the selection is pretty noisy, whereas the non-local and region based scheme show way more consistent order selection.
+Another insight is, that at motion discontinuities (i.e. edges) first order regularization is chosen predominantely.
+Finally it is important to note that different order selection than indicated by the ground truth does not necessarily imply a bad result, since second order regularization contains first order as previously explained.
+
+The quality of the approach has not only been aseesed visually, but also quantitatively by evaluating the average end point error (AEE).
+The four regularizers (global , local, non-local, region) therefore were put up against a purely first and purely second order regularizer on the class room sequences.
+The second order regularizer used for comparison is an anisotropic variant of total generalized variation (TGV~\cite{tgv}) which also uses a coupling term and is not a purposefully chosen inferior regularizer.
+The AEE results for the different sequences are shown in \cref{tbl:classroom_aee}.
+
+\begin{table}[htb]
+%% Table captions on top in journal version
+\caption{\label{tab:vis_accept} Average Endpoint Error - Classroom sequences}
+\scriptsize
+\begin{center}
+\begin{tabular}{r|cccc|c}
+regularization & Seq. 1 & Seq. 2 & Seq. 3 & Seq. 4 & runtime 
+\\\hline
+first order        & 0.129 & 0.358 & 2.038 & \textbf{0.088} & 17s
+\\
+second order       & 0.141 & 0.370 & 0.669 & 0.102 & 75s
+\\\hline
+adaptive global    & 0.141 & 0.365 & 0.667 & 0.095 & 100s
+\\
+adaptive local     & \textbf{0.111} & \textbf{0.260} & 1.115 & \textbf{0.088} & 105s
+\\
+adaptive non-local & 0.116 & 0.275 & 0.737 & 0.095 & 120s
+\\
+adaptive region    & 0.125 & 0.366 & \textbf{0.662} & 0.098 & 180s
+\end{tabular}
+\end{center}
+\end{table}
+
 \section{Exposition}

### 2018-08-07 seminar_vof/paper/paper.tex
benchmark table

@@
 The AEE results for the different sequences are shown in \cref{tbl:classroom_aee}.
+From the table it can be seen that the best results are found in the order adaptive segment except for the fourth sequence where the first order regularizer shares first place with the adaptive local regularizer.
+Also the second order regularizer is already outperformed by the global order adaptive regularizer.
+For the rather fronto-parallel sequences, the local adaptive regularizer shines, but has really poor performance on the third sequence which is heavily affine in motion.
+The non-local and region order adaptive regularizers on the other hand, do quite well in all of the sequences, the region order adaptive however performs best on the affine motion sequence.
+Also note that the global order adaptive regularizer does not simply amount to choose best out of first and second order, but instead relies on the same hyperparamters as the other adaptive regularizers (which were optimized jointly).
+Due to the number variables to be evolved in order to approximate the flow ($\flow, \aaux, \baux, c$) the runtimes of the adaptive regularizers are quite high compared to the simple first order regularizer which does not have a coupling term (and thus no auxiliary functions), especially the the region adaptive order regularizer takes quite a while as it involves the evolution of a smooth level set function.
 
@@
 %% Table captions on top in journal version
-\caption{\label{tab:vis_accept} Average Endpoint Error - Classroom sequences}
+\caption{\label{tbl:classroom_aee} Average endpoint error - Classroom sequences~\cite{daspaper}}
 \scriptsize
@@
 adaptive region    & 0.125 & 0.366 & \textbf{0.662} & 0.098 & 180s
+\end{tabular}
+\end{center}
+\end{table}
+
+Of course the results on the self made image sequences are not easily put into perspective with other recent approaches regarding optic flow, which is why the authors also put the regularizers up against the Middlebury benchmark~\cite{middlebury}, the Sintel benchmark~\cite{sintel} and the KITTI benchmarks of 2012~\cite{kitti12} and 2015~\cite{kitti15}.
+
+\begin{table}[htb]
+%% Table captions on top in journal version
+\caption{\label{tbl:benchmarks} Average endpoint error (AEE) and bad pixel error (BP) in four different benchmarks~\cite{daspaper}}
+\scriptsize
+\begin{center}
+\begin{tabular}{r|cccc}
+regularization & Middlebury & Sintel & KITTI '12 & KITTI '15
+\\
+& (AEE) & (AEE) & (BP) & (BP)
+\\\hline
+first order        & 0.213 & 4.327 & 18.026 & 30.053
+\\
+second order       & 0.222 & 6.518 & 9.461 & 22.736
+\\\hline
+adaptive global    & 0.211 & 4.213 & 9.423 & 22.424
+\\
+adaptive local     & 0.211 & 4.082 & 11.537 & 24.938
+\\
+adaptive non-local & 0.211 & 4.145 & 9.468 & 22.158
+\\
+adaptive region    & 0.208 & 4.358 & 9.415 & 22.343 
 \end{tabular}

### 2018-08-09 seminar_vof/paper/paper.tex
written conclusion

@@
 \\
-adaptive local     & 0.211 & 4.082 & 11.537 & 24.938
+adaptive local     & 0.211 & \textbf{4.082} & 11.537 & 24.938
 \\
-adaptive non-local & 0.211 & 4.145 & 9.468 & 22.158
+adaptive non-local & 0.211 & 4.145 & 9.468 & \textbf{22.158}
 \\
-adaptive region    & 0.208 & 4.358 & 9.415 & 22.343 
+adaptive region    & \textbf{0.208} & 4.358 & \textbf{9.415} & 22.343 
 \end{tabular}
@@
 
-\section{Exposition}
+The results obtained from these benchmark underline the findings from the experiments with the class room sequences.
+Again the non-local and region order adaptive regularizer perform pretty good overall.
+The reason for the local order adaptive scoring so high in the sintel benchmark is the fact that the sintel benchmark consists mainly on fronto-parallel motion~\cite{sintel}.
+For the KITTI benchmarks the local order adaptive regularizer yields worse results than the simple second order and other adaptive regularizers.
+So even on benchmark data, it can be seen that order adaption is worthwhile as it can provide better results than comparable single order approaches.
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum~\cite{ware:2004:IVP}. Stet clita kasd gubergren, no sea takimata
-sanctus est Lorem ipsum dolor sit amet.
-
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat~\cite{kindlmann:1999:SAG}.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo
-consequat~\cite{levoy:1989:DSV}. Duis autem vel eum iriure dolor in
-hendrerit in vulputate velit esse molestie consequat, vel illum dolore
-eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
-dignissim qui blandit praesent luptatum zzril delenit augue duis
-dolore te feugait nulla facilisi.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=1.5in]{sample}
-  \caption{\label{fig:sample} Beispielillustration.}
-\end{figure}
-
-\begin{equation}
-  \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
-\end{equation}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-\begin{table}
-  %% Table captions on top in journal version
-  \caption{\label{tab:vis_accept} Vis Paper Acceptance Rate}
-  \scriptsize
-  \begin{center}
-    \begin{tabular}{cccc}
-      Year & Submitted & Accepted & Accepted (\%)\\
-      \hline
-      1994 &  91 & 41 & 45.1\\
-      1995 & 102 & 41 & 40.2\\
-      1996 & 101 & 43 & 42.6\\
-      1997 & 117 & 44 & 37.6\\
-      1998 & 118 & 50 & 42.4\\
-      1999 & 129 & 47 & 36.4\\
-      2000 & 151 & 52 & 34.4\\
-      2001 & 152 & 51 & 33.6\\
-      2002 & 172 & 58 & 33.7\\
-      2003 & 192 & 63 & 32.8\\
-      2004 & 167 & 46 & 27.6\\
-      2005 & 268 & 88 & 32.8\\
-      2006 & 228 & 63 & 27.6
-    \end{tabular}
-  \end{center}
-\end{table}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-\subsection{Mezcal Head}
-
-Duis autem~\cite{Lorensen:1987:MCA} vel eum iriure dolor in hendrerit
-in vulputate velit esse molestie consequat, vel illum dolore eu
-feugiat nulla facilisis at vero eros et accumsan et iusto odio
-dignissim qui blandit praesent luptatum zzril delenit augue duis
-dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
-consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
-ut laoreet dolore magna aliquam erat volutpat%
-\footnote{Fu"snoten erscheinen an der Unterkante der Spalte. Sie
-  sollten jedoch vermieden werden, da der Lesefluss gest"ort wird.}.
-
-\subsubsection{Ejector Seat Reservation}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo
-consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
-hendrerit in vulputate velit esse molestie consequat, vel illum dolore
-eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
-dignissim qui blandit praesent luptatum zzril delenit augue duis
-dolore te feugait nulla facilisi.
-
-\paragraph{Rejected Ejector Seat Reservation}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
+The authors decided on submitting the results of the non-local order adaptive regularizer to the respective benchmarks for comparison against other approaches of found in literature.
+It showed that the novel approach yields comparable or even better results than other first order approaches as for example Zimmer~et~al.~\cite{opticflowinharmony} on the Sintel and Middlebury benchmarks.
+For the KITTI benchmarks the approach as well yields results that can compete with other second order approaches, like for example non-local total generalized variation (NLTGV) by Ranftl~et~al.~\cite{nltgv}.
 
 \section{Conclusion}
+In this paper, the approach to variational optic flow estimation by Maurer~et~al.~\cite{daspaper} was explained.
+First the problem of exclusive first or second order regularization was explained, which is the suitability of the regularization order that is situation dependent.
+It was explained that first order regularization is well suited for fronto-parallel motion, whereas second order works for affine motion.
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
+The variational model of the novel approach was then introduced which uses a dataterm that combines a brightness and gradient constancy assumption and an order adaptive regularization term.
+The different parts of the regularization term were explained, which consists of an anisotropic first order part and anisotropic second order part that relies on the coupling of the gradients to two separate auxiliary functions.
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum.
+Following that, the global adaptive scheme that selects order for the whole image domain was derived and three other adaptive schemes were shown (local, non-local, region based).
+
+In the end, the novel regularizers were put to the test on synthetic sequences created by the authors which showed the adaption behavior of the four different schemes.
+The reulsts of the regularizers on various benchmarks were presented as well.
+It showed that allowing for adaptively choosing regularization order is actually beneficial and yields a flexible variational model that performs well in both domains, that is fronto-parallel sequences and affine motion sequences.
 

### 2018-08-09 seminar_vof/paper/paper.tex
written abstract and filled in citations

@@
 \abstract{%
-  Hier sollte eine kurze Zusammenfassung der Ausarbeitung stehen. Der
-  Umfang sollte 150 bis 250 Worte betragen. Duis autem vel eum iriure
-  dolor in hendrerit in vulputate velit esse molestie consequat, vel
-  illum dolore eu feugiat nulla facilisis at vero eros et accumsan et
-  iusto odio dignissim qui blandit praesent luptatum zzril delenit
-  augue duis dolore te feugait nulla facilisi. Lorem ipsum dolor sit
-  amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod
-  tincidunt ut laoreet dolore magna aliquam erat volutpat.
+  In variational optic flow estimation, the flow field between two consecutive images of a sequence is estimated by minimizing an energy functional.
+  Usually this functional uses a regularization term in order to avoid unreasonable kinds of flow fields and allow for a unique solution to the ill-posed problem.
+  Often the regularization is formulated as a constraint in terms of vanishing flow derivative of either first or second order which makes an algorithm suitable for different kinds of motion to be estimated.
 
-  Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-  suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis
-  autem vel eum iriure dolor in hendrerit in vulputate velit esse
-  molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-  vero eros et accumsan et iusto odio dignissim qui blandit praesent
-  luptatum zzril delenit augue duis dolore te feugait nulla facilisi.
+  This paper discusses the recent approach by Maurer~et~al.
+on adaptively selecting regularization order to allow for a flexible method that can solve typical first and second order flow problems equally well.
+  This approach also overcomes framewise exclusive regularization order, which is desireable as fronto-parallel and affine motions may be present at the same time. 
 } % end of abstract
@@
 In computer vision, the field of optic flow deals with the estimation of a 2-dimensional flow field based on two consecutive frames of an image sequence.
-\\(mehr dazu warum optic flow geil ist)\\
-Different methods for flow estimation exist, such as ... ... and variational optic flow which is based on the calculus of variations.
-In the following the concept of variational flow estimation is recapitulated and the approach by Maurer, Stoll and Bruhn \cite{daspaper} with order adaptive regulalarization is presented.
+This is useful for the detection and tracking of objects, e.g. in robotics~\cite{robotics}, or in video compression tasks~\cite{videocoding}.
+Different methods for flow estimation exist, Barron~et~al. classified the different approaches into differential, matching, energy-based and phase-based methods~\cite{opticflowtechniques}.
+The differential methods are more commonly known as variational methods as they rely on the calculus of variations.
+Typically variational models use an energy functional consisting of a data term and a regularization term which are to be minimized.
+The regularization term is used to overcome ambigous solutions to the optic flow problem by placing constraints on the flow field.
+Commonly the regularization terms are either of first or second order which allows for different kinds of flow fields to evolve.
+In the following the concept of variational flow estimation is recapitulated and the variational approach by Maurer, Stoll and Bruhn~\cite{daspaper}, which allows for adapting regularization order, is motivated and presented.
 
@@
 
-Unfortunately this sophisticated scheme comes at a larger computational cost as $z$ has to be evolved due to its smoothenss term \cite{someone}.
+Unfortunately this sophisticated scheme comes at a larger computational cost as $z$ has to be evolved due to its smoothenss term (compare~\cite{someone}).
 
@@
 Again the non-local and region order adaptive regularizer perform pretty good overall.
-The reason for the local order adaptive scoring so high in the sintel benchmark is the fact that the sintel benchmark consists mainly on fronto-parallel motion~\cite{sintel}.
-For the KITTI benchmarks the local order adaptive regularizer yields worse results than the simple second order and other adaptive regularizers.
+The local order adaptive regularizer scores high in the sintel benchmark yields rather poor results in the KITTI benchmarks where the simple second order and other adaptive regularizers worked quite well.
 So even on benchmark data, it can be seen that order adaption is worthwhile as it can provide better results than comparable single order approaches.

### 2018-08-09 seminar_vof/paper/paper.tex
some typos, more citations

@@
 The optic flow in this situation would be constant as all cars move in the same direction parallel to the image plane.
-Now imagine the car driver moving forwards as the crosisng is now free, the whole scene around him will appear to move towards him.
+Now imagine the car driver moving forwards as the crossing is now free, the whole scene around him will appear to move towards him.
 The optic flow is now linear as the surroundings move perpendicular to the image plane.
@@
 Lets postpone the discussion of the regularization tensor for a second and get some intuition on the way $S_1$ works.
-As can be seen from \cref{eq:S1}, the term uses two penalizer functions, the Charbonnier penalizer which was introduced in \cref{sec:dataterm} in \cref{eq:charbonnier} and the Perona~Malik penalizer $\peronamalik$.
+As can be seen from \cref{eq:S1}, the term uses two penalizer functions, the Charbonnier penalizer which was introduced in \cref{sec:dataterm} in \cref{eq:charbonnier} and the Perona~Malik penalizer $\peronamalik$~\cite{peronamalik}.
 \begin{flalign}\label{eq:peronamalik}
